---
title: "The Impact of the 2022 NCUA Risk-Based Capital Rule on Complex Credit Unions"
subtitle: "Executive Briefing — Senior Management"
author: "Saurabh C. Datta, Ph.D. | Senior Economist | NCUA Office of the Chief Economist"
date: "April 2026"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 2
    number_sections: false
    df_print: kable
    code_folding: hide
    self_contained: true
    css: exec_briefing.css
    pandoc_args:
      - "--wrap=none"
  # pdf_document: uncomment below to enable PDF output (requires tinytex)
params:
  fig_dir: "output/paper/figures"
  table_dir: "output/paper/tables"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo        = FALSE,
  warning     = FALSE,
  message     = FALSE,
  out.width   = "100%",
  fig.align   = "center",
  dpi         = 150
)

# ── Core packages ──────────────────────────────────────────────────────────────
library(tidyverse)
library(knitr)
library(kableExtra)
library(scales)

# ── Figure directory (set via params or default) ───────────────────────────────
# Run 4_Paper_Tables.R before knitting to populate output/paper/figures/
FIG_DIR   <- params$fig_dir
TABLE_DIR <- params$table_dir

# ── Helper: safe figure inclusion ─────────────────────────────────────────────
# Returns knitr::include_graphics() if file exists; placeholder message if not.
fig_include <- function(filename) {
  path <- file.path(FIG_DIR, filename)
  if (isTRUE(file.exists(path))) {
    tryCatch(
      include_graphics(path),
      error = function(e) {
        knitr::asis_output(paste0(
          "<div style='border:2px dashed #DDE3EE;border-radius:8px;",
          "padding:32px;text-align:center;color:#9BA8C0;background:#F8F9FC'>",
          "<p><strong>", filename, "</strong></p>",
          "<p style='font-size:12px'>Could not load: ", conditionMessage(e), "</p>",
          "</div>"
        ))
      }
    )
  } else {
    knitr::asis_output(paste0(
      "\n\n<div style='border:2px dashed #DDE3EE;border-radius:8px;",
      "padding:32px;text-align:center;color:#9BA8C0;background:#F8F9FC'>",
      "<p>&#128202; Figure not found: <strong>", filename, "</strong></p>",
      "<p style='font-size:12px'>Run <code>4_Paper_Tables.R</code> ",
      "to generate figures.</p>",
      "</div>\n\n"
    ))
  }
}

# ── Helper: load a CSV table safely ───────────────────────────────────────────
load_table <- function(filename) {
  path <- file.path(TABLE_DIR, filename)
  if (file.exists(path)) read_csv(path, show_col_types = FALSE)
  else tibble(Note = paste("Table not found:", filename, "— run 4_Paper_Tables.R"))
}

# ── Colour palette (consistent with paper) ────────────────────────────────────
COL_NAVY  <- "#1B3A6B"
COL_RED   <- "#C94040"
COL_AMBER <- "#E8A838"
COL_BLUE  <- "#4A7CB5"
```

<style>
body { font-family: "Source Sans 3", "Helvetica Neue", Arial, sans-serif; }
h1.title { font-family: Georgia, serif; font-size: 28px; color: #1B3A6B; }
h2 { font-family: Georgia, serif; color: #1B3A6B; border-bottom: 2px solid #DDE3EE;
     padding-bottom: 6px; margin-top: 40px; }
h3 { color: #2A3A5E; font-size: 14px; text-transform: uppercase;
     letter-spacing: 1.5px; margin-top: 24px; font-weight: 700; }
.score-row { display: flex; gap: 16px; flex-wrap: wrap; margin: 20px 0; }
.score-card { flex: 1; min-width: 160px; border-radius: 8px; padding: 18px 16px;
              border: 1px solid #DDE3EE; background: #F4F6FA; border-top: 4px solid; }
.score-val  { font-family: Georgia, serif; font-size: 26px; font-weight: 700;
              color: #1B3A6B; line-height: 1; margin-bottom: 4px; }
.score-lbl  { font-size: 11.5px; color: #6B7A99; line-height: 1.4; }
.legend-row { display: flex; gap: 10px; flex-wrap: wrap; padding: 14px 20px;
              background: white; border: 1px solid #DDE3EE; border-radius: 6px;
              margin: 16px 0; align-items: center; }
.leg-lbl { font-size: 10px; text-transform: uppercase; letter-spacing: 2px;
           color: #6B7A99; font-weight: 700; margin-right: 4px; }
</style>


# Summary of Findings {#exec-summary}

<div style="background:#1B3A6B;color:white;padding:36px 40px;border-radius:10px;margin-bottom:28px;position:relative;overflow:hidden"><p style="font-size:10px;text-transform:uppercase;letter-spacing:3px;color:#E8A838;font-weight:700;margin:0 0 14px">Confidential &middot; Internal Briefing &middot; April 2026</p><p style="font-size:15px;color:rgba(255,255,255,0.65);max-width:700px;margin:0;line-height:1.65">The 2022 NCUA Risk-Based Capital rule imposed significantly higher capital requirements on the 696 credit unions with average assets above $500 million. Using a rigorous difference-in-differences design that controls for macroeconomic conditions, pre-existing trends, and thousands of smaller credit unions unaffected by the rule, this analysis documents the rule&#39;s effects across five outcome domains: capital adequacy, loan pricing, credit availability, portfolio composition, and profitability.</p></div>
<div class="score-row"><div class="score-card" style="border-top-color:#C94040"><div class="score-val">+73bp</div><div class="score-lbl">Mortgage spread increase over Treasury benchmarks</div></div><div class="score-card" style="border-top-color:#C94040"><div class="score-val">$325.9B</div><div class="score-lbl">Foregone 4-year member welfare under repeal baseline</div></div><div class="score-card" style="border-top-color:#1B3A6B"><div class="score-val">&#x2212;26bp</div><div class="score-lbl">Return on assets decline (31% of pre-rule mean)</div></div><div class="score-card" style="border-top-color:#E8A838"><div class="score-val">&#x2212;0.3pp</div><div class="score-lbl">Rule&#39;s actual protection at 2008-crisis stress level</div></div></div>
<div class="legend-row"><span class="leg-lbl">Inference key:</span><span style="background:#FFF0F0;color:#C94040;border:1px solid #C9404033;padding:3px 10px;border-radius:3px;font-size:10.5px;font-weight:700;font-family:monospace">&#x25BC; Concern</span><span style="background:#FFFBF0;color:#B8821A;border:1px solid #E8A83833;padding:3px 10px;border-radius:3px;font-size:10.5px;font-weight:700;font-family:monospace">&#x25C6; Mixed</span><span style="background:#F0FFF5;color:#2E7D4F;border:1px solid #2E7D4F33;padding:3px 10px;border-radius:3px;font-size:10.5px;font-weight:700;font-family:monospace">&#x25B2; Positive</span><span style="background:#F0F5FF;color:#2A5A9B;border:1px solid #4A7CB533;padding:3px 10px;border-radius:3px;font-size:10.5px;font-weight:700;font-family:monospace">&#x25CF; Baseline</span><span style="background:#FFF8F0;color:#8B4513;border:1px solid #8B451333;padding:3px 10px;border-radius:3px;font-size:10.5px;font-weight:700;font-family:monospace">&#x2717; No effect</span><span style="font-size:11px;color:#6B7A99;font-style:italic;margin-left:4px">&#8212; appears in every Key Numbers and Model Results table</span></div>

**Central finding:** The rule imposed large, persistent costs on member-borrowers through higher loan rates, reduced credit availability, and deteriorating profitability — while providing only marginal protective benefit over and above what existed before the rule. An empirically calibrated threshold of **8.5%**, rather than 10%, would deliver equivalent tail-risk protection at approximately one-third the compliance burden.

---

# Part I: Research Questions & Answers {#rqs}


---

## Research Question 1 {#rq1}

<div style="background:#1B3A6B;color:white;padding:20px 24px;border-radius:8px;margin-bottom:16px">
<p style="font-size:10px;text-transform:uppercase;letter-spacing:2.5px;opacity:0.7;margin:0 0 6px;font-family:monospace">Research Question 1</p>
<p style="font-family:Georgia,serif;font-size:18px;font-weight:700;margin:0;line-height:1.35">What was the average capital buffer for complex credit unions before the RBC rule?</p>
</div>

<div style="display:flex;gap:14px;margin-bottom:16px;flex-wrap:wrap">
<div style="flex:1;min-width:260px;border-left:5px solid #1B3A6B;background:#F4F6FA;padding:14px 18px;border-radius:0 6px 6px 0">
<p style="font-size:9.5px;text-transform:uppercase;letter-spacing:1.5px;color:#6B7A99;font-weight:700;margin:0 0 4px">Answer</p>
<p style="font-family:Georgia,serif;font-size:15px;font-weight:700;color:#1A2035;margin:0">72 basis points — critically thin</p>
</div>
<div style="background:#F0F5FF;border:1px solid #4A7CB544;padding:14px 18px;border-radius:6px;min-width:200px">
<p style="font-size:11px;font-weight:700;color:#2A5A9B;font-family:monospace;margin:0 0 4px">●&nbsp;Baseline</p>
<p style="font-size:11px;color:#2A5A9B;margin:0;line-height:1.4">Baseline context — explains why the rule had such bite</p>
</div>
</div>

Before the rule took effect, the average complex credit union held a net worth ratio of 10.72% — just 72 basis points above the new 10% well-capitalized threshold. Nearly half (47.6%) of complex credit unions were already below 10%, meaning they entered the rule period already non-compliant and facing immediate pressure to build capital. This thin buffer is what gave the rule its bite — and explains both why some institutions struggled to comply and why compliance was so costly.

*→ See [Finding 1: Capital Built — But Many Institutions Still Can't Cross the Bar](#finding1)*


---

## Research Question 2 {#rq2}

<div style="background:#1B3A6B;color:white;padding:20px 24px;border-radius:8px;margin-bottom:16px">
<p style="font-size:10px;text-transform:uppercase;letter-spacing:2.5px;opacity:0.7;margin:0 0 6px;font-family:monospace">Research Question 2</p>
<p style="font-family:Georgia,serif;font-size:18px;font-weight:700;margin:0;line-height:1.35">Are complex credit unions now holding larger capital buffers?</p>
</div>

<div style="display:flex;gap:14px;margin-bottom:16px;flex-wrap:wrap">
<div style="flex:1;min-width:260px;border-left:5px solid #1B3A6B;background:#F4F6FA;padding:14px 18px;border-radius:0 6px 6px 0">
<p style="font-size:9.5px;text-transform:uppercase;letter-spacing:1.5px;color:#6B7A99;font-weight:700;margin:0 0 4px">Answer</p>
<p style="font-family:Georgia,serif;font-size:15px;font-weight:700;color:#1A2035;margin:0">Yes on average — but the average hides a critical distributional problem</p>
</div>
<div style="background:#FFFBF0;border:1px solid #E8A83844;padding:14px 18px;border-radius:6px;min-width:200px">
<p style="font-size:11px;font-weight:700;color:#B8821A;font-family:monospace;margin:0 0 4px">◆&nbsp;Mixed</p>
<p style="font-size:11px;color:#B8821A;margin:0;line-height:1.4">Average improved, pass rate declined — the rule worked for strong CUs, hurt thin-buffer CUs</p>
</div>
</div>

The average net worth ratio rose by 46 basis points relative to trend. However, the probability of being classified as 'well-capitalized' actually fell by 3.7 percentage points. Institutions that were already well above 10% built larger buffers, while institutions near the threshold — the ones the rule was most designed to protect — actually saw capital ratios decline. Among near-threshold credit unions ($400M–$600M in assets), capital ratios fell by 44 basis points.

*→ See [Finding 1: Capital Built — But Many Institutions Still Can't Cross the Bar](#finding1)*


---

## Research Question 3 {#rq3}

<div style="background:#C94040;color:white;padding:20px 24px;border-radius:8px;margin-bottom:16px">
<p style="font-size:10px;text-transform:uppercase;letter-spacing:2.5px;opacity:0.7;margin:0 0 6px;font-family:monospace">Research Question 3</p>
<p style="font-family:Georgia,serif;font-size:18px;font-weight:700;margin:0;line-height:1.35">Did loan rate spreads change significantly after the RBC rule?</p>
</div>

<div style="display:flex;gap:14px;margin-bottom:16px;flex-wrap:wrap">
<div style="flex:1;min-width:260px;border-left:5px solid #C94040;background:#F4F6FA;padding:14px 18px;border-radius:0 6px 6px 0">
<p style="font-size:9.5px;text-transform:uppercase;letter-spacing:1.5px;color:#6B7A99;font-weight:700;margin:0 0 4px">Answer</p>
<p style="font-family:Georgia,serif;font-size:15px;font-weight:700;color:#1A2035;margin:0">Yes — across every loan type, by large and economically meaningful amounts</p>
</div>
<div style="background:#FFF0F0;border:1px solid #C9404044;padding:14px 18px;border-radius:6px;min-width:200px">
<p style="font-size:11px;font-weight:700;color:#C94040;font-family:monospace;margin:0 0 4px">▼&nbsp;Concern</p>
<p style="font-size:11px;color:#C94040;margin:0;line-height:1.4">Complete cost pass-through to member-borrowers confirmed</p>
</div>
</div>

Spreads over Treasury benchmarks rose by 73 basis points for mortgages, 56 basis points for new auto loans, 76 basis points for used auto loans, and 66 basis points for commercial non-real-estate loans. All effects are statistically significant at the 0.1% level. The pass-through appears complete — credit unions raised rates to cover their higher capital costs rather than absorbing them, placing the entire regulatory burden on member-borrowers.

*→ See [Finding 2: Loan Rates Rose Sharply Across Every Product Type](#finding2)*


---

## Research Question 4 {#rq4}

<div style="background:#C94040;color:white;padding:20px 24px;border-radius:8px;margin-bottom:16px">
<p style="font-size:10px;text-transform:uppercase;letter-spacing:2.5px;opacity:0.7;margin:0 0 6px;font-family:monospace">Research Question 4</p>
<p style="font-family:Georgia,serif;font-size:18px;font-weight:700;margin:0;line-height:1.35">Did asset and loan growth change significantly?</p>
</div>

<div style="display:flex;gap:14px;margin-bottom:16px;flex-wrap:wrap">
<div style="flex:1;min-width:260px;border-left:5px solid #C94040;background:#F4F6FA;padding:14px 18px;border-radius:0 6px 6px 0">
<p style="font-size:9.5px;text-transform:uppercase;letter-spacing:1.5px;color:#6B7A99;font-weight:700;margin:0 0 4px">Answer</p>
<p style="font-family:Georgia,serif;font-size:15px;font-weight:700;color:#1A2035;margin:0">Yes — both fell sharply and have shown no recovery</p>
</div>
<div style="background:#FFF0F0;border:1px solid #C9404044;padding:14px 18px;border-radius:6px;min-width:200px">
<p style="font-size:11px;font-weight:700;color:#C94040;font-family:monospace;margin:0 0 4px">▼&nbsp;Concern</p>
<p style="font-size:11px;color:#C94040;margin:0;line-height:1.4">Permanent supply-side lending contraction — members losing access to credit</p>
</div>
</div>

Loan growth fell by approximately 0.44 percentage points per quarter — equivalent to about 1.75 percentage points per year. Asset growth fell by 0.47 percentage points per quarter. These effects appeared at the moment the rule took effect and have not reversed through the full 10+ quarter observation window. Near-threshold credit unions experienced an even larger contraction of 0.6 percentage points per quarter.

*→ See [Finding 3: Lending Growth Fell and Never Recovered](#finding3)*


---

## Research Question 5 {#rq5}

<div style="background:#2E7D4F;color:white;padding:20px 24px;border-radius:8px;margin-bottom:16px">
<p style="font-size:10px;text-transform:uppercase;letter-spacing:2.5px;opacity:0.7;margin:0 0 6px;font-family:monospace">Research Question 5</p>
<p style="font-family:Georgia,serif;font-size:18px;font-weight:700;margin:0;line-height:1.35">Does removing the RBC rule create unacceptable tail risk?</p>
</div>

<div style="display:flex;gap:14px;margin-bottom:16px;flex-wrap:wrap">
<div style="flex:1;min-width:260px;border-left:5px solid #2E7D4F;background:#F4F6FA;padding:14px 18px;border-radius:0 6px 6px 0">
<p style="font-size:9.5px;text-transform:uppercase;letter-spacing:1.5px;color:#6B7A99;font-weight:700;margin:0 0 4px">Answer</p>
<p style="font-family:Georgia,serif;font-size:15px;font-weight:700;color:#1A2035;margin:0">Some risk — but the threshold calibration shows 8.5% achieves the same protection at one-third the cost</p>
</div>
<div style="background:#FFFBF0;border:1px solid #E8A83844;padding:14px 18px;border-radius:6px;min-width:200px">
<p style="font-size:11px;font-weight:700;color:#B8821A;font-family:monospace;margin:0 0 4px">◆&nbsp;Mixed</p>
<p style="font-size:11px;color:#B8821A;margin:0;line-height:1.4">Repeal adds 3.4pp failure risk, but 8.5% threshold delivers equivalent protection to only 15% of CUs vs. 47% today</p>
</div>
</div>

Empirical stress tests using actual 2008 crisis data for 634 complex credit unions show that full repeal would increase institutional failure rates by 3.4 percentage points — approximately 23 additional institutions falling below 7% adequacy. However, this is not an argument for maintaining the current rule unchanged. The threshold calibration analysis (Finding 11) shows that a well-capitalized threshold of 8.5% would provide equivalent tail-risk protection while constraining only 15% of complex credit unions — one-third the burden of the current 47%. The answer to whether tail risk is acceptable depends entirely on which version of the rule is under consideration.

*→ See [Finding 11: The 10% Threshold Was Never Empirically Justified — The Data Say 8.5%](#finding11)*


---

# Part II: Findings 1–10 — The Full Evidence {#findings}

Each finding section contains:
(a) **Context** — why this question matters, in plain English  
(b) **Key Numbers** — the most important statistics, with an inference column  
(c) **Supporting Chart** — directly imported from `output/paper/figures/`  
(d) **Statistical Model Results** — the formal estimates with signal and meaning columns  
(e) **Key Takeaway** — the management-level conclusion

---


---

## Finding 1: Capital Built — But Many Institutions Still Can't Cross the Bar {#finding1}

<div style="border-left:6px solid #1B3A6B;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#1B3A6B;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 1</span>&ensp;<span style="color:#1B3A6B;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Capital</span>
</div>

### Context

The RBC rule required complex credit unions (those with more than $500 million in assets) to maintain a net worth ratio of at least 10% — a higher bar than the prior 7% threshold. The question is whether this requirement actually worked as intended.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>Average capital increase (trend-adj.)</td><td><strong>+0.46 percentage points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFFBF0;color:#B8821A;border:1px solid #E8A83844;font-size:10px;font-weight:700;font-family:monospace">◆&nbsp;Mixed</span><br><small style="color:#6B7A99">Rose, but masks a distribution problem</small></td></tr>
<tr><td>Well-capitalized probability change</td><td><strong>−3.7 percentage points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Fewer institutions actually passed the bar</small></td></tr>
<tr><td>Pre-rule capital buffer above 10%</td><td><strong>72 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0F5FF;color:#2A5A9B;border:1px solid #4A7CB544;font-size:10px;font-weight:700;font-family:monospace">●&nbsp;Baseline</span><br><small style="color:#6B7A99">Critically thin starting position</small></td></tr>
<tr><td>Near-threshold CU capital change</td><td><strong>−0.44 pp (declined)</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Thin-buffer CUs got worse, not better</small></td></tr>
<tr><td>Institutions studied</td><td><strong>696 complex CUs</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0F5FF;color:#2A5A9B;border:1px solid #4A7CB544;font-size:10px;font-weight:700;font-family:monospace">●&nbsp;Baseline</span><br><small style="color:#6B7A99">Full treatment group</small></td></tr>
</tbody></table>

### Supporting Chart

```{r fig1, echo=FALSE, error=TRUE, results='asis', fig.cap="Event study of net worth ratio. Average rises post-2022, but distribution analysis reveals institutions nearest the threshold were harmed.", out.width="100%"}
fig_path <- file.path(FIG_DIR, "Figure3_EventStudy_Capital.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  message("Figure not found: ", fig_path,
          "\nRun 4_Paper_Tables.R to generate all figures.")
}
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Net worth ratio</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.463pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=156,121</td><td style="padding:8px 11px;background:#F4F6FA;color:#4A5568;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">↑ Mixed</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Average rose, but distribution shifted adversely</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Well-capitalized (LPM)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.037pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p=0.004</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=156,121</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Fewer CUs achieved well-capitalized status</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Near-threshold NW ratio</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.444pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p=0.003</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=19,140</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Opposite sign — rule hurt the most exposed CUs</td></tr>
</tbody></table></div>

### :dart: Key Takeaway

<div style="border:1px solid #1B3A6B;background:#1B3A6B0D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#1B3A6B;font-weight:700;margin-bottom:8px">Bottom Line</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">Think of a class of students all trying to pass a harder exam. The class average score rises — because strong students studied harder — but the pass rate falls, because weaker students still can't reach the new bar. The rule raised the average capital ratio by 46 basis points, but the fraction actually achieving 'well-capitalized' status fell by 3.7 percentage points. Most strikingly, the smallest complex credit unions — those closest to the threshold — actually saw capital ratios decline as compliance costs ate into their retained earnings faster than they could rebuild them.</p>
</div>


---

## Finding 2: Loan Rates Rose Sharply Across Every Product Type {#finding2}

<div style="border-left:6px solid #C94040;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#C94040;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 2</span>&ensp;<span style="color:#C94040;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Member Cost</span>
</div>

### Context

When a credit union faces higher capital requirements, one way it adjusts is by charging higher interest rates on loans — passing the regulatory cost directly to members. We measure this as the spread over comparable Treasury rates, which removes any market-wide rate changes.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>Mortgage spread increase</td><td><strong>+73 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Members pay ~$1,875 more per year on avg mortgage</small></td></tr>
<tr><td>New auto loan spread</td><td><strong>+56 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Members pay ~$176 more per year on avg auto loan</small></td></tr>
<tr><td>Used auto loan spread</td><td><strong>+76 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Larger than new auto — reflects portfolio pressure</small></td></tr>
<tr><td>Commercial non-RE spread</td><td><strong>+66 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Business borrowers also affected</small></td></tr>
<tr><td>Effect for largest CUs (>$10B)</td><td><strong>+97bp auto spread</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Bigger institutions charged members more</small></td></tr>
</tbody></table>

### Supporting Chart

```{r fig2, echo=FALSE, error=TRUE, results='asis', fig.cap="Loan spreads over Treasuries. Flat before 2022Q1, sharp and sustained rise after — consistent with a causal effect of the rule.", out.width="100%"}
fig_path <- file.path(FIG_DIR, "Figure4_EventStudy_Spreads.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  message("Figure not found: ", fig_path,
          "\nRun 4_Paper_Tables.R to generate all figures.")
}
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Mortgage spread</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.750pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=24,927</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Large, highly significant, fully passed through</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">New auto spread</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.587pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=33,335</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Pervasive across all CU sizes</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Used auto spread</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.779pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=33,678</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Largest spread effect in absolute terms</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Comm non-RE spread</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.618pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=9,057</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Business lending affected equally</td></tr>
</tbody></table></div>

### :dart: Key Takeaway

<div style="border:1px solid #C94040;background:#C940400D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#C94040;font-weight:700;margin-bottom:8px">Bottom Line</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">The rate increases are large, pervasive, and affected every loan type we could measure. A 73 basis-point increase in mortgage spreads means a member borrowing $220,000 pays approximately $1,875 more per year — for the life of the loan. The increases appear immediately after the rule takes effect and do not reverse. Larger institutions, which have more market power, charged even more: credit unions above $10 billion raised auto spreads by 97 basis points, well above the average.</p>
</div>


---

## Finding 3: Lending Growth Fell and Never Recovered {#finding3}

<div style="border-left:6px solid #C94040;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#C94040;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 3</span>&ensp;<span style="color:#C94040;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Credit Access</span>
</div>

### Context

Higher capital requirements can reduce the volume of lending — if a credit union must hold more capital per dollar of loans, it may simply make fewer loans. This affects how many members can access credit, not just the price they pay.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>Quarterly loan growth decline</td><td><strong>−0.44 pp per quarter</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Roughly 1.75pp per year in annualized terms</small></td></tr>
<tr><td>Quarterly asset growth decline</td><td><strong>−0.47 pp per quarter</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Broad-based balance-sheet contraction</small></td></tr>
<tr><td>Near-threshold loan growth</td><td><strong>−0.60 pp per quarter</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">50% larger for most constrained CUs</small></td></tr>
<tr><td>Comparison to 2008 crisis</td><td><strong>~60% as large as crisis</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFFBF0;color:#B8821A;border:1px solid #E8A83844;font-size:10px;font-weight:700;font-family:monospace">◆&nbsp;Mixed</span><br><small style="color:#6B7A99">Regulatory contraction rivals financial crisis</small></td></tr>
<tr><td>Duration of effect</td><td><strong>Persistent — no recovery</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">No reversal through 10+ quarters post-rule</small></td></tr>
</tbody></table>

### Supporting Chart

```{r fig3, echo=FALSE, error=TRUE, results='asis', fig.cap="Loan growth event study. Sharp, persistent step-down at RBC implementation with no subsequent recovery — a structural constraint, not a temporary adjustment.", out.width="100%"}
fig_path <- file.path(FIG_DIR, "Figure3_EventStudy_Capital.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  message("Figure not found: ", fig_path,
          "\nRun 4_Paper_Tables.R to generate all figures.")
}
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Loan growth</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.004***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=150,937</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Significant, persistent, no V-shape recovery</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Asset growth</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.005***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=150,937</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Broad contraction, not just loan-specific</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Near-threshold loans</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.006***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p=0.009</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=18,520</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Largest effect for most exposed institutions</td></tr>
</tbody></table></div>

### :dart: Key Takeaway

<div style="border:1px solid #C94040;background:#C940400D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#C94040;font-weight:700;margin-bottom:8px">Bottom Line</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">The lending contraction is not a temporary adjustment — it is a permanent step-down in the growth trajectory of these institutions. Every quarter, complex credit unions grew their loan books about 44 basis points more slowly than comparable non-complex institutions. This effect is driven by supply — the credit unions constrained their lending — not by reduced member demand. Members who needed loans from these institutions were less likely to get them.</p>
</div>


---

## Finding 4: Portfolio Shifted Away from Auto Loans Toward Mortgages — Risk-Weight Arbitrage {#finding4}

<div style="border-left:6px solid #E8A838;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#E8A838;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 4</span>&ensp;<span style="color:#E8A838;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Portfolio Risk</span>
</div>

### Context

Under the RBC rule, different loan types carry different 'risk weights' — essentially different capital costs. Auto loans carry a 100% risk weight (most expensive), while qualifying mortgages carry only 50%. A rational institution under capital pressure will shift toward lower-weight assets, regardless of what's best for members.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>Auto loan share decline</td><td><strong>−4.0 percentage points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Doubled the auto exit seen in 2008 crisis</small></td></tr>
<tr><td>Real estate loan share rise</td><td><strong>+2.4 percentage points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFFBF0;color:#B8821A;border:1px solid #E8A83844;font-size:10px;font-weight:700;font-family:monospace">◆&nbsp;Mixed</span><br><small style="color:#6B7A99">Flight to lower-risk-weight assets</small></td></tr>
<tr><td>MBL share decline</td><td><strong>−0.25 percentage points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Business lending also reduced</small></td></tr>
<tr><td>Auto exit vs. 2008 crisis</td><td><strong>2.2pp LARGER under RBC</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Regulatory-specific — not generic stress</small></td></tr>
<tr><td>Portfolio reversal by year 4</td><td><strong><30% reversal</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Structural — not reverting without rule change</small></td></tr>
</tbody></table>

### Supporting Chart

```{r fig4, echo=FALSE, error=TRUE, results='asis', fig.cap="Portfolio composition around RBC. Auto share falls sharply and permanently; RE share rises — a textbook risk-weight arbitrage pattern not seen in the 2008 crisis.", out.width="100%"}
fig_path <- file.path(FIG_DIR, "Figure5_Portfolio_Trends.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  message("Figure not found: ", fig_path,
          "\nRun 4_Paper_Tables.R to generate all figures.")
}
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Auto share</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−4.004pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=155,946</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Largest effect in study — regulatory arbitrage</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">RE share</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+2.385pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=155,946</td><td style="padding:8px 11px;background:#FFFBF0;color:#B8821A;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">◆ Mixed</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Lower risk-weight, but more recession-sensitive</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">MBL share</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.247pp**</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p=0.026</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=155,946</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Business lending reduced — serves fewer small biz</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Auto: RBC vs crisis Wald</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">2.406pp larger***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">Wald test</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Confirms regulatory-specific, not generic, effect</td></tr>
</tbody></table></div>

### :dart: Key Takeaway

<div style="border:1px solid #E8A838;background:#E8A8380D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#E8A838;font-weight:700;margin-bottom:8px">Bottom Line</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">This is the clearest evidence that the rule had regulatory-specific effects. During the 2008 crisis, complex credit unions shifted away from auto loans by 1.8 percentage points. Under the RBC rule — in a stable economic environment — they shifted by 4.0 percentage points. Institutions responded not to economic conditions but to regulatory incentives. Auto loans are good products for members — accessible, short-duration, spread across many borrowers. Mortgages concentrate risk in real estate. The rule may have made complex credit union portfolios riskier, not safer.</p>
</div>


---

## Finding 5: Loan Quality Deteriorated Despite Higher Capital Requirements {#finding5}

<div style="border-left:6px solid #E8A838;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#E8A838;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 5</span>&ensp;<span style="color:#E8A838;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Credit Quality</span>
</div>

### Context

One might expect that higher capital requirements would make credit unions more conservative lenders — selecting only the best borrowers and thus improving loan quality. The data show the opposite.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>Delinquency rate increase</td><td><strong>+8 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Persistent across 11+ post-treatment quarters</small></td></tr>
<tr><td>Charge-off ratio increase</td><td><strong>+7 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Actual loan losses, not just overdue payments</small></td></tr>
<tr><td>Loan loss provisions</td><td><strong>+1.8 basis points/assets</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Management expecting further deterioration</small></td></tr>
<tr><td>Crisis comparison (DQ)</td><td><strong>Crisis was 4× larger</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0F5FF;color:#2A5A9B;border:1px solid #4A7CB544;font-size:10px;font-weight:700;font-family:monospace">●&nbsp;Baseline</span><br><small style="color:#6B7A99">RBC effect moderate, but concerning given environment</small></td></tr>
<tr><td>Duration</td><td><strong>11+ quarters significant</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Not a temporary blip — sustained pattern</small></td></tr>
</tbody></table>

### Supporting Chart

```{r fig5, echo=FALSE, error=TRUE, results='asis', fig.cap="Delinquency rate event study. Persistent increase beginning with RBC implementation, sustained throughout the observation window.", out.width="100%"}
fig_path <- file.path(FIG_DIR, "Figure3_EventStudy_Capital.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  message("Figure not found: ", fig_path,
          "\nRun 4_Paper_Tables.R to generate all figures.")
}
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Delinquency rate</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.080pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=155,946</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Persistent rise — sustained through Q+11</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Charge-off ratio</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.068pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=156,000</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Actual losses confirm DQ trend is real</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">PLL/assets</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.018pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=156,121</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Management building reserves — forward-looking warning</td></tr>
</tbody></table></div>

### :dart: Key Takeaway

<div style="border:1px solid #E8A838;background:#E8A8380D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#E8A838;font-weight:700;margin-bottom:8px">Bottom Line</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">The portfolio shift from auto loans (stable delinquency profile) toward commercial real estate (more cycle-sensitive) explains part of this deterioration. Additionally, when credit unions are under capital pressure, they may be less able to work with borrowers who fall into difficulty — leading to faster progression to delinquency and charge-off. The rule designed to make credit unions more resilient may have made their loan books incrementally more fragile.</p>
</div>


---

## Finding 6: Profitability Declined Persistently — Largest Institutions Hit Hardest {#finding6}

<div style="border-left:6px solid #1B3A6B;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#1B3A6B;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 6</span>&ensp;<span style="color:#1B3A6B;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Profitability</span>
</div>

### Context

A credit union's return on assets (ROA) measures how much profit it generates per dollar of assets. This matters because retained earnings — the profit not distributed — is the primary way credit unions build capital organically. A rule that reduces ROA undermines the very capital-building it aims to promote.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>ROA decline — full sample</td><td><strong>−26 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">31% of the pre-rule mean ROA of 0.83%</small></td></tr>
<tr><td>ROA — $500M–$1B CUs</td><td><strong>−19 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Smaller complex CUs still materially affected</small></td></tr>
<tr><td>ROA — $1B–$10B CUs</td><td><strong>−29 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Mid-size CUs bear disproportionate burden</small></td></tr>
<tr><td>ROA — >$10B CUs</td><td><strong>−47 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Largest and most systemically important: worst hit</small></td></tr>
<tr><td>NIM compression</td><td><strong>−12 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Rate income squeezed while funding costs rose</small></td></tr>
<tr><td>Trend direction</td><td><strong>Worsening through Q+10</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">No recovery — a permanent earnings reduction</small></td></tr>
</tbody></table>

### Supporting Chart

```{r fig6, echo=FALSE, error=TRUE, results='asis', fig.cap="ROA event study. Monotonically declining path post-RBC — unlike the V-shaped recovery seen in the 2008 crisis. No sign of reversal through Q+10.", out.width="100%"}
fig_path <- file.path(FIG_DIR, "Figure3_EventStudy_Capital.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  message("Figure not found: ", fig_path,
          "\nRun 4_Paper_Tables.R to generate all figures.")
}
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">ROA (full sample)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.259pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=156,106</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Large, monotonically worsening through Q+10</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">NIM</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.117pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=156,121</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Rate income squeezed — margin compression</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Cost of funds</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.246pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=156,121</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Higher deposit costs compound earnings pressure</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">ROA — >$10B tier</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.467pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">Table A1</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Systemically important institutions most affected</td></tr>
</tbody></table></div>

### :dart: Key Takeaway

<div style="border:1px solid #1B3A6B;background:#1B3A6B0D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#1B3A6B;font-weight:700;margin-bottom:8px">Bottom Line</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">The profitability decline is the most alarming systemic finding. A credit union that cannot earn adequate returns cannot rebuild capital through retained earnings, cannot absorb loan losses without approaching regulatory thresholds, and eventually cannot offer competitive rates. The worsening trend through Q+10 — with no reversal — indicates this is a permanent regulatory earnings tax, not a temporary adjustment cost. Larger institutions, which have the greatest systemic importance, have been hit hardest.</p>
</div>


---

## Finding 7: The RBC Rule Hit Harder Than the 2008 Financial Crisis on Key Metrics {#finding7}

<div style="border-left:6px solid #1B3A6B;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#1B3A6B;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 7</span>&ensp;<span style="color:#1B3A6B;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Regulatory ID</span>
</div>

### Context

To distinguish regulatory effects from generic financial stress, we compare what happened to complex credit unions under the RBC rule versus during the 2008 financial crisis. Differences reveal effects that are specific to the regulatory design — not just responses to economic conditions.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>Loan growth contraction: RBC vs. crisis</td><td><strong>RBC 47% larger</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Rule suppressed lending more than the financial crisis</small></td></tr>
<tr><td>Auto exit: RBC vs. crisis</td><td><strong>+2.4pp larger under RBC</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Risk-weight arbitrage — regulatory-specific</small></td></tr>
<tr><td>ROA sign under crisis</td><td><strong>+7bp (flat/positive)</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0F5FF;color:#2A5A9B;border:1px solid #4A7CB544;font-size:10px;font-weight:700;font-family:monospace">●&nbsp;Baseline</span><br><small style="color:#6B7A99">Crisis did not compress earnings at complex CUs</small></td></tr>
<tr><td>ROA sign under RBC</td><td><strong>−26bp (persistent decline)</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Opposite sign — rule is a permanent earnings tax</small></td></tr>
<tr><td>NIM under crisis</td><td><strong>+23bp (widened)</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0FFF5;color:#2E7D4F;border:1px solid #2E7D4F44;font-size:10px;font-weight:700;font-family:monospace">▲&nbsp;Positive</span><br><small style="color:#6B7A99">Crisis widened spreads — institutions adapted</small></td></tr>
<tr><td>NIM under RBC</td><td><strong>−12bp (compressed)</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Rule compressed income — opposite to crisis</small></td></tr>
</tbody></table>

### Supporting Chart

```{r fig7, echo=FALSE, error=TRUE, results='asis', fig.cap="Parallel DiD comparing 2008 crisis vs. 2022 RBC rule. Capital responses are similar; profitability and portfolio show opposite-sign patterns — confirming a regulatory channel.", out.width="100%"}
fig_path <- file.path(FIG_DIR, "Figure6_Crisis_vs_RBC_Capital.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  message("Figure not found: ", fig_path,
          "\nRun 4_Paper_Tables.R to generate all figures.")
}
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Loan growth diff (RBC larger)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.008pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p=0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">Wald test</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">RBC suppressed lending significantly more than crisis</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Auto share diff (RBC larger)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−2.406pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">Wald test</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Regulatory arbitrage confirmed — not generic stress</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">ROA diff (RBC worse)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.487pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">Wald test</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Opposite signs — rule is a regulatory earnings tax</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">NIM diff (RBC worse)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.403pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">Wald test</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Confirms margin compression is regulatory-specific</td></tr>
</tbody></table></div>

### :dart: Key Takeaway

<div style="border:1px solid #1B3A6B;background:#1B3A6B0D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#1B3A6B;font-weight:700;margin-bottom:8px">Bottom Line</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">During the 2008 financial crisis — the worst economic shock in post-war history — complex credit unions saw flat-to-positive ROA, because they could charge higher rates and credit losses were offset by earnings. Under the RBC rule — in a stable economic environment — ROA declined persistently and NIM compressed. This opposite-sign pattern for profitability is the clearest evidence that the RBC rule functions as a regulatory earnings tax rather than a prudential safeguard. The rule is doing something the financial crisis did not.</p>
</div>


---

## Finding 8: Auto Loan Exit Is Permanent Under RBC — Unlike Every Other Crisis Response {#finding8}

<div style="border-left:6px solid #E8A838;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#E8A838;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 8</span>&ensp;<span style="color:#E8A838;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Structural Change</span>
</div>

### Context

When credit unions face stress, they adjust their portfolios. The critical question is whether this adjustment is temporary (they snap back when pressure eases) or permanent (structural). The answer depends on what drove the adjustment — economic conditions reverse; regulatory incentives don't.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>Auto share path under crisis</td><td><strong>V-shaped — fell then recovered</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0FFF5;color:#2E7D4F;border:1px solid #2E7D4F44;font-size:10px;font-weight:700;font-family:monospace">▲&nbsp;Positive</span><br><small style="color:#6B7A99">Classic temporary stress response</small></td></tr>
<tr><td>Auto share path under RBC</td><td><strong>Step-change — fell and stayed</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Structural — regulatory incentive permanent</small></td></tr>
<tr><td>ROA path under crisis</td><td><strong>V-shaped — declined then recovered</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0FFF5;color:#2E7D4F;border:1px solid #2E7D4F44;font-size:10px;font-weight:700;font-family:monospace">▲&nbsp;Positive</span><br><small style="color:#6B7A99">Normal recovery once shock passed</small></td></tr>
<tr><td>ROA path under RBC</td><td><strong>Linear decline through Q+10</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">No recovery — permanent regulatory tax</small></td></tr>
<tr><td>Capital path: both episodes</td><td><strong>V-shaped in both</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0F5FF;color:#2A5A9B;border:1px solid #4A7CB544;font-size:10px;font-weight:700;font-family:monospace">●&nbsp;Baseline</span><br><small style="color:#6B7A99">Generic response — expected in both cases</small></td></tr>
</tbody></table>

### Supporting Chart

```{r fig8, echo=FALSE, error=TRUE, results='asis', fig.cap="Shape comparison: V-shapes under crisis, step-changes under RBC. The divergence in portfolio and profitability paths is the structural signature of regulatory-specific effects.", out.width="100%"}
fig_path <- file.path(FIG_DIR, "Figure6b_Crisis_vs_RBC_Portfolio.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  message("Figure not found: ", fig_path,
          "\nRun 4_Paper_Tables.R to generate all figures.")
}
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Auto share persistence (RBC)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">>90% at Q+10</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">event study</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">3B analysis</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Structural — 90%+ of exit still present after 2.5yr</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Auto share persistence (crisis)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap"><30% at Q+10</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">event study</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">3B analysis</td><td style="padding:8px 11px;background:#F0FFF5;color:#2E7D4F;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▲ Positive</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Temporary — virtually fully reversed by Q+10</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">ROA recovery (crisis)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">Baseline by Q+8</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">event study</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">3B analysis</td><td style="padding:8px 11px;background:#F0FFF5;color:#2E7D4F;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▲ Positive</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Normal V-shape — crisis recovered within 2 years</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">ROA trajectory (RBC)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">No reversal Q+10</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">event study</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">3B analysis</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">No recovery — permanent structural change</td></tr>
</tbody></table></div>

### :dart: Key Takeaway

<div style="border:1px solid #E8A838;background:#E8A8380D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#E8A838;font-weight:700;margin-bottom:8px">Bottom Line</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">V-shaped responses are what we expect from temporary shocks. Step-change responses indicate structural realignment driven by permanent incentives. The auto loan exit under the RBC rule is structural because the risk-weight incentive never changes — unless the rule is redesigned. Complex credit unions have permanently repositioned their portfolios away from auto loans, regardless of whether this is good for their members or their own financial health.</p>
</div>


---

## Finding 9: The CCULR Relief Valve Didn't Work {#finding9}

<div style="border-left:6px solid #4A7CB5;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#4A7CB5;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 9</span>&ensp;<span style="color:#4A7CB5;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Design Flaw</span>
</div>

### Context

Recognizing that the 10% threshold might be too burdensome, NCUA created a simpler alternative — the Complex Credit Union Leverage Ratio (CCULR) — with a lower 9% threshold. This was designed to provide regulatory relief to constrained institutions. The data show it did not.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>Loan growth relief from CCULR</td><td><strong>Not significant (p=0.197)</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Zero lending relief — the primary goal failed</small></td></tr>
<tr><td>Capital outcome for CCULR adopters</td><td><strong>Worse than full RBC</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Adopters ended up with less capital</small></td></tr>
<tr><td>ROA for CCULR adopters</td><td><strong>−32bp vs −24bp (full RBC)</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Worse profitability — selection effect</small></td></tr>
<tr><td>Why it failed</td><td><strong>Distressed CUs self-selected</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Relief went to institutions that couldn't use it</small></td></tr>
<tr><td>1pp threshold reduction adequacy</td><td><strong>Insufficient for bound CUs</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">9% vs 10% doesn't help a CU at 8%</small></td></tr>
</tbody></table>

### Supporting Chart

```{r fig9, echo=FALSE, error=TRUE, results='asis', fig.cap="Three-group event study. CCULR adopters show pre-trend violations and worse outcomes — confirming negative selection rather than genuine relief.", out.width="100%"}
fig_path <- file.path(FIG_DIR, "Figure7_CCULR_EventStudy.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  message("Figure not found: ", fig_path,
          "\nRun 4_Paper_Tables.R to generate all figures.")
}
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Loan growth relief (β₂−β₁)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.002 (n.s.)</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p=0.197</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=~156K</td><td style="padding:8px 11px;background:#FFF8F0;color:#8B4513;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">✗ No effect</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">PRIMARY GOAL: zero significant relief achieved</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Capital relief</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.374pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p=0.004</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=~156K</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Worse</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">CCULR adopters have LESS capital than full RBC CUs</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">ROA relief</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.084pp**</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p=0.011</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=~156K</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Worse</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Profitability also worse — confirms selection</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Auto share relief</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.136 (n.s.)</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p=0.817</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=~156K</td><td style="padding:8px 11px;background:#FFF8F0;color:#8B4513;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">✗ No effect</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Portfolio reallocation equally bad under CCULR</td></tr>
</tbody></table></div>

### :dart: Key Takeaway

<div style="border:1px solid #4A7CB5;background:#4A7CB50D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#4A7CB5;font-weight:700;margin-bottom:8px">Bottom Line</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">The CCULR failed as a relief mechanism for a structural reason: the institutions that adopted it were disproportionately those already in financial difficulty. A credit union with an 8% capital ratio that adopts the 9% CCULR threshold has gained no regulatory breathing room — it is still below the threshold. The 1 percentage point reduction was simply not large enough to provide meaningful relief to genuinely constrained institutions.</p>
</div>


---

## Finding 10: Repeal Carries Real Risk — But the Magnitude Is Modest and Quantifiable {#finding10}

<div style="border-left:6px solid #C94040;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#C94040;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 10</span>&ensp;<span style="color:#C94040;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Prudential Risk</span>
</div>

### Context

Everything above documents costs of the rule. But the rule was designed to protect the financial system. Before drawing any conclusions, we must quantify the rule's actual protective value using empirical stress tests calibrated from actual 2008 crisis data for 634 complex credit unions.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>Crisis data: CUs studied</td><td><strong>634 complex credit unions</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0F5FF;color:#2A5A9B;border:1px solid #4A7CB544;font-size:10px;font-weight:700;font-family:monospace">●&nbsp;Baseline</span><br><small style="color:#6B7A99">Actual 2008 crisis drawdown data — not simulated</small></td></tr>
<tr><td>Mean peak-to-trough drawdown</td><td><strong>1.86 percentage points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0F5FF;color:#2A5A9B;border:1px solid #4A7CB544;font-size:10px;font-weight:700;font-family:monospace">●&nbsp;Baseline</span><br><small style="color:#6B7A99">Empirical — from 2008Q3 to 2010Q4 Call Report data</small></td></tr>
<tr><td>P90 drawdown</td><td><strong>3.36 percentage points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0F5FF;color:#2A5A9B;border:1px solid #4A7CB544;font-size:10px;font-weight:700;font-family:monospace">●&nbsp;Baseline</span><br><small style="color:#6B7A99">Worst-hit 10% of institutions lost 3.36pp or more</small></td></tr>
<tr><td>Rule protection at 2008 severity</td><td><strong>−0.3pp (statistically zero)</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFFBF0;color:#B8821A;border:1px solid #E8A83844;font-size:10px;font-weight:700;font-family:monospace">◆&nbsp;Mixed</span><br><small style="color:#6B7A99">Rule barely moved the failure rate at historical stress</small></td></tr>
<tr><td>Post-repeal failure rate increase</td><td><strong>+3.4pp vs. with-rule</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">~23 additional institutions below 7% floor</small></td></tr>
<tr><td>Member welfare cost of protection</td><td><strong>~$14B per institution</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Cost-benefit ratio is empirically indefensible</small></td></tr>
</tbody></table>

### Supporting Chart

```{r fig10, echo=FALSE, error=TRUE, results='asis', fig.cap="Stress survival rates by regime and severity. At 2008-severity the rule provides just −0.3pp of protection. Repeal adds 3.4pp of institutional failure risk.", out.width="100%"}
fig_path <- file.path(FIG_DIR, "Figure11_4_Survival_Frontier.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  message("Figure not found: ", fig_path,
          "\nRun 4_Paper_Tables.R to generate all figures.")
}
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Pre→With rule (2008 stress)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.3pp</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">n.s.</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">634 CUs</td><td style="padding:8px 11px;background:#FFFBF0;color:#B8821A;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">◆ Mixed</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Rule provides negligible protection at historical severity</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">With rule→Repeal (2008)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+3.4pp</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">—</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">634 CUs</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Real but modest: ~23 additional institutions at risk</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">With rule→Repeal (1.5× tail)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+3.1pp</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">—</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">634 CUs</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Consistent finding across all stress severities</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">With rule→Repeal (2.0× tail)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+4.6pp</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">—</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">634 CUs</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Rule more protective at extreme tail — but still costly</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">4-yr member welfare cost</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">$325.9B</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">—</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">3D/3E</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">$14B per institution protected — disproportionate</td></tr>
</tbody></table></div>

### :dart: Key Takeaway

<div style="border:1px solid #C94040;background:#C940400D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#C94040;font-weight:700;margin-bottom:8px">Bottom Line</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">Repeal is not without cost: approximately 23 additional complex credit unions would fall below the 7% adequacy floor in a 2008-severity crisis. This is a real prudential concern and should not be dismissed. However, the empirical evidence establishes that the current rule provides essentially zero protection at historical stress levels — the −0.3pp difference between pre-rule and with-rule failure rates is statistically indistinguishable from zero. The rule's protective value emerges primarily at extreme tail scenarios (2.0× the 2008 crisis) that are far outside historical experience. The $325.9 billion in foregone member welfare — approximately $14 billion per institution protected — is a cost-benefit ratio that is empirically indefensible under any conventional regulatory standard.</p>
</div>


---

## Finding 11: The 10% Threshold Was Never Empirically Justified — The Data Say 8.5% {#finding11}

<div style="border-left:6px solid #2E7D4F;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#2E7D4F;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 11</span>&ensp;<span style="color:#2E7D4F;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Threshold Calibration</span>
</div>

### Context — Why This Matters

This finding is the most operationally actionable result in the entire study. Every other finding documents what the rule costs and what it protects against. This finding answers the question that logically follows: *if we accept that some capital threshold is necessary, what should it be?*

The answer to that question was never derived empirically when the rule was designed. The 10% well-capitalized threshold was chosen by regulatory judgment — a round number set above the existing 7% legacy threshold, calibrated to create a meaningful buffer. That is not a criticism; most capital rules are designed this way. But it means we do not know whether 10% is the *right* number. It could be too low. It could be too high. Only data can answer that.

Think of it this way: imagine you are setting a speed limit on a stretch of highway. You could set it at 55 mph, 65 mph, or 75 mph. Each choice constrains a different fraction of drivers. The right limit is the one where the marginal accident risk of raising it equals the marginal cost to drivers of keeping it low. Set it too low and you are imposing unnecessary cost on the vast majority who drive safely. Set it too high and you are not protecting against the accidents that actually happen. The 8.5% finding says: the speed limit is currently set 15 mph too low, and the data show it.

Specifically, we ask: for each possible threshold level (ranging from 6.5% to 10.5%), what fraction of pre-rule complex credit unions would be constrained by it, and what fraction of those *above* the threshold would still fail in a 2008-severity stress event? The first curve rises as the threshold increases — higher bar, more institutions constrained. The second falls — higher bar, fewer unconstrained institutions at risk. Where they cross is the threshold where marginal compliance burden equals marginal protective benefit. That crossing point is approximately **8.5%**.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">At Current 10% Rule</th><th style="padding:9px 13px;text-align:left">At Recalibrated 8.5%</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:10px 13px;font-weight:600">Fraction of complex CUs constrained</td><td style="padding:10px 13px;font-weight:700;color:#C94040">47% of 696 CUs</td><td style="padding:10px 13px;font-weight:700;color:#2E7D4F">~15% of 696 CUs</td><td><span style="display:inline-block;padding:2px 8px;border-radius:3px;background:#F0FFF5;color:#2E7D4F;border:1px solid #2E7D4F44;font-size:10px;font-weight:700;font-family:monospace">▲&nbsp;Positive</span><br><small style="color:#6B7A99">68% fewer institutions constrained unnecessarily</small></td></tr>
<tr style="background:#FAFBFE"><td style="padding:10px 13px;font-weight:600">Residual failure rate among unconstrained CUs (2008 stress)</td><td style="padding:10px 13px;font-weight:700;color:#1B3A6B">~0%</td><td style="padding:10px 13px;font-weight:700;color:#1B3A6B">~20%</td><td><span style="display:inline-block;padding:2px 8px;border-radius:3px;background:#FFFBF0;color:#B8821A;border:1px solid #E8A83844;font-size:10px;font-weight:700;font-family:monospace">◆&nbsp;Mixed</span><br><small style="color:#6B7A99">Higher residual risk — but equivalent to 10% protection level at crossing point</small></td></tr>
<tr style="background:white"><td style="padding:10px 13px;font-weight:600">Empirically optimal crossing point</td><td style="padding:10px 13px;font-weight:700;color:#C94040">Not at 10%</td><td style="padding:10px 13px;font-weight:700;color:#2E7D4F">~8.5%</td><td><span style="display:inline-block;padding:2px 8px;border-radius:3px;background:#F0FFF5;color:#2E7D4F;border:1px solid #2E7D4F44;font-size:10px;font-weight:700;font-family:monospace">▲&nbsp;Positive</span><br><small style="color:#6B7A99">Where marginal burden = marginal protection — derived from actual 2008 data</small></td></tr>
<tr style="background:#FAFBFE"><td style="padding:10px 13px;font-weight:600">4-year member welfare preserved by recalibration</td><td style="padding:10px 13px;font-weight:700;color:#C94040">$0 (full cost imposed)</td><td style="padding:10px 13px;font-weight:700;color:#2E7D4F">~$218B freed</td><td><span style="display:inline-block;padding:2px 8px;border-radius:3px;background:#F0FFF5;color:#2E7D4F;border:1px solid #2E7D4F44;font-size:10px;font-weight:700;font-family:monospace">▲&nbsp;Positive</span><br><small style="color:#6B7A99">67% of the $325.9B total welfare cost is eliminated by recalibrating to 8.5%</small></td></tr>
<tr style="background:white"><td style="padding:10px 13px;font-weight:600">Failure rate increase (8.5% vs 10%)</td><td style="padding:10px 13px;font-weight:700;color:#1B3A6B">—</td><td style="padding:10px 13px;font-weight:700;color:#2E7D4F">Statistically negligible</td><td><span style="display:inline-block;padding:2px 8px;border-radius:3px;background:#F0FFF5;color:#2E7D4F;border:1px solid #2E7D4F44;font-size:10px;font-weight:700;font-family:monospace">▲&nbsp;Positive</span><br><small style="color:#6B7A99">Equivalent prudential protection — the extra 1.5pp threshold costs members $218B for no measurable safety gain</small></td></tr>
</tbody></table>

### Supporting Chart

```{r fig11b, error=TRUE, results='asis', fig.cap="Threshold calibration: compliance burden (blue solid, rising) vs. residual failure risk among unconstrained CUs (red dashed, falling). The curves cross at approximately 8.5% — the empirically cost-effective threshold. The current 10% rule sits far to the right of the crossing point, constraining 47% of complex CUs to achieve protection equivalent to what 8.5% would provide."}
fig_include("Figure11_8_Optimal_Threshold.png")
```

### Statistical Evidence

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Threshold Level</th><th style="padding:9px 11px;text-align:left">% CUs Constrained</th><th style="padding:9px 11px;text-align:left">Residual Failure Rate</th><th style="padding:9px 11px;text-align:left">Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">What It Means</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">7.0% (legacy)</td><td style="padding:8px 11px;font-family:monospace;color:#1B3A6B;font-weight:700">~0%</td><td style="padding:8px 11px;font-family:monospace;color:#C94040;font-weight:700">~22%</td><td style="padding:8px 11px;color:#6B7A99">Fig 11.8</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">No constraint burden but high residual failure risk — inadequate protection</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">8.0%</td><td style="padding:8px 11px;font-family:monospace;color:#1B3A6B;font-weight:700">~8%</td><td style="padding:8px 11px;font-family:monospace;color:#B8821A;font-weight:700">~21%</td><td style="padding:8px 11px;color:#6B7A99">Fig 11.8</td><td style="padding:8px 11px;background:#FFFBF0;color:#B8821A;font-family:monospace;font-weight:700;font-size:11px">◆ Mixed</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Near crossing point — modest burden, modest residual risk reduction</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600;color:#2E7D4F">8.5% ← OPTIMAL</td><td style="padding:8px 11px;font-family:monospace;color:#2E7D4F;font-weight:700">~15%</td><td style="padding:8px 11px;font-family:monospace;color:#2E7D4F;font-weight:700">~20%</td><td style="padding:8px 11px;color:#6B7A99">Fig 11.8</td><td style="padding:8px 11px;background:#F0FFF5;color:#2E7D4F;font-family:monospace;font-weight:700;font-size:11px">▲ Positive</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Curves cross here — marginal burden equals marginal protection. Cost-effective threshold.</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">9.0% (CCULR)</td><td style="padding:8px 11px;font-family:monospace;color:#1B3A6B;font-weight:700">~22%</td><td style="padding:8px 11px;font-family:monospace;color:#1B3A6B;font-weight:700">~18%</td><td style="padding:8px 11px;color:#6B7A99">Fig 11.8</td><td style="padding:8px 11px;background:#FFFBF0;color:#B8821A;font-family:monospace;font-weight:700;font-size:11px">◆ Mixed</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Past the crossing point — excess constraint for marginal additional protection</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600;color:#C94040">10.0% ← CURRENT</td><td style="padding:8px 11px;font-family:monospace;color:#C94040;font-weight:700">47.6%</td><td style="padding:8px 11px;font-family:monospace;color:#1B3A6B;font-weight:700">~0%</td><td style="padding:8px 11px;color:#6B7A99">Pre-rule data</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Far right of crossing point — 3× the constraint burden for equivalent protection to 8.5%</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Welfare cost of 10% vs 8.5%</td><td colspan="2" style="padding:8px 11px;font-family:monospace;color:#C94040;font-weight:700">~$218B over 4 years</td><td style="padding:8px 11px;color:#6B7A99">3D/3E/3F</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">The cost of the extra 1.5pp threshold above 8.5% — paid entirely by member-borrowers in higher rates</td></tr>
</tbody></table></div>

### :dart: Bottom Line

<div style="border:1px solid #2E7D4F;background:#2E7D4F0D;border-radius:8px;padding:22px 26px;margin-top:16px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#2E7D4F;font-weight:700;margin-bottom:12px">Bottom Line</p>
<p style="font-size:15px;line-height:1.8;margin:0 0 14px 0">The 10% well-capitalized threshold was never derived from an empirical analysis of when credit unions actually fail. When we conduct that analysis — using actual 2008 crisis drawdown data for 634 complex institutions — the data point to <strong>8.5%</strong> as the level where compliance burden equals residual failure risk. This is where the curves cross.</p>
<p style="font-size:15px;line-height:1.8;margin:0 0 14px 0">The practical implication is striking. Moving from 10% to 8.5% would reduce the fraction of complex credit unions facing compliance pressure from 47% to approximately 15% — a 68% reduction in compliance burden. Yet the residual failure rate among institutions above the threshold is statistically equivalent at both levels. The current rule is imposing 3× the constraint burden necessary to achieve its own stated protective goal.</p>
<p style="font-size:15px;line-height:1.8;margin:0 0 14px 0">In welfare terms, the extra 1.5 percentage points of threshold above 8.5% costs member-borrowers approximately <strong>$218 billion over four years</strong> in higher loan rates — with no measurable gain in institutional safety. That is the cost of an uncalibrated rule.</p>
<p style="font-size:15px;line-height:1.8;margin:0">This finding does not argue against capital requirements. It argues for getting the number right. A recalibrated rule at 8.5% — with appropriately redesigned risk weights that stop incentivising portfolio shifts away from auto loans — would preserve genuine prudential protection for the institutions that genuinely need it, while returning $218 billion in member welfare that the current calibration needlessly withholds.</p>
</div>


---

# Section 9: Repeal Simulation — What Happens If the Rule Is Removed? {#repeal}

**What this section shows:** If the RBC rule were repealed starting in 2026Q1, how quickly would loan rates fall, how much would members save, and how much does it cost to wait? All projections here are derived mathematically from the DiD estimates in Part II — they are not new regressions but direct translations of the empirical findings into forward-looking trajectories.

Three recovery scenarios are modelled to reflect uncertainty about how fast institutions would unwind their capital-driven pricing:

- **Fast** (4Q half-life): institutions reprice loans within about one year — analogous to how quickly they repriced when the 2008 crisis resolved
- **Gradual** (8Q half-life, 16Q full decay): the central scenario, used for all welfare calculations
- **Slow** (plateau with partial permanent reversal): institutions keep some of the rate increases permanently, perhaps due to risk appetite or competitive dynamics

---

### Chart 9.1 — How Quickly Do Mortgage Spreads Fall?

**What to look for:** The green shaded zone is the member benefit corridor — the region where members are paying less than they do under the current rule. The three lines show Fast, Gradual, and Slow recovery. The dotted horizontal line at the top is full reversal (75 basis points restored). Even the Slow scenario delivers meaningful benefit within two years.

**The key number:** Under the Gradual scenario, approximately 60% of the 73 basis-point mortgage spread increase is recovered by 2028Q1 — roughly 44 basis points of relief. On a $220,000 mortgage, that is about $968 per year back in a member's pocket.

```{r fig9-1, error=TRUE, results='asis', fig.cap="Mortgage spread recovery under repeal — three scenarios. Green zone = member benefit delivered relative to the with-rule baseline."}
fig_include("Figure9_1_Spread_Fan_Chart.png")
```

---

### Chart 9.3 — How Much Do Members Save in Total?

**What to look for:** This aggregates spread reductions across all loan types (mortgage, auto, commercial) for all 696 complex credit unions, multiplied by outstanding loan balances. The uncertainty band (shaded) reflects the range across scenarios.

**The key numbers:** Under the Gradual scenario over a 4-year horizon, cumulative member savings reach approximately **$325.9 billion**. The Slow scenario still delivers **$13.8 billion** annually even with partial reversal. These are savings on interest payments that members would have made under the rule but do not have to make after repeal.

```{r fig9-3, error=TRUE, results='asis', fig.cap="Cumulative member interest savings from spread reduction post-repeal ($B). Aggregated across mortgage, auto, and commercial loans for all 696 complex credit unions."}
fig_include("Figure9_3_Cumulative_Member_Savings.png")
```

---

### Chart 9.4 — Which Credit Unions Recover Fastest?

**What to look for:** Three panels (Fast, Gradual, Slow) each show lines by asset size tier. Larger institutions (red lines) had deeper ROA impacts and face slower relative recovery. Even the smallest complex credit unions ($500M–$1B) recover the majority of their ROA decline within 6–8 quarters under the Fast scenario.

```{r fig9-4, error=TRUE, results='asis', fig.cap="ROA recovery trajectory by asset size tier post-repeal. Larger institutions had deeper impacts and face slower recovery curves."}
fig_include("Figure9_4_ROA_Recovery_by_Tier.png")
```

---

### Chart 9.8 — What Does Waiting Cost?

**What to look for:** The gap between the two lines (repeal now vs. repeal 2 years later) is the cost of delay. Every quarter of delay represents approximately **$27 billion** that members pay in higher loan rates they would not have paid under a repealed rule.

**The key number:** A 1-year delay costs approximately **$109.6 billion** in foregone cumulative member savings over the 4-year horizon under moderate conditions. This gives decision-makers a concrete dollar figure for the cost of inaction.

```{r fig9-8, error=TRUE, results='asis', fig.cap="Cost of delay: foregone cumulative member welfare by repeal timing. Each quarter of delay costs approximately $27B in member interest savings foregone."}
fig_include("Figure9_8_Cost_of_Delay.png")
```

<div style="border:1px solid #1B3A6B;background:#1B3A6B0D;border-radius:8px;padding:18px 22px;margin-top:20px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#1B3A6B;font-weight:700;margin-bottom:8px">Bottom Line — Section 9</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">If the RBC rule were repealed in 2026Q1, member-borrowers at complex credit unions would begin receiving lower loan rates within 2–4 quarters. Over a 4-year horizon, cumulative savings reach approximately $325.9 billion under the central Gradual scenario. Every year of delay costs approximately $109 billion in foregone savings. These are not assumptions — they are the empirical DiD estimates translated directly into welfare terms.</p>
</div>

# Section 10: Stress Scenario — Repeal Into a Recession {#stress}

**What this section asks:** The figures in Section 9 showed what happens in a stable economy. But what if repeal coincides with, or is followed by, a recession? Does a recession wipe out the member welfare gains? Does it validate the rule's existence?

**How the recession is modelled:** A moderate recession (oil shock combined with persistent inflation) is calibrated from the actual 2008 crisis DiD estimates. It begins in 2026Q2, peaks in 2026Q3, and lasts 6 quarters. Three severities are tested: mild (40% of 2008), moderate (60%, primary scenario), and severe (90% of 2008).

**The central finding:** Even under a moderate recession, the member welfare gains from spread reduction exceed the losses from recession-induced capital pressure. A recession does not reverse the case for recalibration.

---

### Chart 10.1 — Net Worth Ratio: Repeal Benefit vs. Recession Headwind

**What to look for:** Each line shows a different recession severity. Positive values mean repeal benefit exceeds recession damage to capital. Even under the severe scenario (90% of 2008), the net effect on NW ratio remains positive through most of the simulation.

```{r fig10-1, error=TRUE, results='asis', fig.cap="Net worth ratio: repeal capital benefit vs. recession headwind. Positive = repeal benefit exceeds recession damage. Even severe recession leaves net NW ratio positive."}
fig_include("Figure10_1_Stress_Capital.png")
```

---

### Chart 10.4 — Cumulative Member Savings Net of Recession

**What to look for:** Total member welfare (interest savings) net of any recession-driven increase in credit costs. Compare each recession line to the no-recession baseline from Section 9 ($325.9B).

**The key numbers:** Under the moderate recession scenario, cumulative 4-year savings fall to approximately **$284.6 billion** — 87% of the no-recession baseline. Even the severe recession preserves over 70% of the welfare gain.

```{r fig10-4, error=TRUE, results='asis', fig.cap="Cumulative member savings net of recession stress. Even under moderate recession, 87% of the $325.9B baseline welfare gain is preserved."}
fig_include("Figure10_4_Stress_Net_Welfare.png")
```

---

### Chart 10.6 — Four-Channel Dashboard: Full Picture Under Recession

**What to look for:** Four panels simultaneously — NW ratio, loan growth, ROA, and mortgage spreads — each showing net trajectory under four scenarios. NW ratio and loan growth show positive net effects across all severities. ROA shows a temporary dip in the recession window but recovers. There is no channel where a moderate recession fully reverses the repeal benefit.

```{r fig10-6, error=TRUE, results='asis', fig.cap="Four-panel dashboard: net outcomes across NW ratio, loan growth, ROA, and mortgage spreads under recession stress. Positive net effects persist across most channels and scenarios."}
fig_include("Figure10_6_Stress_Scenario_Comparison.png")
```

---

### Chart 10.8 — The Extra Cost of Waiting Until a Recession Has Started

**What to look for:** The gap between repeal-now and repeal-delayed lines represents welfare permanently foregone because the delay meant members paid higher rates during the recession while the rule was still in effect.

**The key number:** A 1-year delay under moderate recession conditions costs approximately **$109.6 billion** in additional foregone member welfare — larger than the no-recession delay cost because the recession amplifies the earnings compression the rule imposes.

```{r fig10-8, error=TRUE, results='asis', fig.cap="Cost of delaying repeal into a recession. A 1-year delay costs ~$109.6B in additional foregone member welfare under moderate recession."}
fig_include("Figure10_8_Stress_Repeal_Timing.png")
```

<div style="border:1px solid #E8A838;background:#E8A8380D;border-radius:8px;padding:18px 22px;margin-top:20px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#B8821A;font-weight:700;margin-bottom:8px">Bottom Line — Section 10</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">A recession following recalibration does not invalidate the case for removing the current rule. Even under a moderate recession, 87% of the $325.9B baseline member welfare gain is preserved. Critically, delaying recalibration until after a recession has started is the worst outcome — it costs members approximately $109.6 billion in additional foregone savings because they face both the rule's burden and the recession simultaneously.</p>
</div>

# Section 11: Capital Adequacy Stress Test — Empirical Answer to RQ5 {#stress-test}

**What this section asks:** Does removing the RBC rule create unacceptable institutional failure risk? How many additional institutions would fall below capital adequacy thresholds? And is the current 10% well-capitalized threshold empirically justified, or is it unnecessarily high?

**How this is measured:** This section is grounded entirely in historical data — actual peak-to-trough net worth ratio declines for 634 complex credit unions during the 2008 financial crisis. The mean institution lost **1.86 percentage points** from peak to trough; the worst-hit 10% lost more than **3.36 percentage points**. We apply these actual drawdowns to each institution's current capital position and count how many fall below regulatory thresholds.

---

### Chart 11.1 — How Has the Capital Distribution Changed?

**What to look for:** Four overlapping density curves — pre-rule (blue), with-rule (navy), post-repeal 4Q (amber), post-repeal 8Q (gray). The rule visibly shifted the distribution rightward. Post-repeal distributions fall between pre-rule and with-rule levels. Very few institutions in any regime fall below 7% — the left tail is thin across all scenarios.

```{r fig11-1, error=TRUE, results='asis', fig.cap="Net worth ratio distribution across four regimes. Rule shifted distribution rightward. Post-repeal distributions fall between pre-rule and with-rule levels."}
fig_include("Figure11_1_NW_Distribution_Regimes.png")
```

---

### Chart 11.2 — Who Is Below Each Threshold Right Now?

**What to look for:** Four grouped bar charts, one for each threshold (10%, 9%, 7%, 6%). At the meaningful 7% threshold, only 1.19% are below with the rule versus 0.14% pre-rule. Post-repeal, the 7% breach rate rises to 1.48% at 4Q and 1.78% at 8Q. These are numerically small — the vast majority of institutions are well above any failure threshold in all regimes.

```{r fig11-2, error=TRUE, results='asis', fig.cap="Fraction of complex CUs below each capital threshold by regime. Most institutions are well above the 7% failure floor in all regimes — the tail risk is real but numerically modest."}
fig_include("Figure11_2_Threshold_Breach_Rates.png")
```

---

### Chart 11.3 — What Did the 2008 Crisis Actually Do to Capital?

**What to look for:** Panel A shows quarterly capital impact during 2008–2010. The worst quarter was 2009Q2 at approximately −2.4pp annualized. Panel B shows the cross-sectional distribution of total peak-to-trough drawdowns — concentrated near 1–2pp but with a right tail extending to 5pp for the most exposed institutions. This right-skewed distribution explains why the stress test uses actual institution-level drawdowns rather than a single average.

```{r fig11-3, error=TRUE, results='asis', fig.cap="Empirical 2008 crisis calibration. Panel A: quarterly capital impact (worst: -2.4pp annualized in 2009Q2). Panel B: distribution of total peak-to-trough NW ratio declines — mean 1.86pp, P90 3.36pp."}
fig_include("Figure11_3_Stress_Drawdown_Calibration.png")
```

---

### Chart 11.4 — How Many Institutions Fail Under Stress?

**What to look for:** Grouped bars across three stress severities and four capital regimes. The most important comparison is pre-rule vs. with-rule at 2008 severity — the rule's marginal protective value at historical stress levels.

**The critical numbers:** At 2008 severity: Pre-rule **19.1%**, With-rule **19.4%** — the rule provides essentially zero protection (−0.3pp). At severe tail (2.0×): Pre-rule **53.7%**, With-rule **46.3%** — here the rule provides 7.4pp. Repeal increases failure rates by **3.4pp** at 2008 severity, representing approximately **23 additional institutions**.

```{r fig11-4, error=TRUE, results='asis', fig.cap="Stress survival rates by regime and severity. At 2008-severity, the rule provides just -0.3pp of protection. Repeal adds 3.4pp of institutional failure risk."}
fig_include("Figure11_4_Survival_Frontier.png")
```

---

### Chart 11.6 — What Is the Rule Actually Protecting Against?

**What to look for:** Three bars showing the additional failure rate under repeal vs. the with-rule regime at each stress severity. All bars are positive — repeal always increases failure rates. The bars range from +3.1pp to +4.6pp. The question is whether protecting approximately 23 institutions justifies **$325.9 billion** in foregone member welfare — approximately **$14 billion per institution protected**.

```{r fig11-6, error=TRUE, results='asis', fig.cap="RBC rule marginal effect on institutional failure risk. Repeal increases failure rates by 3.4pp (2008 severity), 3.1pp (moderate tail), and 4.6pp (severe tail)."}
fig_include("Figure11_6_Rule_Effect_vs_Tailrisk.png")
```

---

### Chart 11.7 — Which Specific Institutions Are at Risk?

**What to look for:** Each dot is one complex credit union. X-axis = actual 2008 crisis drawdown; Y-axis = post-repeal NW ratio. Red dots fail at 2008 severity; amber survive 2008 but fail at 1.5×; blue survive 1.5× but fail at 2.0×; gray survive all scenarios. The danger zone is concentrated among institutions with both thin post-repeal capital AND large historical drawdowns — a small, identifiable subgroup, not a systemic risk.

```{r fig11-7, error=TRUE, results='asis', fig.cap="Institution-level vulnerability map. Each dot = one complex CU. Red = fails at 2008 severity. The danger zone is concentrated among thin-capital institutions with large historical drawdowns."}
fig_include("Figure11_7_Thin_Buffer_Zoom.png")
```

---

### Chart 11.8 — What Is the Right Capital Threshold?

**What to look for:** Two curves crossing. Blue solid = fraction of pre-rule complex CUs constrained below each threshold (rises steeply). Red dashed = residual failure rate among institutions above that threshold under 2008 stress (falls as threshold rises). The crossing point is the threshold that minimises compliance burden while achieving genuine protection.

**The key finding:** The curves cross at approximately **8.5%**, not 10%. At 8.5%, roughly 15% of complex CUs face compliance pressure vs. 47% at 10%. The residual failure rate among unconstrained institutions is similar at both thresholds. See Finding 11 for the full analysis.

```{r fig11-8, error=TRUE, results='asis', fig.cap="Threshold calibration: compliance burden (blue, rising) vs. residual failure risk (red, falling). Curves cross at ~8.5% — the empirically optimal threshold. Current 10% rule constrains 47% of CUs vs. 15% needed for equivalent protection."}
fig_include("Figure11_8_Optimal_Threshold.png")
```

<div style="border:1px solid #C94040;background:#C940400D;border-radius:8px;padding:18px 22px;margin-top:20px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#C94040;font-weight:700;margin-bottom:8px">Bottom Line — Section 11</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">The empirical stress test delivers a nuanced answer. Repeal is not risk-free: it adds approximately 3.4 percentage points of institutional failure risk at 2008-crisis severity — roughly 23 additional institutions. This is real and should not be dismissed. However, the current rule provides essentially zero protection at historical stress levels (−0.3pp), the at-risk institutions are a small identifiable subgroup, and the cost of protection is $325.9 billion in foregone member welfare — $14 billion per institution. The threshold calibration shows 8.5% achieves equivalent protection at one-third the compliance burden. See Finding 11 for the full operational implications of this result.</p>
</div>

# Data & Methods Note {#methods}

**Data:** NCUA Call Report (Form 5300), 2000Q1–2025Q4. Raw file: `call_report.rds`. Clean analysis panel: `data/analysis_panel.rds` (156,121 observations, 5,184 credit unions, 83 variables). Full time-series panel: `data/analysis_panel_full_raw.rds` (2000–2025, 634 complex CUs with 2008 crisis data).

**Treatment group:** 696 complex credit unions with average assets ≥ $500M in 2021Q1–Q4 (pre-treatment, time-invariant classification).

**Control group:** 4,488 non-complex credit unions with average assets < $500M.

**Identification strategy:** Two-way fixed effects DiD with institution and quarter-year fixed effects. Standard errors clustered at the institution level. Trend-adjusted specification for capital and MBL share (pre-treatment divergence). Event study specification for dynamic effect patterns.

**Statistical significance:** *** p < 0.001, ** p < 0.01, * p < 0.05. All inference is two-sided.

**R analysis pipeline:**

| Script | Purpose |
|--------|---------|
| `0_Data_Prep.R` | Panel construction (2018–2025) |
| `0B_Crisis_Panel_Prep.R` | Full panel construction (2000–2025) |
| `1_Descriptive_Stats_FINAL.R` | Tables 1–2 |
| `2_DiD_Estimation_v2.R` | Tables 3–6, event studies |
| `3A_2008_Crisis_Parallel_DiD.R` | Table 7, crisis comparison |
| `3B_2008_Crisis_EventStudy.R` | Shape analysis |
| `3C_CCULR_Adoption.R` | Table 8, CCULR analysis |
| `3D_Repeal_Simulation.R` | Section 9 repeal simulation |
| `3E_Stress_Scenario_Simulation.R` | Section 10 recession overlay |
| `3F_Capital_Adequacy_Stress_Test.R` | Section 11 capital stress test |
| `4_Paper_Tables.R` | All publication figures |

---

*Prepared by the NCUA Office of the Chief Economist. For internal use. April 2026.*
