---
title: "The 2022 NCUA Risk-Based Capital Rule"
subtitle: "Empirical Findings from an Analysis of Complex Credit Unions"
author: "Saurabh C. Datta, Ph.D. -- Senior Economist, NCUA Office of the Chief Economist"
date: "April 2026"
output:
  revealjs::revealjs_presentation:
    theme: simple
    highlight: pygments
    center: false
    transition: fade
    slide_level: 2
    self_contained: true
    reveal_options:
      width: 1100
      height: 660
      margin: 0.04
      minScale: 0.5
      maxScale: 1.5
      controls: true
      progress: true
      history: true
      keyboard: true
    css: rbc_slides.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)

# Helper: include figure if it exists, otherwise show a styled placeholder
fig_include <- function(path, caption = "") {
  if (file.exists(path)) {
    knitr::include_graphics(path)
  } else {
    fname <- basename(path)
    cat(sprintf("<div class=\"fig-placeholder\"><span class=\"fig-label\">FIGURE</span><span class=\"fig-name\">%s</span><span class=\"fig-caption\">%s</span></div>", fname, caption))
  }
}

# Base figure directory -- adjust path if needed
fig_dir <- "output/paper/figures/"
```

<!-- SLIDE 1 -- TITLE -->

## {.title-slide data-background-color="#1a2744"}

<div class="title-org">NCUA &nbsp;&middot;&nbsp; Office of the Chief Economist</div>

<div class="title-main">The 2022 NCUA Risk-Based Capital Rule</div>

<div class="title-sub">Empirical Findings from an Analysis of Complex Credit Unions</div>

<div class="title-rule"></div>

<div class="title-meta">
Saurabh C. Datta, Ph.D. &nbsp;&middot;&nbsp; Senior Economist<br>
NCUA Call Report Data, 2000--2025 &nbsp;&middot;&nbsp; April 2026
</div>

<!-- SLIDE 2 -- WHAT WE STUDIED -->

## What We Studied -- and How {data-background-color="#f5f1eb"}

<div class="slide-cols">
<div class="col-wide">

<div class="label-row">The Data</div>

- **696 complex credit unions** -- average assets at or above $500M, classified using 2021 data
- **~4,488 non-complex credit unions** serve as the comparison group
- **150,000+ credit union-quarters** of NCUA Call Report data, 2000 through 2025
- **12 outcomes** measured: capital, loan rates, lending volume, portfolio mix, profitability, and loan quality

<hr>

<div class="label-row">The Method: Difference-in-Differences (DiD)</div>

<div class="method-diagram">
<div class="md-box"><strong>Change in complex CUs</strong>Before vs. after the rule</div>
<div class="md-minus">&minus;</div>
<div class="md-box"><strong>Change in non-complex CUs</strong>Same time window</div>
<div class="md-equals">=</div>
<div class="md-result"><strong>The Rule's Effect</strong>What changed because of the rule, not the economy</div>
</div>

<div class="callout">
<strong>Why this works:</strong> Non-complex credit unions faced the same economic conditions but were not subject to the RBC rule. Any difference in outcomes between the two groups, after accounting for pre-existing trends, isolates the rule's effect.
</div>

</div>
<div class="col-divider"></div>
<div class="col-narrow">

<div class="label-row">How to Read the Event Study Charts</div>

<div class="intro-box">
<ul>
<li><strong>Horizontal axis:</strong> quarters relative to the rule's effective date (quarter 0). Left = before; right = after.</li>
<li><strong>Vertical axis:</strong> estimated difference between complex and non-complex CUs, above any pre-existing gap.</li>
<li><strong>Before quarter 0:</strong> dots should sit near zero -- confirming the two groups moved together before the rule (the parallel trends check).</li>
<li><strong>After quarter 0:</strong> a persistent move up or down is the rule's estimated effect. The shaded band is the 95% confidence interval.</li>
<li><strong>If the band crosses zero</strong> at a given quarter, the effect is not statistically distinguishable from no effect.</li>
</ul>
</div>

```{r schematic-chart, fig.height=2.4, fig.width=4.2, out.width="100%"}
library(ggplot2)
library(tibble)
quarters <- c(-4, -2, 0, 2, 4, 6)
est      <- c(0.05, -0.03, 0.00, -0.45, -0.58, -0.62)
se       <- c(0.10,  0.10,  0.00,  0.10,  0.10,  0.10)
d <- tibble(q = quarters, e = est, lo = est - 1.96*se, hi = est + 1.96*se)
ggplot(d, aes(x = q, y = e)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "#999", linewidth = 0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "#c8972a", linewidth = 0.7) +
  geom_ribbon(aes(ymin = lo, ymax = hi), fill = "#1a2744", alpha = 0.15) +
  geom_line(color = "#1a2744", linewidth = 1.2) +
  geom_point(color = "#1a2744", size = 2.5, fill = "white", shape = 21, stroke = 1.5) +
  annotate("text", x = 0.2, y = 0.22, label = "Rule\neffective", hjust = 0,
           color = "#c8972a", size = 2.8, lineheight = 0.9) +
  annotate("text", x = -3.8, y = 0.22, label = "Flat pre-period (check)",
           hjust = 0, color = "#555", size = 2.6) +
  annotate("text", x = 1.8, y = -0.38, label = "Persistent effect",
           hjust = 0, color = "#1a2744", size = 2.8) +
  scale_x_continuous(breaks = quarters,
    labels = paste0(ifelse(quarters > 0, "+", ""), quarters, "Q")) +
  labs(x = NULL, y = NULL,
       caption = "Schematic -- actual charts follow on subsequent slides") +
  theme_minimal(base_size = 9) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "#e8e0d4", linewidth = 0.4),
    plot.caption     = element_text(color = "#aaa", size = 7, hjust = 0.5),
    plot.background  = element_rect(fill = "white", color = "#c8c0b0", linewidth = 0.5),
    axis.text        = element_text(color = "#666")
  )
```

</div>
</div>

<!-- SLIDE 3 -- RQ1 -->

## RQ 1 &nbsp;&middot;&nbsp; Pre-Rule Capital Buffer {data-background-color="#f5f1eb"}

<div class="rq-label">Research Question 1</div>

**How much capital buffer did complex credit unions hold before the rule?**

<div class="answer-banner">
Answer: The average complex credit union entered the rule with a capital buffer of just <strong>72 basis points</strong> above the new 10% well-capitalized threshold -- critically thin.
</div>

<div class="slide-cols">
<div class="col-wide">

<div class="label-row">What the Data Show</div>

- The 696 complex CUs were nominally compliant before the rule -- but with very little margin to absorb a new, higher threshold
- A **72 basis point buffer** means a credit union holding 10.72% net worth when the new well-capitalized definition shifted upward simultaneously
- The capital distribution was **tightly clustered just above the old threshold**, with a large mass of institutions barely on the compliant side
- This thin margin directly explains why the rule had such large downstream effects -- institutions had to act immediately to avoid falling below the new line

</div>
<div class="col-divider"></div>
<div class="col-narrow">

<div class="callout">
<strong>What is a basis point?</strong> One basis point (bp) is one one-hundredth of a percentage point. A 72bp buffer means a credit union holding 10.72% capital when the threshold is 10.00%.
</div>

<div class="stat-row">
<div class="stat-card"><div class="stat-val">72bp</div><div class="stat-lbl">Average pre-rule capital buffer above the 10% threshold</div></div>
<div class="stat-card stat-red"><div class="stat-val">696</div><div class="stat-lbl">Complex CUs subject to the rule (fixed 2021 classification)</div></div>
</div>

<div class="callout" style="margin-top:10px;">
<strong>Why it matters:</strong> A rule requiring additional capital reserves is always more costly when institutions start with little room to spare. The thin pre-rule buffer set the stage for the compliance pressures documented in RQ2 through RQ4.
</div>

</div>
</div>

<!-- SLIDE 4 -- RQ2 -->

## RQ 2 &nbsp;&middot;&nbsp; Did Capital Levels Rise? {data-background-color="#f5f1eb"}

<div class="rq-label">Research Question 2</div>

**Did the rule cause complex credit unions to hold more capital?**

<div class="answer-banner">
Answer: <strong>Mixed.</strong> Average capital levels rose -- but the share of institutions meeting the new well-capitalized standard actually fell. The bar rose faster than capital could accumulate.
</div>

<div class="slide-cols">
<div class="col-wide">

<div class="stat-row" style="margin-bottom:14px;">
<div class="stat-card stat-green"><div class="stat-val">+46bp</div><div class="stat-lbl">Increase in net worth ratio (p &lt; 0.01)</div></div>
<div class="stat-card stat-red"><div class="stat-val">-3.7pp</div><div class="stat-lbl">Change in probability of being well-capitalized (p &lt; 0.01)</div></div>
</div>

<div class="label-row">The Paradox Explained</div>

- Capital levels rose in absolute terms -- institutions responded to the rule's incentive to build reserves
- But the **definition of well-capitalized changed** at the same time, raising the threshold institutions must reach
- Result: more capital on average, but **fewer institutions clearing the new bar**

<div class="callout" style="margin-top:10px;">
<strong>An analogy:</strong> Think of a class where every student studied harder and raised their score -- but the passing grade was raised even more. The average score went up, yet more students are now technically failing. Both facts are true simultaneously.
</div>

</div>
<div class="col-divider"></div>
<div class="col-narrow">

<div class="label-row">What the Event Study Shows</div>

```{r fig-rq2, out.width="100%", results="asis"}
fig_include(
  path    = file.path(fig_dir, "Figure3_EventStudy_Capital.png"),
  caption = "4-panel: NW ratio, well-cap probability, tier 1 ratio, risk-based capital ratio"
)
```

<div class="fig-caption-note">Pre-period flat confirms parallel trends. Post-period divergence is attributed to the rule.</div>

</div>
</div>

<!-- SLIDE 5 -- RQ3 -->

## RQ 3 &nbsp;&middot;&nbsp; Member Cost: Higher Rates and Less Lending {data-background-color="#f5f1eb"}

<div class="rq-label">Research Question 3</div>

**What did the rule cost members -- in higher loan rates and reduced lending?**

<div class="answer-banner">
Answer: Loan rates rose across every product and loan growth contracted persistently. The estimated member cost totals <strong>$325.9 billion</strong> over four years -- approximately <strong>$865 per member per year</strong>.
</div>

<div class="slide-cols">
<div class="col-wide">

<div class="label-row">Loan Rate Increases vs. Non-Complex Credit Unions</div>

<div class="stat-row">
<div class="stat-card stat-red"><div class="stat-val">+75bp</div><div class="stat-lbl">Mortgage spread increase</div></div>
<div class="stat-card stat-red"><div class="stat-val">+59bp</div><div class="stat-lbl">Auto loan spread increase</div></div>
<div class="stat-card stat-red"><div class="stat-val">+58bp</div><div class="stat-lbl">Commercial spread increase</div></div>
</div>

- All increases are **statistically significant** and persist through 10+ quarters with no sign of convergence
- A 75bp mortgage increase on a $300,000 loan adds roughly **$135 per month** in payments

```{r fig-rq3-spreads, out.width="100%", results="asis"}
fig_include(
  path    = file.path(fig_dir, "Figure4_EventStudy_Spreads.png"),
  caption = "Loan spread event study -- all products"
)
```

</div>
<div class="col-divider"></div>
<div class="col-narrow">

<div class="label-row">Lending Volume and Total Member Cost</div>

<div class="stat-row">
<div class="stat-card stat-red"><div class="stat-val">-0.44pp</div><div class="stat-lbl">Quarterly loan growth reduction -- persistent, no recovery</div></div>
<div class="stat-card stat-red"><div class="stat-val">$325.9B</div><div class="stat-lbl">Total estimated member welfare cost over 4 years</div></div>
</div>

- Loan growth fell **0.44 percentage points per quarter** -- compounding over time into a substantial reduction in credit availability
- The $325.9BN estimate aggregates the rate-increase effect across all member-borrowers of all 696 complex CUs over four years
- On a per-member basis: approximately **$865 per year** in higher borrowing costs

```{r fig-rq3-growth, out.width="100%", results="asis"}
fig_include(
  path    = file.path(fig_dir, "Figure3b_EventStudy_LoanGrowth.png"),
  caption = "Loan growth event study"
)
```

</div>
</div>

<!-- SLIDE 6 -- RQ4 -->

## RQ 4 &nbsp;&middot;&nbsp; Portfolio Shift and the Crisis Comparison {data-background-color="#f5f1eb"}

<div class="rq-label">Research Question 4</div>

**Did the rule change what loans credit unions make -- and how does this compare to past stress?**

<div class="answer-banner">
Answer: Yes. Complex CUs shifted away from auto loans toward mortgages -- consistent with <strong>risk-weight arbitrage</strong>. This shift is unique to the RBC rule and did not occur during the 2008 financial crisis.
</div>

<div class="slide-cols">
<div class="col-wide">

<div class="label-row">Portfolio Reallocation</div>

<div class="stat-row" style="margin-bottom:12px;">
<div class="stat-card stat-red"><div class="stat-val">-4.0pp</div><div class="stat-lbl">Auto loan share (relative to controls)</div></div>
<div class="stat-card stat-green"><div class="stat-val">+2.4pp</div><div class="stat-lbl">Real estate loan share</div></div>
</div>

- **Risk-weight arbitrage:** auto loans carry a higher risk weight than mortgages under the RBC rule. By shifting to mortgages, complex CUs reduce their risk-weighted capital requirement -- a rational regulatory response
- Loan quality also deteriorated: **delinquency rates rose +8bp**, driven by a capacity channel -- the CUs with the largest profitability declines showed 2x the delinquency increase

```{r fig-rq4-portfolio, out.width="100%", results="asis"}
fig_include(
  path    = file.path(fig_dir, "Figure5_Portfolio_Trends.png"),
  caption = "Portfolio reallocation -- auto vs. real estate loan share"
)
```

</div>
<div class="col-divider"></div>
<div class="col-narrow">

<div class="label-row">Comparison to the 2008 Financial Crisis</div>

<div class="callout" style="margin-bottom:10px;">
We ran the same DiD analysis using 2008 as the event -- measuring how complex CUs responded to the financial crisis versus the RBC rule.
</div>

| Outcome | 2008 Crisis | RBC Rule |
|:--------|:------------|:---------|
| Capital accumulation | Similar magnitude | Similar magnitude |
| Auto loan exit | Not observed | Significant and persistent |
| Loan growth contraction | Temporary | Persistent -- no recovery |
| Profitability decline | Temporary | Persistent -- no recovery |

- Capital behavior during RBC resembles a crisis response -- but the **lending and profitability effects are distinctly regulatory** in nature

```{r fig-rq4-crisis, out.width="100%", results="asis"}
fig_include(
  path    = file.path(fig_dir, "Figure6_Crisis_vs_RBC_Capital.png"),
  caption = "Crisis vs. RBC comparison"
)
```

</div>
</div>

<!-- SLIDE 7 -- PROFITABILITY -->

## Supporting Finding &nbsp;&middot;&nbsp; Profitability {data-background-color="#f5f1eb"}

<div class="badge-intro-label">Supporting Finding</div>

**Profitability Declined Persistently -- Largest Institutions Hit Hardest**

<div class="answer-banner">
Return on assets fell across complex credit unions -- with the largest institutions experiencing the sharpest and most sustained declines. This finding is directly relevant to the financial stability implications in RQ5.
</div>

<div class="slide-cols">
<div class="col-wide">

<div class="label-row">Key Estimates</div>

<div class="stat-row" style="margin-bottom:14px;">
<div class="stat-card stat-red"><div class="stat-val">-26bp</div><div class="stat-lbl">Average ROA decline -- all complex CUs (p &lt; 0.01)</div></div>
<div class="stat-card stat-red"><div class="stat-val">-47bp</div><div class="stat-lbl">ROA decline for CUs above $10B in assets</div></div>
</div>

- The ROA decline is **persistent** across all post-rule quarters -- no evidence of adaptation or recovery through the observation window
- The largest complex CUs -- those with over $10 billion in assets -- experience a decline **nearly double the average**
- CCULR provided **no statistically significant relief** on lending outcomes; CCULR adopters showed worse baseline outcomes, consistent with distressed self-selection

</div>
<div class="col-divider"></div>
<div class="col-narrow">

<div class="callout" style="margin-bottom:10px;">
<strong>Lower profitability reduces a credit union's ability to build capital organically.</strong> A credit union that cannot retain earnings must either raise rates further, cut lending, or run down its capital ratio -- all three patterns appear in the data.
</div>

<div class="callout" style="margin-bottom:10px;">
<strong>The CCULR finding:</strong> The Community Credit Union Leverage Ratio was designed as a simpler compliance pathway. The data suggest institutions most likely to adopt it were already under financial stress -- so CCULR adoption did not cause their worse outcomes, but it also did not reverse them.
</div>

```{r fig-roa, out.width="100%", results="asis"}
fig_include(
  path    = file.path(fig_dir, "Figure3d_EventStudy_ROA.png"),
  caption = "ROA event study"
)
```

</div>
</div>

<!-- SLIDE 8 -- RQ5 -->

## RQ 5 &nbsp;&middot;&nbsp; Failure Risk from Removing the Rule {data-background-color="#f5f1eb"}

<div class="rq-label">Research Question 5</div>

**Does removing the RBC rule create unacceptable failure risk?**

<div class="answer-banner">
Answer: Removing the rule carries <strong>real but quantifiable and modest risk</strong>. At 2008 financial crisis severity, the rule reduces failure risk by 3.4 percentage points -- protecting approximately 24 institutions.
</div>

<div class="slide-cols">
<div class="col-wide">

<div class="label-row">Capital Adequacy Stress Test -- At 2008 GFC Severity</div>

<div class="stat-row" style="margin-bottom:14px;">
<div class="stat-card"><div class="stat-val">19.1%</div><div class="stat-lbl">Pre-rule failure rate -- baseline before the rule</div></div>
<div class="stat-card stat-green"><div class="stat-val">19.4%</div><div class="stat-lbl">With-rule failure rate -- rule adds -0.3pp incremental protection</div></div>
<div class="stat-card stat-red"><div class="stat-val">22.5%</div><div class="stat-lbl">Post-removal failure rate -- +3.4pp increase, approximately 24 institutions</div></div>
</div>

- The rule's **incremental protection is only -0.3pp** relative to pre-rule capital levels -- complex CUs were already building substantial buffers before the rule
- Removing the rule raises simulated failure risk by **+3.4pp at 2008 severity**, corresponding to approximately **24 additional institutions**
- Each institution protected represents approximately **$14 billion in assets** sheltered from distress

</div>
<div class="col-divider"></div>
<div class="col-narrow">

<div class="label-row">Expected Insurance Cost</div>

<div class="stat-row" style="margin-bottom:12px;">
<div class="stat-card"><div class="stat-val" style="font-size:18px;">$0.4B--$2.2B</div><div class="stat-lbl">Probability-weighted NCUSIF cost of removal</div></div>
</div>

<div class="label-row">Stress Scenario Sensitivity</div>

| Recession Severity | Welfare Retained | Member Cost |
|:-------------------|:----------------:|:-----------:|
| Moderate recession | 87% | $284BN |
| Severe recession   | 73% | $238BN |

<div class="callout" style="margin-top:10px;">
Even in a severe recession scenario, the estimated member cost of the rule substantially outweighs the estimated insurance benefit under any plausible combination of assumptions tested. No scenario produces a reversal of this finding.
</div>

```{r fig-rq5, out.width="100%", results="asis"}
fig_include(
  path    = file.path(fig_dir, "Figure9_9_Net_Benefit_Analysis.png"),
  caption = "Net benefit analysis"
)
```

</div>
</div>

<!-- SLIDE 9 -- SUMMARY -->

## Five Questions. Five Answers. {data-background-color="#f5f1eb"}

<div class="badge-summary-label">Summary</div>

```{r summary-table}
library(knitr)
library(kableExtra)

df <- data.frame(
  RQ = c("RQ 1", "RQ 2", "RQ 3", "RQ 4", "RQ 5"),
  Question = c(
    "How much capital buffer before the rule?",
    "Did complex CUs hold more capital?",
    "What did it cost members in higher rates and lower lending?",
    "Did the rule change what loans CUs make?",
    "Does removing the rule create unacceptable failure risk?"
  ),
  Finding = c(
    "Critically thin -- institutions had very little room to absorb a higher threshold",
    "Mixed -- average capital rose, but fewer institutions are now well-capitalized",
    "Rates rose on every loan product; lending growth contracted persistently and has not recovered",
    "Yes -- shifted from auto to mortgage lending (risk-weight arbitrage). Pattern did not occur during 2008.",
    "Real but modest and quantifiable -- the rule protects approximately 24 institutions at 2008 crisis severity"
  ),
  Key_Number = c(
    "72bp average buffer above 10% threshold",
    "NW ratio +46bp | Well-cap probability -3.7pp",
    "$325.9BN total | $865/member/yr | Loan growth -0.44pp/qtr",
    "Auto share -4.0pp | RE share +2.4pp",
    "Failure risk +3.4pp | ~24 institutions | NCUSIF cost $0.4-$2.2BN"
  ),
  stringsAsFactors = FALSE
)

kable(
  df,
  col.names = c("RQ", "Question", "Core Finding", "Key Number"),
  align     = c("c", "l", "l", "l"),
  escape    = TRUE
) |>
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width        = TRUE,
    font_size         = 13
  ) |>
  column_spec(1, bold = TRUE, width = "4%",  color = "#1a2744") |>
  column_spec(2, width = "22%") |>
  column_spec(3, width = "38%", bold = TRUE, color = "#1a2744") |>
  column_spec(4, width = "36%") |>
  row_spec(0, background = "#1a2744", color = "white", bold = TRUE)
```

<div class="summary-footer-row">
<div class="summary-callout">
<strong>Across all five research questions</strong>, the findings are consistent: complex credit unions responded to the rule's capital incentive, but the adjustment came primarily through higher loan rates, reduced lending volume, portfolio reallocation, and lower profitability -- rather than through material improvements in financial stability outcomes at the system level.
</div>
<div class="summary-meta-box">
<strong>Data and Methods</strong><br>
NCUA Form 5300 &nbsp;&middot;&nbsp; 2000--2025<br>
Difference-in-differences<br>
696 treated &nbsp;&middot;&nbsp; ~4,488 controls<br>
150,000+ CU-quarters<br>
12 outcome variables
</div>
</div>
