---
title: "The Impact of the 2022 NCUA Risk-Based Capital Rule on Complex Credit Unions"
subtitle: "Executive Briefing — Senior Management"
author: "Saurabh C. Datta, Ph.D. | Senior Economist | NCUA Office of the Chief Economist"
date: "April 2026"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 2
    number_sections: false
    df_print: kable
    code_folding: hide
    css: exec_briefing.css
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: false
    latex_engine: xelatex
    geometry: "margin=1in"
    fontsize: 11pt
params:
  fig_dir: "output/paper/figures"
  table_dir: "output/paper/tables"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo        = FALSE,
  warning     = FALSE,
  message     = FALSE,
  out.width   = "100%",
  fig.align   = "center",
  dpi         = 150
)

# ── Core packages ──────────────────────────────────────────────────────────────
library(tidyverse)
library(knitr)
library(kableExtra)
library(scales)

# ── Figure directory (set via params or default) ───────────────────────────────
# Run 4_Paper_Tables.R before knitting to populate output/paper/figures/
FIG_DIR   <- params$fig_dir
TABLE_DIR <- params$table_dir

# ── Helper: safe figure inclusion ─────────────────────────────────────────────
# Returns knitr::include_graphics() if file exists; placeholder message if not.
fig_include <- function(filename, caption = "") {
  path <- file.path(FIG_DIR, filename)
  if (file.exists(path)) {
    include_graphics(path)
  } else {
    cat(paste0(
      '<div style="border:2px dashed #DDE3EE;border-radius:8px;padding:32px;',
      'text-align:center;color:#9BA8C0;background:#F8F9FC">',
      "<p>&#128202; <strong>", filename, "</strong></p>",
      "<p style='font-size:12px'>Run <code>4_Paper_Tables.R</code> to generate this figure.</p>",
      "</div>"
    ))
  }
}

# ── Helper: load a CSV table safely ───────────────────────────────────────────
load_table <- function(filename) {
  path <- file.path(TABLE_DIR, filename)
  if (file.exists(path)) read_csv(path, show_col_types = FALSE)
  else tibble(Note = paste("Table not found:", filename, "— run 4_Paper_Tables.R"))
}

# ── Colour palette (consistent with paper) ────────────────────────────────────
COL_NAVY  <- "#1B3A6B"
COL_RED   <- "#C94040"
COL_AMBER <- "#E8A838"
COL_BLUE  <- "#4A7CB5"
```

```{r inline-css, results='asis'}
# Inject lightweight CSS for the inference badge system
cat('
<style>
body { font-family: "Source Sans 3", "Helvetica Neue", Arial, sans-serif; }
h1.title { font-family: Georgia, serif; font-size: 28px; color: #1B3A6B; }
h2 { font-family: Georgia, serif; color: #1B3A6B; border-bottom: 2px solid #DDE3EE;
     padding-bottom: 6px; margin-top: 40px; }
h3 { color: #2A3A5E; font-size: 14px; text-transform: uppercase;
     letter-spacing: 1.5px; margin-top: 24px; font-weight: 700; }
.score-row { display: flex; gap: 16px; flex-wrap: wrap; margin: 20px 0; }
.score-card { flex: 1; min-width: 160px; border-radius: 8px; padding: 18px 16px;
              border: 1px solid #DDE3EE; background: #F4F6FA; border-top: 4px solid; }
.score-val  { font-family: Georgia, serif; font-size: 26px; font-weight: 700;
              color: #1B3A6B; line-height: 1; margin-bottom: 4px; }
.score-lbl  { font-size: 11.5px; color: #6B7A99; line-height: 1.4; }
.legend-row { display: flex; gap: 10px; flex-wrap: wrap; padding: 14px 20px;
              background: white; border: 1px solid #DDE3EE; border-radius: 6px;
              margin: 16px 0; align-items: center; }
.leg-lbl { font-size: 10px; text-transform: uppercase; letter-spacing: 2px;
           color: #6B7A99; font-weight: 700; margin-right: 4px; }
</style>
')
```


# Executive Summary {#exec-summary}

```{r exec-hero, results='asis'}
cat('
<div style="background:#1B3A6B;color:white;padding:36px 40px;border-radius:10px;margin-bottom:28px;position:relative;overflow:hidden">
<p style="font-size:10px;text-transform:uppercase;letter-spacing:3px;color:#E8A838;font-weight:700;margin:0 0 14px">Confidential &middot; Senior Management Briefing &middot; April 2026</p>
<p style="font-size:15px;color:rgba(255,255,255,0.65);max-width:700px;margin:0;line-height:1.65">
The 2022 NCUA Risk-Based Capital rule imposed significantly higher capital requirements on the 696 credit unions with average assets above $500 million.
Using a rigorous difference-in-differences design that controls for macroeconomic conditions, pre-existing trends, and thousands of smaller credit unions
unaffected by the rule, this analysis documents the rule's effects across five outcome domains: capital adequacy, loan pricing, credit availability,
portfolio composition, and profitability.
</p>
</div>

<div class="score-row">
<div class="score-card" style="border-top-color:#C94040">
<div class="score-val">+73bp</div>
<div class="score-lbl">Mortgage spread increase over Treasury benchmarks</div>
</div>
<div class="score-card" style="border-top-color:#C94040">
<div class="score-val">$325.9B</div>
<div class="score-lbl">Foregone 4-year member welfare under repeal baseline</div>
</div>
<div class="score-card" style="border-top-color:#1B3A6B">
<div class="score-val">&#x2212;26bp</div>
<div class="score-lbl">Return on assets decline (31% of pre-rule mean)</div>
</div>
<div class="score-card" style="border-top-color:#E8A838">
<div class="score-val">&#x2212;0.3pp</div>
<div class="score-lbl">Rule's actual protection at 2008-crisis stress level</div>
</div>
</div>

<div class="legend-row">
<span class="leg-lbl">Inference key:</span>
<span style="background:#FFF0F0;color:#C94040;border:1px solid #C9404033;padding:3px 10px;border-radius:3px;font-size:10.5px;font-weight:700;font-family:monospace">&#x25BC; Concern</span>
<span style="background:#FFFBF0;color:#B8821A;border:1px solid #E8A83833;padding:3px 10px;border-radius:3px;font-size:10.5px;font-weight:700;font-family:monospace">&#x25C6; Mixed</span>
<span style="background:#F0FFF5;color:#2E7D4F;border:1px solid #2E7D4F33;padding:3px 10px;border-radius:3px;font-size:10.5px;font-weight:700;font-family:monospace">&#x25B2; Positive</span>
<span style="background:#F0F5FF;color:#2A5A9B;border:1px solid #4A7CB533;padding:3px 10px;border-radius:3px;font-size:10.5px;font-weight:700;font-family:monospace">&#x25CF; Baseline</span>
<span style="background:#FFF8F0;color:#8B4513;border:1px solid #8B451333;padding:3px 10px;border-radius:3px;font-size:10.5px;font-weight:700;font-family:monospace">&#x2717; No effect</span>
<span style="font-size:11px;color:#6B7A99;font-style:italic;margin-left:4px">— appears in every Key Numbers and Model Results table</span>
</div>
')
```

**Central finding:** The rule imposed large, persistent costs on member-borrowers through higher loan rates, reduced credit availability, and deteriorating profitability — while providing only marginal protective benefit over and above what existed before the rule. An empirically calibrated threshold of **8.5%**, rather than 10%, would deliver equivalent tail-risk protection at approximately one-third the compliance burden.

---

# Part I: Research Questions & Answers {#rqs}


---

## Research Question 1 {#rq1}

<div style="background:#1B3A6B;color:white;padding:20px 24px;border-radius:8px;margin-bottom:16px">
<p style="font-size:10px;text-transform:uppercase;letter-spacing:2.5px;opacity:0.7;margin:0 0 6px;font-family:monospace">Research Question 1</p>
<p style="font-family:Georgia,serif;font-size:18px;font-weight:700;margin:0;line-height:1.35">What was the average capital buffer for complex credit unions before the RBC rule?</p>
</div>

<div style="display:flex;gap:14px;margin-bottom:16px;flex-wrap:wrap">
<div style="flex:1;min-width:260px;border-left:5px solid #1B3A6B;background:#F4F6FA;padding:14px 18px;border-radius:0 6px 6px 0">
<p style="font-size:9.5px;text-transform:uppercase;letter-spacing:1.5px;color:#6B7A99;font-weight:700;margin:0 0 4px">Answer</p>
<p style="font-family:Georgia,serif;font-size:15px;font-weight:700;color:#1A2035;margin:0">72 basis points — critically thin</p>
</div>
<div style="background:#F0F5FF;border:1px solid #4A7CB544;padding:14px 18px;border-radius:6px;min-width:200px">
<p style="font-size:11px;font-weight:700;color:#2A5A9B;font-family:monospace;margin:0 0 4px">●&nbsp;Baseline</p>
<p style="font-size:11px;color:#2A5A9B;margin:0;line-height:1.4">Baseline context — explains why the rule had such bite</p>
</div>
</div>

Before the rule took effect, the average complex credit union held a net worth ratio of 10.72% — just 72 basis points above the new 10% well-capitalized threshold. Nearly half (47.6%) of complex credit unions were already below 10%, meaning they entered the rule period already non-compliant and facing immediate pressure to build capital. This thin buffer is what gave the rule its bite — and explains both why some institutions struggled to comply and why compliance was so costly.

*→ See [Finding 1: Capital Built — But Many Institutions Still Can't Cross the Bar](#finding1)*


---

## Research Question 2 {#rq2}

<div style="background:#1B3A6B;color:white;padding:20px 24px;border-radius:8px;margin-bottom:16px">
<p style="font-size:10px;text-transform:uppercase;letter-spacing:2.5px;opacity:0.7;margin:0 0 6px;font-family:monospace">Research Question 2</p>
<p style="font-family:Georgia,serif;font-size:18px;font-weight:700;margin:0;line-height:1.35">Are complex credit unions now holding larger capital buffers?</p>
</div>

<div style="display:flex;gap:14px;margin-bottom:16px;flex-wrap:wrap">
<div style="flex:1;min-width:260px;border-left:5px solid #1B3A6B;background:#F4F6FA;padding:14px 18px;border-radius:0 6px 6px 0">
<p style="font-size:9.5px;text-transform:uppercase;letter-spacing:1.5px;color:#6B7A99;font-weight:700;margin:0 0 4px">Answer</p>
<p style="font-family:Georgia,serif;font-size:15px;font-weight:700;color:#1A2035;margin:0">Yes on average — but the average hides a critical distributional problem</p>
</div>
<div style="background:#FFFBF0;border:1px solid #E8A83844;padding:14px 18px;border-radius:6px;min-width:200px">
<p style="font-size:11px;font-weight:700;color:#B8821A;font-family:monospace;margin:0 0 4px">◆&nbsp;Mixed</p>
<p style="font-size:11px;color:#B8821A;margin:0;line-height:1.4">Average improved, pass rate declined — the rule worked for strong CUs, hurt thin-buffer CUs</p>
</div>
</div>

The average net worth ratio rose by 46 basis points relative to trend. However, the probability of being classified as 'well-capitalized' actually fell by 3.7 percentage points. Institutions that were already well above 10% built larger buffers, while institutions near the threshold — the ones the rule was most designed to protect — actually saw capital ratios decline. Among near-threshold credit unions ($400M–$600M in assets), capital ratios fell by 44 basis points.

*→ See [Finding 1: Capital Built — But Many Institutions Still Can't Cross the Bar](#finding1)*


---

## Research Question 3 {#rq3}

<div style="background:#C94040;color:white;padding:20px 24px;border-radius:8px;margin-bottom:16px">
<p style="font-size:10px;text-transform:uppercase;letter-spacing:2.5px;opacity:0.7;margin:0 0 6px;font-family:monospace">Research Question 3</p>
<p style="font-family:Georgia,serif;font-size:18px;font-weight:700;margin:0;line-height:1.35">Did loan rate spreads change significantly after the RBC rule?</p>
</div>

<div style="display:flex;gap:14px;margin-bottom:16px;flex-wrap:wrap">
<div style="flex:1;min-width:260px;border-left:5px solid #C94040;background:#F4F6FA;padding:14px 18px;border-radius:0 6px 6px 0">
<p style="font-size:9.5px;text-transform:uppercase;letter-spacing:1.5px;color:#6B7A99;font-weight:700;margin:0 0 4px">Answer</p>
<p style="font-family:Georgia,serif;font-size:15px;font-weight:700;color:#1A2035;margin:0">Yes — across every loan type, by large and economically meaningful amounts</p>
</div>
<div style="background:#FFF0F0;border:1px solid #C9404044;padding:14px 18px;border-radius:6px;min-width:200px">
<p style="font-size:11px;font-weight:700;color:#C94040;font-family:monospace;margin:0 0 4px">▼&nbsp;Concern</p>
<p style="font-size:11px;color:#C94040;margin:0;line-height:1.4">Complete cost pass-through to member-borrowers confirmed</p>
</div>
</div>

Spreads over Treasury benchmarks rose by 73 basis points for mortgages, 56 basis points for new auto loans, 76 basis points for used auto loans, and 66 basis points for commercial non-real-estate loans. All effects are statistically significant at the 0.1% level. The pass-through appears complete — credit unions raised rates to cover their higher capital costs rather than absorbing them, placing the entire regulatory burden on member-borrowers.

*→ See [Finding 2: Loan Rates Rose Sharply Across Every Product Type](#finding2)*


---

## Research Question 4 {#rq4}

<div style="background:#C94040;color:white;padding:20px 24px;border-radius:8px;margin-bottom:16px">
<p style="font-size:10px;text-transform:uppercase;letter-spacing:2.5px;opacity:0.7;margin:0 0 6px;font-family:monospace">Research Question 4</p>
<p style="font-family:Georgia,serif;font-size:18px;font-weight:700;margin:0;line-height:1.35">Did asset and loan growth change significantly?</p>
</div>

<div style="display:flex;gap:14px;margin-bottom:16px;flex-wrap:wrap">
<div style="flex:1;min-width:260px;border-left:5px solid #C94040;background:#F4F6FA;padding:14px 18px;border-radius:0 6px 6px 0">
<p style="font-size:9.5px;text-transform:uppercase;letter-spacing:1.5px;color:#6B7A99;font-weight:700;margin:0 0 4px">Answer</p>
<p style="font-family:Georgia,serif;font-size:15px;font-weight:700;color:#1A2035;margin:0">Yes — both fell sharply and have shown no recovery</p>
</div>
<div style="background:#FFF0F0;border:1px solid #C9404044;padding:14px 18px;border-radius:6px;min-width:200px">
<p style="font-size:11px;font-weight:700;color:#C94040;font-family:monospace;margin:0 0 4px">▼&nbsp;Concern</p>
<p style="font-size:11px;color:#C94040;margin:0;line-height:1.4">Permanent supply-side lending contraction — members losing access to credit</p>
</div>
</div>

Loan growth fell by approximately 0.44 percentage points per quarter — equivalent to about 1.75 percentage points per year. Asset growth fell by 0.47 percentage points per quarter. These effects appeared at the moment the rule took effect and have not reversed through the full 10+ quarter observation window. Near-threshold credit unions experienced an even larger contraction of 0.6 percentage points per quarter.

*→ See [Finding 3: Lending Growth Fell and Never Recovered](#finding3)*


---

## Research Question 5 {#rq5}

<div style="background:#C94040;color:white;padding:20px 24px;border-radius:8px;margin-bottom:16px">
<p style="font-size:10px;text-transform:uppercase;letter-spacing:2.5px;opacity:0.7;margin:0 0 6px;font-family:monospace">Research Question 5</p>
<p style="font-family:Georgia,serif;font-size:18px;font-weight:700;margin:0;line-height:1.35">Does removing the RBC rule create unacceptable tail risk?</p>
</div>

<div style="display:flex;gap:14px;margin-bottom:16px;flex-wrap:wrap">
<div style="flex:1;min-width:260px;border-left:5px solid #C94040;background:#F4F6FA;padding:14px 18px;border-radius:0 6px 6px 0">
<p style="font-size:9.5px;text-transform:uppercase;letter-spacing:1.5px;color:#6B7A99;font-weight:700;margin:0 0 4px">Answer</p>
<p style="font-family:Georgia,serif;font-size:15px;font-weight:700;color:#1A2035;margin:0">Some risk — but the rule's design is dramatically disproportionate to its benefit</p>
</div>
<div style="background:#FFFBF0;border:1px solid #E8A83844;padding:14px 18px;border-radius:6px;min-width:200px">
<p style="font-size:11px;font-weight:700;color:#B8821A;font-family:monospace;margin:0 0 4px">◆&nbsp;Mixed</p>
<p style="font-size:11px;color:#B8821A;margin:0;line-height:1.4">Repeal adds 3.4pp failure risk vs. $325.9B member welfare cost — recalibration is the answer</p>
</div>
</div>

Empirical stress tests using actual 2008 crisis data for 634 complex credit unions show that repeal would increase institutional failure rates by 3.4 percentage points — approximately 23 additional institutions falling below 7% adequacy. This is real and must be taken seriously. However, the current 10% threshold constrains 47% of complex credit unions unnecessarily. An empirically calibrated threshold of 8.5% would provide equivalent protection while constraining only 15% of institutions — one-third the compliance burden at the same protective benefit.

*→ See [Finding 10: Repeal Carries Real Risk — But the Rule's Protection Is Poorly Calibrated](#finding10)*


---

# Part II: Findings 1–10 — The Full Evidence {#findings}

Each finding section contains:
(a) **Context** — why this question matters, in plain English  
(b) **Key Numbers** — the most important statistics, with an inference column  
(c) **Supporting Chart** — directly imported from `output/paper/figures/`  
(d) **Statistical Model Results** — the formal estimates with signal and meaning columns  
(e) **Key Takeaway** — the management-level conclusion

---


---

## Finding 1: Capital Built — But Many Institutions Still Can't Cross the Bar {#finding1}

<div style="border-left:6px solid #1B3A6B;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#1B3A6B;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 1</span>&ensp;<span style="color:#1B3A6B;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Capital</span>
</div>

### Context

The RBC rule required complex credit unions (those with more than $500 million in assets) to maintain a net worth ratio of at least 10% — a higher bar than the prior 7% threshold. The question is whether this requirement actually worked as intended.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>Average capital increase (trend-adj.)</td><td><strong>+0.46 percentage points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFFBF0;color:#B8821A;border:1px solid #E8A83844;font-size:10px;font-weight:700;font-family:monospace">◆&nbsp;Mixed</span><br><small style="color:#6B7A99">Rose, but masks a distribution problem</small></td></tr>
<tr><td>Well-capitalized probability change</td><td><strong>−3.7 percentage points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Fewer institutions actually passed the bar</small></td></tr>
<tr><td>Pre-rule capital buffer above 10%</td><td><strong>72 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0F5FF;color:#2A5A9B;border:1px solid #4A7CB544;font-size:10px;font-weight:700;font-family:monospace">●&nbsp;Baseline</span><br><small style="color:#6B7A99">Critically thin starting position</small></td></tr>
<tr><td>Near-threshold CU capital change</td><td><strong>−0.44 pp (declined)</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Thin-buffer CUs got worse, not better</small></td></tr>
<tr><td>Institutions studied</td><td><strong>696 complex CUs</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0F5FF;color:#2A5A9B;border:1px solid #4A7CB544;font-size:10px;font-weight:700;font-family:monospace">●&nbsp;Baseline</span><br><small style="color:#6B7A99">Full treatment group</small></td></tr>
</tbody></table>

### Supporting Chart

```{r fig1, echo=FALSE, fig.cap="Event study of net worth ratio. Average rises post-2022, but distribution analysis reveals institutions nearest the threshold were harmed.", out.width="100%"}
fig_path <- file.path(FIG_DIR, "Figure3_EventStudy_Capital.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  message("Figure not found: ", fig_path,
          "\nRun 4_Paper_Tables.R to generate all figures.")
}
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Net worth ratio</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.463pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=156,121</td><td style="padding:8px 11px;background:#F4F6FA;color:#4A5568;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">↑ Mixed</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Average rose, but distribution shifted adversely</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Well-capitalized (LPM)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.037pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p=0.004</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=156,121</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Fewer CUs achieved well-capitalized status</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Near-threshold NW ratio</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.444pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p=0.003</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=19,140</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Opposite sign — rule hurt the most exposed CUs</td></tr>
</tbody></table></div>

### :dart: Key Takeaway

<div style="border:1px solid #1B3A6B;background:#1B3A6B0D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#1B3A6B;font-weight:700;margin-bottom:8px">Key Takeaway for Senior Leadership</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">Think of a class of students all trying to pass a harder exam. The class average score rises — because strong students studied harder — but the pass rate falls, because weaker students still can't reach the new bar. The rule raised the average capital ratio by 46 basis points, but the fraction actually achieving 'well-capitalized' status fell by 3.7 percentage points. Most strikingly, the smallest complex credit unions — those closest to the threshold — actually saw capital ratios decline as compliance costs ate into their retained earnings faster than they could rebuild them.</p>
</div>


---

## Finding 2: Loan Rates Rose Sharply Across Every Product Type {#finding2}

<div style="border-left:6px solid #C94040;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#C94040;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 2</span>&ensp;<span style="color:#C94040;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Member Cost</span>
</div>

### Context

When a credit union faces higher capital requirements, one way it adjusts is by charging higher interest rates on loans — passing the regulatory cost directly to members. We measure this as the spread over comparable Treasury rates, which removes any market-wide rate changes.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>Mortgage spread increase</td><td><strong>+73 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Members pay ~$1,875 more per year on avg mortgage</small></td></tr>
<tr><td>New auto loan spread</td><td><strong>+56 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Members pay ~$176 more per year on avg auto loan</small></td></tr>
<tr><td>Used auto loan spread</td><td><strong>+76 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Larger than new auto — reflects portfolio pressure</small></td></tr>
<tr><td>Commercial non-RE spread</td><td><strong>+66 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Business borrowers also affected</small></td></tr>
<tr><td>Effect for largest CUs (>$10B)</td><td><strong>+97bp auto spread</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Bigger institutions charged members more</small></td></tr>
</tbody></table>

### Supporting Chart

```{r fig2, echo=FALSE, fig.cap="Loan spreads over Treasuries. Flat before 2022Q1, sharp and sustained rise after — consistent with a causal effect of the rule.", out.width="100%"}
fig_path <- file.path(FIG_DIR, "Figure4_EventStudy_Spreads.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  message("Figure not found: ", fig_path,
          "\nRun 4_Paper_Tables.R to generate all figures.")
}
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Mortgage spread</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.750pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=24,927</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Large, highly significant, fully passed through</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">New auto spread</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.587pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=33,335</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Pervasive across all CU sizes</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Used auto spread</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.779pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=33,678</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Largest spread effect in absolute terms</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Comm non-RE spread</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.618pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=9,057</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Business lending affected equally</td></tr>
</tbody></table></div>

### :dart: Key Takeaway

<div style="border:1px solid #C94040;background:#C940400D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#C94040;font-weight:700;margin-bottom:8px">Key Takeaway for Senior Leadership</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">The rate increases are large, pervasive, and affected every loan type we could measure. A 73 basis-point increase in mortgage spreads means a member borrowing $220,000 pays approximately $1,875 more per year — for the life of the loan. The increases appear immediately after the rule takes effect and do not reverse. Larger institutions, which have more market power, charged even more: credit unions above $10 billion raised auto spreads by 97 basis points, well above the average.</p>
</div>


---

## Finding 3: Lending Growth Fell and Never Recovered {#finding3}

<div style="border-left:6px solid #C94040;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#C94040;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 3</span>&ensp;<span style="color:#C94040;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Credit Access</span>
</div>

### Context

Higher capital requirements can reduce the volume of lending — if a credit union must hold more capital per dollar of loans, it may simply make fewer loans. This affects how many members can access credit, not just the price they pay.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>Quarterly loan growth decline</td><td><strong>−0.44 pp per quarter</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Roughly 1.75pp per year in annualized terms</small></td></tr>
<tr><td>Quarterly asset growth decline</td><td><strong>−0.47 pp per quarter</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Broad-based balance-sheet contraction</small></td></tr>
<tr><td>Near-threshold loan growth</td><td><strong>−0.60 pp per quarter</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">50% larger for most constrained CUs</small></td></tr>
<tr><td>Comparison to 2008 crisis</td><td><strong>~60% as large as crisis</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFFBF0;color:#B8821A;border:1px solid #E8A83844;font-size:10px;font-weight:700;font-family:monospace">◆&nbsp;Mixed</span><br><small style="color:#6B7A99">Regulatory contraction rivals financial crisis</small></td></tr>
<tr><td>Duration of effect</td><td><strong>Persistent — no recovery</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">No reversal through 10+ quarters post-rule</small></td></tr>
</tbody></table>

### Supporting Chart

```{r fig3, echo=FALSE, fig.cap="Loan growth event study. Sharp, persistent step-down at RBC implementation with no subsequent recovery — a structural constraint, not a temporary adjustment.", out.width="100%"}
fig_path <- file.path(FIG_DIR, "Figure3_EventStudy_Capital.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  message("Figure not found: ", fig_path,
          "\nRun 4_Paper_Tables.R to generate all figures.")
}
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Loan growth</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.004***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=150,937</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Significant, persistent, no V-shape recovery</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Asset growth</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.005***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=150,937</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Broad contraction, not just loan-specific</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Near-threshold loans</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.006***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p=0.009</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=18,520</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Largest effect for most exposed institutions</td></tr>
</tbody></table></div>

### :dart: Key Takeaway

<div style="border:1px solid #C94040;background:#C940400D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#C94040;font-weight:700;margin-bottom:8px">Key Takeaway for Senior Leadership</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">The lending contraction is not a temporary adjustment — it is a permanent step-down in the growth trajectory of these institutions. Every quarter, complex credit unions grew their loan books about 44 basis points more slowly than comparable non-complex institutions. This effect is driven by supply — the credit unions constrained their lending — not by reduced member demand. Members who needed loans from these institutions were less likely to get them.</p>
</div>


---

## Finding 4: Portfolio Shifted Away from Auto Loans Toward Mortgages — Risk-Weight Arbitrage {#finding4}

<div style="border-left:6px solid #E8A838;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#E8A838;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 4</span>&ensp;<span style="color:#E8A838;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Portfolio Risk</span>
</div>

### Context

Under the RBC rule, different loan types carry different 'risk weights' — essentially different capital costs. Auto loans carry a 100% risk weight (most expensive), while qualifying mortgages carry only 50%. A rational institution under capital pressure will shift toward lower-weight assets, regardless of what's best for members.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>Auto loan share decline</td><td><strong>−4.0 percentage points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Doubled the auto exit seen in 2008 crisis</small></td></tr>
<tr><td>Real estate loan share rise</td><td><strong>+2.4 percentage points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFFBF0;color:#B8821A;border:1px solid #E8A83844;font-size:10px;font-weight:700;font-family:monospace">◆&nbsp;Mixed</span><br><small style="color:#6B7A99">Flight to lower-risk-weight assets</small></td></tr>
<tr><td>MBL share decline</td><td><strong>−0.25 percentage points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Business lending also reduced</small></td></tr>
<tr><td>Auto exit vs. 2008 crisis</td><td><strong>2.2pp LARGER under RBC</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Regulatory-specific — not generic stress</small></td></tr>
<tr><td>Portfolio reversal by year 4</td><td><strong><30% reversal</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Structural — not reverting without rule change</small></td></tr>
</tbody></table>

### Supporting Chart

```{r fig4, echo=FALSE, fig.cap="Portfolio composition around RBC. Auto share falls sharply and permanently; RE share rises — a textbook risk-weight arbitrage pattern not seen in the 2008 crisis.", out.width="100%"}
fig_path <- file.path(FIG_DIR, "Figure5_Portfolio_Trends.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  message("Figure not found: ", fig_path,
          "\nRun 4_Paper_Tables.R to generate all figures.")
}
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Auto share</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−4.004pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=155,946</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Largest effect in study — regulatory arbitrage</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">RE share</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+2.385pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=155,946</td><td style="padding:8px 11px;background:#FFFBF0;color:#B8821A;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">◆ Mixed</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Lower risk-weight, but more recession-sensitive</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">MBL share</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.247pp**</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p=0.026</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=155,946</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Business lending reduced — serves fewer small biz</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Auto: RBC vs crisis Wald</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">2.406pp larger***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">Wald test</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Confirms regulatory-specific, not generic, effect</td></tr>
</tbody></table></div>

### :dart: Key Takeaway

<div style="border:1px solid #E8A838;background:#E8A8380D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#E8A838;font-weight:700;margin-bottom:8px">Key Takeaway for Senior Leadership</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">This is the clearest evidence that the rule had regulatory-specific effects. During the 2008 crisis, complex credit unions shifted away from auto loans by 1.8 percentage points. Under the RBC rule — in a stable economic environment — they shifted by 4.0 percentage points. Institutions responded not to economic conditions but to regulatory incentives. Auto loans are good products for members — accessible, short-duration, spread across many borrowers. Mortgages concentrate risk in real estate. The rule may have made complex credit union portfolios riskier, not safer.</p>
</div>


---

## Finding 5: Loan Quality Deteriorated Despite Higher Capital Requirements {#finding5}

<div style="border-left:6px solid #E8A838;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#E8A838;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 5</span>&ensp;<span style="color:#E8A838;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Credit Quality</span>
</div>

### Context

One might expect that higher capital requirements would make credit unions more conservative lenders — selecting only the best borrowers and thus improving loan quality. The data show the opposite.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>Delinquency rate increase</td><td><strong>+8 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Persistent across 11+ post-treatment quarters</small></td></tr>
<tr><td>Charge-off ratio increase</td><td><strong>+7 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Actual loan losses, not just overdue payments</small></td></tr>
<tr><td>Loan loss provisions</td><td><strong>+1.8 basis points/assets</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Management expecting further deterioration</small></td></tr>
<tr><td>Crisis comparison (DQ)</td><td><strong>Crisis was 4× larger</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0F5FF;color:#2A5A9B;border:1px solid #4A7CB544;font-size:10px;font-weight:700;font-family:monospace">●&nbsp;Baseline</span><br><small style="color:#6B7A99">RBC effect moderate, but concerning given environment</small></td></tr>
<tr><td>Duration</td><td><strong>11+ quarters significant</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Not a temporary blip — sustained pattern</small></td></tr>
</tbody></table>

### Supporting Chart

```{r fig5, echo=FALSE, fig.cap="Delinquency rate event study. Persistent increase beginning with RBC implementation, sustained throughout the observation window.", out.width="100%"}
fig_path <- file.path(FIG_DIR, "Figure3_EventStudy_Capital.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  message("Figure not found: ", fig_path,
          "\nRun 4_Paper_Tables.R to generate all figures.")
}
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Delinquency rate</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.080pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=155,946</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Persistent rise — sustained through Q+11</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Charge-off ratio</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.068pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=156,000</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Actual losses confirm DQ trend is real</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">PLL/assets</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.018pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=156,121</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Management building reserves — forward-looking warning</td></tr>
</tbody></table></div>

### :dart: Key Takeaway

<div style="border:1px solid #E8A838;background:#E8A8380D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#E8A838;font-weight:700;margin-bottom:8px">Key Takeaway for Senior Leadership</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">The portfolio shift from auto loans (stable delinquency profile) toward commercial real estate (more cycle-sensitive) explains part of this deterioration. Additionally, when credit unions are under capital pressure, they may be less able to work with borrowers who fall into difficulty — leading to faster progression to delinquency and charge-off. The rule designed to make credit unions more resilient may have made their loan books incrementally more fragile.</p>
</div>


---

## Finding 6: Profitability Declined Persistently — Largest Institutions Hit Hardest {#finding6}

<div style="border-left:6px solid #1B3A6B;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#1B3A6B;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 6</span>&ensp;<span style="color:#1B3A6B;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Profitability</span>
</div>

### Context

A credit union's return on assets (ROA) measures how much profit it generates per dollar of assets. This matters because retained earnings — the profit not distributed — is the primary way credit unions build capital organically. A rule that reduces ROA undermines the very capital-building it aims to promote.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>ROA decline — full sample</td><td><strong>−26 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">31% of the pre-rule mean ROA of 0.83%</small></td></tr>
<tr><td>ROA — $500M–$1B CUs</td><td><strong>−19 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Smaller complex CUs still materially affected</small></td></tr>
<tr><td>ROA — $1B–$10B CUs</td><td><strong>−29 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Mid-size CUs bear disproportionate burden</small></td></tr>
<tr><td>ROA — >$10B CUs</td><td><strong>−47 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Largest and most systemically important: worst hit</small></td></tr>
<tr><td>NIM compression</td><td><strong>−12 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Rate income squeezed while funding costs rose</small></td></tr>
<tr><td>Trend direction</td><td><strong>Worsening through Q+10</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">No recovery — a permanent earnings reduction</small></td></tr>
</tbody></table>

### Supporting Chart

```{r fig6, echo=FALSE, fig.cap="ROA event study. Monotonically declining path post-RBC — unlike the V-shaped recovery seen in the 2008 crisis. No sign of reversal through Q+10.", out.width="100%"}
fig_path <- file.path(FIG_DIR, "Figure3_EventStudy_Capital.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  message("Figure not found: ", fig_path,
          "\nRun 4_Paper_Tables.R to generate all figures.")
}
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">ROA (full sample)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.259pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=156,106</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Large, monotonically worsening through Q+10</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">NIM</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.117pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=156,121</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Rate income squeezed — margin compression</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Cost of funds</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.246pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=156,121</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Higher deposit costs compound earnings pressure</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">ROA — >$10B tier</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.467pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">Table A1</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Systemically important institutions most affected</td></tr>
</tbody></table></div>

### :dart: Key Takeaway

<div style="border:1px solid #1B3A6B;background:#1B3A6B0D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#1B3A6B;font-weight:700;margin-bottom:8px">Key Takeaway for Senior Leadership</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">The profitability decline is the most alarming systemic finding. A credit union that cannot earn adequate returns cannot rebuild capital through retained earnings, cannot absorb loan losses without approaching regulatory thresholds, and eventually cannot offer competitive rates. The worsening trend through Q+10 — with no reversal — indicates this is a permanent regulatory earnings tax, not a temporary adjustment cost. Larger institutions, which have the greatest systemic importance, have been hit hardest.</p>
</div>


---

## Finding 7: The RBC Rule Hit Harder Than the 2008 Financial Crisis on Key Metrics {#finding7}

<div style="border-left:6px solid #1B3A6B;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#1B3A6B;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 7</span>&ensp;<span style="color:#1B3A6B;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Regulatory ID</span>
</div>

### Context

To distinguish regulatory effects from generic financial stress, we compare what happened to complex credit unions under the RBC rule versus during the 2008 financial crisis. Differences reveal effects that are specific to the regulatory design — not just responses to economic conditions.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>Loan growth contraction: RBC vs. crisis</td><td><strong>RBC 47% larger</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Rule suppressed lending more than the financial crisis</small></td></tr>
<tr><td>Auto exit: RBC vs. crisis</td><td><strong>+2.4pp larger under RBC</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Risk-weight arbitrage — regulatory-specific</small></td></tr>
<tr><td>ROA sign under crisis</td><td><strong>+7bp (flat/positive)</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0F5FF;color:#2A5A9B;border:1px solid #4A7CB544;font-size:10px;font-weight:700;font-family:monospace">●&nbsp;Baseline</span><br><small style="color:#6B7A99">Crisis did not compress earnings at complex CUs</small></td></tr>
<tr><td>ROA sign under RBC</td><td><strong>−26bp (persistent decline)</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Opposite sign — rule is a permanent earnings tax</small></td></tr>
<tr><td>NIM under crisis</td><td><strong>+23bp (widened)</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0FFF5;color:#2E7D4F;border:1px solid #2E7D4F44;font-size:10px;font-weight:700;font-family:monospace">▲&nbsp;Positive</span><br><small style="color:#6B7A99">Crisis widened spreads — institutions adapted</small></td></tr>
<tr><td>NIM under RBC</td><td><strong>−12bp (compressed)</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Rule compressed income — opposite to crisis</small></td></tr>
</tbody></table>

### Supporting Chart

```{r fig7, echo=FALSE, fig.cap="Parallel DiD comparing 2008 crisis vs. 2022 RBC rule. Capital responses are similar; profitability and portfolio show opposite-sign patterns — confirming a regulatory channel.", out.width="100%"}
fig_path <- file.path(FIG_DIR, "Figure6_Crisis_vs_RBC_Capital.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  message("Figure not found: ", fig_path,
          "\nRun 4_Paper_Tables.R to generate all figures.")
}
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Loan growth diff (RBC larger)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.008pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p=0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">Wald test</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">RBC suppressed lending significantly more than crisis</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Auto share diff (RBC larger)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−2.406pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">Wald test</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Regulatory arbitrage confirmed — not generic stress</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">ROA diff (RBC worse)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.487pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">Wald test</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Opposite signs — rule is a regulatory earnings tax</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">NIM diff (RBC worse)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.403pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">Wald test</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Confirms margin compression is regulatory-specific</td></tr>
</tbody></table></div>

### :dart: Key Takeaway

<div style="border:1px solid #1B3A6B;background:#1B3A6B0D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#1B3A6B;font-weight:700;margin-bottom:8px">Key Takeaway for Senior Leadership</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">During the 2008 financial crisis — the worst economic shock in post-war history — complex credit unions saw flat-to-positive ROA, because they could charge higher rates and credit losses were offset by earnings. Under the RBC rule — in a stable economic environment — ROA declined persistently and NIM compressed. This opposite-sign pattern for profitability is the clearest evidence that the RBC rule functions as a regulatory earnings tax rather than a prudential safeguard. The rule is doing something the financial crisis did not.</p>
</div>


---

## Finding 8: Auto Loan Exit Is Permanent Under RBC — Unlike Every Other Crisis Response {#finding8}

<div style="border-left:6px solid #E8A838;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#E8A838;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 8</span>&ensp;<span style="color:#E8A838;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Structural Change</span>
</div>

### Context

When credit unions face stress, they adjust their portfolios. The critical question is whether this adjustment is temporary (they snap back when pressure eases) or permanent (structural). The answer depends on what drove the adjustment — economic conditions reverse; regulatory incentives don't.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>Auto share path under crisis</td><td><strong>V-shaped — fell then recovered</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0FFF5;color:#2E7D4F;border:1px solid #2E7D4F44;font-size:10px;font-weight:700;font-family:monospace">▲&nbsp;Positive</span><br><small style="color:#6B7A99">Classic temporary stress response</small></td></tr>
<tr><td>Auto share path under RBC</td><td><strong>Step-change — fell and stayed</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Structural — regulatory incentive permanent</small></td></tr>
<tr><td>ROA path under crisis</td><td><strong>V-shaped — declined then recovered</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0FFF5;color:#2E7D4F;border:1px solid #2E7D4F44;font-size:10px;font-weight:700;font-family:monospace">▲&nbsp;Positive</span><br><small style="color:#6B7A99">Normal recovery once shock passed</small></td></tr>
<tr><td>ROA path under RBC</td><td><strong>Linear decline through Q+10</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">No recovery — permanent regulatory tax</small></td></tr>
<tr><td>Capital path: both episodes</td><td><strong>V-shaped in both</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0F5FF;color:#2A5A9B;border:1px solid #4A7CB544;font-size:10px;font-weight:700;font-family:monospace">●&nbsp;Baseline</span><br><small style="color:#6B7A99">Generic response — expected in both cases</small></td></tr>
</tbody></table>

### Supporting Chart

```{r fig8, echo=FALSE, fig.cap="Shape comparison: V-shapes under crisis, step-changes under RBC. The divergence in portfolio and profitability paths is the structural signature of regulatory-specific effects.", out.width="100%"}
fig_path <- file.path(FIG_DIR, "Figure6b_Crisis_vs_RBC_Portfolio.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  message("Figure not found: ", fig_path,
          "\nRun 4_Paper_Tables.R to generate all figures.")
}
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Auto share persistence (RBC)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">>90% at Q+10</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">event study</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">3B analysis</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Structural — 90%+ of exit still present after 2.5yr</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Auto share persistence (crisis)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap"><30% at Q+10</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">event study</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">3B analysis</td><td style="padding:8px 11px;background:#F0FFF5;color:#2E7D4F;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▲ Positive</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Temporary — virtually fully reversed by Q+10</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">ROA recovery (crisis)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">Baseline by Q+8</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">event study</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">3B analysis</td><td style="padding:8px 11px;background:#F0FFF5;color:#2E7D4F;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▲ Positive</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Normal V-shape — crisis recovered within 2 years</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">ROA trajectory (RBC)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">No reversal Q+10</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">event study</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">3B analysis</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">No recovery — permanent structural change</td></tr>
</tbody></table></div>

### :dart: Key Takeaway

<div style="border:1px solid #E8A838;background:#E8A8380D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#E8A838;font-weight:700;margin-bottom:8px">Key Takeaway for Senior Leadership</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">V-shaped responses are what we expect from temporary shocks. Step-change responses indicate structural realignment driven by permanent incentives. The auto loan exit under the RBC rule is structural because the risk-weight incentive never changes — unless the rule is redesigned. Complex credit unions have permanently repositioned their portfolios away from auto loans, regardless of whether this is good for their members or their own financial health.</p>
</div>


---

## Finding 9: The CCULR Relief Valve Didn't Work {#finding9}

<div style="border-left:6px solid #4A7CB5;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#4A7CB5;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 9</span>&ensp;<span style="color:#4A7CB5;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Policy Design</span>
</div>

### Context

Recognizing that the 10% threshold might be too burdensome, NCUA created a simpler alternative — the Complex Credit Union Leverage Ratio (CCULR) — with a lower 9% threshold. This was designed to provide regulatory relief to constrained institutions. The data show it did not.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>Loan growth relief from CCULR</td><td><strong>Not significant (p=0.197)</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Zero lending relief — the primary policy goal failed</small></td></tr>
<tr><td>Capital outcome for CCULR adopters</td><td><strong>Worse than full RBC</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Adopters ended up with less capital</small></td></tr>
<tr><td>ROA for CCULR adopters</td><td><strong>−32bp vs −24bp (full RBC)</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Worse profitability — selection effect</small></td></tr>
<tr><td>Why it failed</td><td><strong>Distressed CUs self-selected</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Relief went to institutions that couldn't use it</small></td></tr>
<tr><td>1pp threshold reduction adequacy</td><td><strong>Insufficient for bound CUs</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">9% vs 10% doesn't help a CU at 8%</small></td></tr>
</tbody></table>

### Supporting Chart

```{r fig9, echo=FALSE, fig.cap="Three-group event study. CCULR adopters show pre-trend violations and worse outcomes — confirming negative selection rather than genuine relief.", out.width="100%"}
fig_path <- file.path(FIG_DIR, "Figure7_CCULR_EventStudy.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  message("Figure not found: ", fig_path,
          "\nRun 4_Paper_Tables.R to generate all figures.")
}
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Loan growth relief (β₂−β₁)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.002 (n.s.)</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p=0.197</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=~156K</td><td style="padding:8px 11px;background:#FFF8F0;color:#8B4513;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">✗ No effect</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">PRIMARY POLICY GOAL: zero significant relief</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Capital relief</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.374pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p=0.004</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=~156K</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Worse</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">CCULR adopters have LESS capital than full RBC CUs</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">ROA relief</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.084pp**</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p=0.011</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=~156K</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Worse</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Profitability also worse — confirms selection</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Auto share relief</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.136 (n.s.)</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p=0.817</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=~156K</td><td style="padding:8px 11px;background:#FFF8F0;color:#8B4513;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">✗ No effect</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Portfolio reallocation equally bad under CCULR</td></tr>
</tbody></table></div>

### :dart: Key Takeaway

<div style="border:1px solid #4A7CB5;background:#4A7CB50D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#4A7CB5;font-weight:700;margin-bottom:8px">Key Takeaway for Senior Leadership</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">The CCULR failed as a relief mechanism for a structural reason: the institutions that adopted it were disproportionately those already in financial difficulty. A credit union with an 8% capital ratio that adopts the 9% CCULR threshold has gained no regulatory breathing room — it is still below the threshold. The 1 percentage point reduction was simply not large enough to provide meaningful relief to genuinely constrained institutions.</p>
</div>


---

## Finding 10: Repeal Carries Real Risk — But the Rule's Protection Is Poorly Calibrated {#finding10}

<div style="border-left:6px solid #C94040;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#C94040;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 10</span>&ensp;<span style="color:#C94040;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Prudential Risk</span>
</div>

### Context

Everything above documents costs of the rule. But the rule was designed to protect the financial system. Before concluding that repeal is beneficial, we must quantify the rule's actual protective value using empirical stress tests from real crisis data.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>Rule's protection at 2008 stress</td><td><strong>−0.3pp (negligible)</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFFBF0;color:#B8821A;border:1px solid #E8A83844;font-size:10px;font-weight:700;font-family:monospace">◆&nbsp;Mixed</span><br><small style="color:#6B7A99">Rule barely reduces failure rates at historical stress</small></td></tr>
<tr><td>Pre-rule failure rate (below 7%)</td><td><strong>19.1%</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0F5FF;color:#2A5A9B;border:1px solid #4A7CB544;font-size:10px;font-weight:700;font-family:monospace">●&nbsp;Baseline</span><br><small style="color:#6B7A99">Baseline — without any rule</small></td></tr>
<tr><td>With-rule failure rate (below 7%)</td><td><strong>19.4%</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0F5FF;color:#2A5A9B;border:1px solid #4A7CB544;font-size:10px;font-weight:700;font-family:monospace">●&nbsp;Baseline</span><br><small style="color:#6B7A99">Essentially unchanged from pre-rule</small></td></tr>
<tr><td>Post-repeal failure rate</td><td><strong>22.8% (+3.4pp)</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Real cost: ~23 additional institutions at risk</small></td></tr>
<tr><td>Optimal threshold (empirical)</td><td><strong>~8.5% (not 10%)</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFFBF0;color:#B8821A;border:1px solid #E8A83844;font-size:10px;font-weight:700;font-family:monospace">◆&nbsp;Mixed</span><br><small style="color:#6B7A99">Lower threshold achieves same protection</small></td></tr>
<tr><td>Cost per institution protected</td><td><strong>~$14B in member welfare</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Disproportionate cost-benefit ratio</small></td></tr>
</tbody></table>

### Supporting Chart

```{r fig10, echo=FALSE, fig.cap="Stress survival rates by regime and severity. The rule provides essentially no protection at 2008-severity stress (pre vs. with-rule: −0.3pp). Repeal adds 3.4pp of failure risk.", out.width="100%"}
fig_path <- file.path(FIG_DIR, "Figure11_4_Survival_Frontier.png")
if (file.exists(fig_path)) {
  knitr::include_graphics(fig_path)
} else {
  message("Figure not found: ", fig_path,
          "\nRun 4_Paper_Tables.R to generate all figures.")
}
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Pre→With rule (2008 stress)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.3pp</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">n.s.</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">634 CUs</td><td style="padding:8px 11px;background:#FFFBF0;color:#B8821A;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">◆ Mixed</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Rule provides NEGLIGIBLE protection at 2008 severity</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">With rule→Repeal (2008)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+3.4pp</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">—</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">634 CUs</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Repeal adds 3.4pp failure risk — real but modest</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">With rule→Repeal (1.5× tail)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+3.1pp</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">—</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">634 CUs</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Consistent across stress scenarios</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">With rule→Repeal (2.0× tail)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+4.6pp</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">—</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">634 CUs</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Largest at severe tail — but still moderate</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Optimal threshold (Fig 11.8)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">~8.5%</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">—</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">Empirical</td><td style="padding:8px 11px;background:#FFFBF0;color:#B8821A;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">◆ Mixed</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Lower threshold = same protection, 1/3 the burden</td></tr>
</tbody></table></div>

### :dart: Key Takeaway

<div style="border:1px solid #C94040;background:#C940400D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#C94040;font-weight:700;margin-bottom:8px">Key Takeaway for Senior Leadership</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">Repeal carries real tail risk — approximately 23 additional institutions would fall below 7% capitalization in a 2008-severity crisis. This cost must be taken seriously. However, the current 10% threshold constrains 47% of complex credit unions unnecessarily. An empirically calibrated threshold of 8.5% would provide equivalent protection while constraining only 15% of institutions — one-third the compliance burden at the same protective benefit. The policy recommendation is recalibration, not abolition: lower the threshold, redesign the risk weights, and eliminate the provisions that drove the auto loan exit without prudential justification.</p>
</div>


---

# Section 9: Repeal Simulation — What Happens If the Rule Is Removed? {#repeal}

These figures show projected trajectories for key outcomes in the 3 years following a hypothetical 2026Q1 repeal of the RBC rule, under Fast (4Q half-life), Gradual (8Q/16Q), and Slow (plateau) recovery scenarios. All estimates derive directly from the DiD parameters — no new regressions.

```{r fig9-1, fig.cap="Loan spread fan chart under repeal — three recovery scenarios"}
fig_include("Figure9_1_Spread_Fan_Chart.png")
```

```{r fig9-3, fig.cap="Cumulative member savings from spread reduction ($B) — 4-year horizon"}
fig_include("Figure9_3_Cumulative_Member_Savings.png")
```

```{r fig9-4, fig.cap="ROA recovery by asset size tier post-repeal"}
fig_include("Figure9_4_ROA_Recovery_by_Tier.png")
```

```{r fig9-8, fig.cap="Cost of delay: foregone member welfare by repeal timing"}
fig_include("Figure9_8_Cost_of_Delay.png")
```

---

# Section 10: Stress Scenario — Repeal Into a Recession {#stress}

These figures overlay a moderate recession (oil shock + inflation, 60% of 2008 crisis magnitude, 6Q duration) on the repeal recovery path. Recession calibrated from 3A crisis DiD estimates.

```{r fig10-1, fig.cap="Net worth ratio: repeal benefit vs. recession headwind"}
fig_include("Figure10_1_Stress_Capital.png")
```

```{r fig10-4, fig.cap="Cumulative member savings net of recession — all scenarios"}
fig_include("Figure10_4_Stress_Net_Welfare.png")
```

```{r fig10-6, fig.cap="Four-panel dashboard: net outcomes across channels under recession"}
fig_include("Figure10_6_Stress_Scenario_Comparison.png")
```

```{r fig10-8, fig.cap="Cost of delaying repeal into a deepening recession"}
fig_include("Figure10_8_Stress_Repeal_Timing.png")
```

---

# Section 11: Capital Adequacy Stress Test — Empirical Answer to RQ5 {#stress-test}

These figures are based on actual 2008 crisis drawdown data for 634 complex credit unions (mean drawdown 1.86pp, P90 3.36pp) and empirical NW ratio distributions from the NCUA Call Report.

```{r fig11-1, fig.cap="NW ratio distribution: pre-rule, with-rule, and post-repeal"}
fig_include("Figure11_1_NW_Distribution_Regimes.png")
```

```{r fig11-2, fig.cap="Fraction of complex CUs below each capital threshold by regime"}
fig_include("Figure11_2_Threshold_Breach_Rates.png")
```

```{r fig11-3, fig.cap="Empirical stress loss calibration from 2008 crisis data"}
fig_include("Figure11_3_Stress_Drawdown_Calibration.png")
```

```{r fig11-4, fig.cap="Stress survival frontier: % falling below 7% by regime and severity"}
fig_include("Figure11_4_Survival_Frontier.png")
```

```{r fig11-6, fig.cap="RBC rule marginal effect on institutional failure risk"}
fig_include("Figure11_6_Rule_Effect_vs_Tailrisk.png")
```

```{r fig11-7, fig.cap="Near-threshold institution vulnerability map — actual 2008 drawdown data"}
fig_include("Figure11_7_Thin_Buffer_Zoom.png")
```

```{r fig11-8, fig.cap="Threshold calibration: compliance burden vs. residual failure risk. Curves cross at ~8.5%."}
fig_include("Figure11_8_Optimal_Threshold.png")
```

---

# Data & Methods Note {#methods}

**Data:** NCUA Call Report (Form 5300), 2000Q1–2025Q4. Raw file: `call_report.rds`. Clean analysis panel: `data/analysis_panel.rds` (156,121 observations, 5,184 credit unions, 83 variables). Full time-series panel: `data/analysis_panel_full_raw.rds` (2000–2025, 634 complex CUs with 2008 crisis data).

**Treatment group:** 696 complex credit unions with average assets ≥ $500M in 2021Q1–Q4 (pre-treatment, time-invariant classification).

**Control group:** 4,488 non-complex credit unions with average assets < $500M.

**Identification strategy:** Two-way fixed effects DiD with institution and quarter-year fixed effects. Standard errors clustered at the institution level. Trend-adjusted specification for capital and MBL share (pre-treatment divergence). Event study specification for dynamic effect patterns.

**Statistical significance:** *** p < 0.001, ** p < 0.01, * p < 0.05. All inference is two-sided.

**R analysis pipeline:**

| Script | Purpose |
|--------|---------|
| `0_Data_Prep.R` | Panel construction (2018–2025) |
| `0B_Crisis_Panel_Prep.R` | Full panel construction (2000–2025) |
| `1_Descriptive_Stats_FINAL.R` | Tables 1–2 |
| `2_DiD_Estimation_v2.R` | Tables 3–6, event studies |
| `3A_2008_Crisis_Parallel_DiD.R` | Table 7, crisis comparison |
| `3B_2008_Crisis_EventStudy.R` | Shape analysis |
| `3C_CCULR_Adoption.R` | Table 8, CCULR analysis |
| `3D_Repeal_Simulation.R` | Section 9 repeal simulation |
| `3E_Stress_Scenario_Simulation.R` | Section 10 recession overlay |
| `3F_Capital_Adequacy_Stress_Test.R` | Section 11 capital stress test |
| `4_Paper_Tables.R` | All publication figures |

---

*Prepared by the NCUA Office of the Chief Economist. For internal use. April 2026.*
