---
title: "The 2022 NCUA Risk-Based Capital Rule"
subtitle: "Empirical Findings from an Analysis of Complex Credit Unions"
author: "Saurabh C. Datta, Ph.D. | Senior Economist, NCUA Office of the Chief Economist"
date: "April 2026"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 2
    self_contained: true
    css: rbc_pres.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
fig_dir <- "output/paper/figures/"
fig_include <- function(path, caption = "") {
  if (file.exists(path)) {
    knitr::include_graphics(path)
  } else {
    cat(paste0('<div class="fig-ph"><span>Figure: ', basename(path), '</span></div>'))
  }
}
```

<div class="cover-slide">
<div class="cover-org">NCUA &middot; Office of the Chief Economist</div>
<div class="cover-title">The 2022 NCUA Risk-Based Capital Rule</div>
<div class="cover-sub">Empirical Findings from an Analysis of Complex Credit Unions</div>
<div class="cover-rule"></div>
<div class="cover-meta">Saurabh C. Datta, Ph.D. &middot; Senior Economist<br>NCUA Call Report Data, 2000&ndash;2025 &middot; April 2026</div>
</div>

---

## What We Studied {.slide-section}

<div class="two-col">
<div class="col-l">

<div class="lbl">The Data</div>

- **696 complex credit unions** &mdash; average assets above $500M, classified using 2021 data
- **~4,488 non-complex credit unions** as the comparison group
- **150,000+ CU-quarters** of NCUA Call Report data, 2000&ndash;2025
- **12 outcomes** measured: capital, loan rates, lending volume, portfolio mix, profitability, and loan quality

<div class="lbl" style="margin-top:12px;">The Method: Difference-in-Differences (DiD)</div>

<div class="method-row">
<div class="mbox"><strong>Change in complex CUs</strong><br><small>Before vs. after rule</small></div>
<div class="mop">&minus;</div>
<div class="mbox"><strong>Change in non-complex CUs</strong><br><small>Same time window</small></div>
<div class="mop">=</div>
<div class="mres"><strong>The Rule's Effect</strong><br><small>What changed because of the rule, not the economy</small></div>
</div>

<div class="cb"><strong>Why this works:</strong> Non-complex credit unions faced the same economic conditions but were not subject to the RBC rule. Any difference in outcomes between the two groups, after accounting for pre-existing trends, isolates the rule's effect.</div>

</div>
<div class="col-r">

<div class="lbl">How to Read the Event Study Charts</div>

<div class="ib">
- **Horizontal axis:** quarters relative to the rule's effective date (Q0). Left = before; right = after.
- **Vertical axis:** the estimated gap between complex and non-complex CUs, above any pre-existing difference.
- **Before Q0:** dots near zero confirm the two groups moved together &mdash; the parallel trends check.
- **After Q0:** a persistent move up or down is the rule's estimated effect. The shaded band is the 95% confidence interval.
- **If the band crosses zero** at a given quarter, the effect is not statistically distinguishable from no effect.
</div>

<svg viewBox="0 0 300 100" style="width:100%;border:1px solid #ddd;border-radius:4px;background:#fff;margin-top:8px;">
<line x1="20" y1="55" x2="285" y2="55" stroke="#ccc" stroke-width="0.8" stroke-dasharray="3,2"/>
<line x1="155" y1="8" x2="155" y2="90" stroke="#c8972a" stroke-width="1.2" stroke-dasharray="3,2"/>
<polygon points="60,50 105,52 155,55 198,43 240,38 282,36 282,48 240,50 198,55 155,62 105,60 60,62" fill="#1a2744" opacity="0.12"/>
<polyline points="60,51 105,53 155,55 198,44 240,39 282,37" fill="none" stroke="#1a2744" stroke-width="2"/>
<circle cx="60" cy="51" r="3" fill="#1a2744"/><circle cx="105" cy="53" r="3" fill="#1a2744"/>
<circle cx="155" cy="55" r="3" fill="#1a2744"/><circle cx="198" cy="44" r="3" fill="#1a2744"/>
<circle cx="240" cy="39" r="3" fill="#1a2744"/><circle cx="282" cy="37" r="3" fill="#1a2744"/>
<text x="60" y="100" font-size="8" text-anchor="middle" fill="#888" font-family="Arial">-4Q</text>
<text x="105" y="100" font-size="8" text-anchor="middle" fill="#888" font-family="Arial">-2Q</text>
<text x="155" y="100" font-size="8" text-anchor="middle" fill="#888" font-family="Arial">0</text>
<text x="198" y="100" font-size="8" text-anchor="middle" fill="#888" font-family="Arial">+2Q</text>
<text x="240" y="100" font-size="8" text-anchor="middle" fill="#888" font-family="Arial">+4Q</text>
<text x="282" y="100" font-size="8" text-anchor="middle" fill="#888" font-family="Arial">+6Q</text>
<text x="158" y="20" font-size="7.5" fill="#c8972a" font-family="Arial">Rule effective</text>
<text x="62" y="45" font-size="7.5" fill="#555" font-family="Arial">Flat pre-period</text>
<text x="162" y="38" font-size="7.5" fill="#1a2744" font-family="Arial">Persistent effect</text>
</svg>
<p style="font-size:10px;color:#aaa;text-align:center;margin-top:4px;">Schematic only &mdash; actual charts appear in each section below</p>

</div>
</div>

---

## RQ 1 &mdash; Pre-Rule Capital Buffer {.slide-section}

<span class="bdg-rq">Research Question 1</span>

**How much capital buffer did complex credit unions hold before the rule?**

<div class="abar">Answer: The average complex credit union entered the rule with a buffer of just <strong>72 basis points</strong> above the new 10% threshold &mdash; critically thin.</div>

<div class="two-col">
<div class="col-l">

<div class="lbl">What the Data Show</div>

- The 696 complex CUs were nominally compliant before the rule &mdash; but with very little margin to absorb a new, higher threshold
- A **72 basis point buffer** means holding 10.72% net worth when the definition of well-capitalized shifted upward simultaneously
- Capital was **tightly clustered just above the old threshold** &mdash; a large mass barely on the compliant side
- This thin margin directly explains why the rule had such large downstream effects &mdash; institutions had to act immediately

</div>
<div class="col-r">

<div class="cb"><strong>What is a basis point?</strong> One basis point (bp) equals one one-hundredth of a percentage point. A 72bp buffer means a credit union holding 10.72% capital against a 10.00% threshold.</div>

<div class="stat-row">
<div class="stat-card"><div class="sn">72bp</div><div class="sl">Average pre-rule buffer above the 10% threshold</div></div>
<div class="stat-card red"><div class="sn">696</div><div class="sl">Complex CUs subject to the rule (fixed 2021 classification)</div></div>
</div>

<div class="cb" style="margin-top:10px;"><strong>Why it matters:</strong> A rule requiring additional capital is always more costly when institutions start with little room to spare. The thin pre-rule buffer set the stage for the compliance pressures documented in RQ2&ndash;RQ4.</div>

</div>
</div>

---

## RQ 2 &mdash; Did Capital Levels Rise? {.slide-section}

<span class="bdg-rq">Research Question 2</span>

**Did the rule cause complex credit unions to hold more capital?**

<div class="abar">Answer: <strong>Mixed.</strong> Average capital levels rose &mdash; but fewer institutions are now meeting the well-capitalized standard. The bar rose faster than capital could accumulate.</div>

<div class="two-col">
<div class="col-l">

<div class="stat-row" style="margin-bottom:14px;">
<div class="stat-card green"><div class="sn">+46bp</div><div class="sl">Net worth ratio increase (p &lt; 0.01)</div></div>
<div class="stat-card red"><div class="sn">-3.7pp</div><div class="sl">Well-capitalized probability change (p &lt; 0.01)</div></div>
</div>

<div class="lbl">The Paradox Explained</div>

- Capital rose &mdash; institutions did respond to the rule's incentive to build reserves
- But the **definition of well-capitalized changed** simultaneously, raising the bar
- Result: more capital on average, but **fewer institutions clearing the new bar**

<div class="cb" style="margin-top:10px;"><strong>Analogy:</strong> Every student studied harder and raised their score &mdash; but the passing grade was raised even more. Average scores went up, yet more students are technically failing. Both facts are simultaneously true.</div>

</div>
<div class="col-r">

<div class="lbl">Event Study &mdash; Capital</div>

```{r fig-rq2, out.width="100%", results="asis"}
fig_include(file.path(fig_dir, "Figure3_EventStudy_Capital.png"))
```

<p class="fn">Pre-period flat confirms parallel trends. Post-period divergence attributed to the rule.</p>

</div>
</div>

---

## RQ 3 &mdash; Member Cost: Higher Rates and Less Lending {.slide-section}

<span class="bdg-rq">Research Question 3</span>

**What did the rule cost members in higher loan rates and reduced lending?**

<div class="abar">Answer: Loan rates rose across every product; lending contracted persistently. Estimated total member cost: <strong>$325.9 billion over 4 years</strong> &mdash; approximately <strong>$865 per member per year</strong>.</div>

<div class="two-col">
<div class="col-l">

<div class="lbl">Loan Rate Increases vs. Non-Complex CUs</div>

<div class="stat-row" style="margin-bottom:10px;">
<div class="stat-card red"><div class="sn sm">+75bp</div><div class="sl">Mortgage spread</div></div>
<div class="stat-card red"><div class="sn sm">+59bp</div><div class="sl">Auto loan spread</div></div>
<div class="stat-card red"><div class="sn sm">+58bp</div><div class="sl">Commercial spread</div></div>
</div>

- All increases **statistically significant**, persist 10+ quarters with no sign of convergence
- A 75bp mortgage increase on a $300,000 loan adds roughly **$135 per month** in payments

```{r fig-rq3a, out.width="100%", results="asis"}
fig_include(file.path(fig_dir, "Figure4_EventStudy_Spreads.png"))
```

</div>
<div class="col-r">

<div class="lbl">Lending Volume and Total Member Cost</div>

<div class="stat-row" style="margin-bottom:10px;">
<div class="stat-card red"><div class="sn sm">-0.44pp</div><div class="sl">Quarterly loan growth &mdash; persistent, no recovery</div></div>
<div class="stat-card red"><div class="sn sm">$325.9B</div><div class="sl">Total member welfare cost over 4 years</div></div>
</div>

- Loan growth fell **0.44pp per quarter** &mdash; compounding into a substantial reduction in credit availability
- $325.9BN aggregates the rate-increase effect across all member-borrowers of all 696 complex CUs over four years
- Per-member: approximately **$865/year** in higher borrowing costs

```{r fig-rq3b, out.width="100%", results="asis"}
fig_include(file.path(fig_dir, "Figure3b_EventStudy_LoanGrowth.png"))
```

</div>
</div>

---

## RQ 4 &mdash; Portfolio Shift and the Crisis Comparison {.slide-section}

<span class="bdg-rq">Research Question 4</span>

**Did the rule change what loans credit unions make &mdash; and how does this compare to past stress?**

<div class="abar">Answer: Yes. Complex CUs shifted from auto loans toward mortgages &mdash; consistent with <strong>risk-weight arbitrage</strong>. This shift is unique to the RBC rule; it did not occur during the 2008 financial crisis.</div>

<div class="two-col">
<div class="col-l">

<div class="stat-row" style="margin-bottom:10px;">
<div class="stat-card red"><div class="sn">-4.0pp</div><div class="sl">Auto loan share (vs. controls)</div></div>
<div class="stat-card green"><div class="sn">+2.4pp</div><div class="sl">Real estate loan share</div></div>
</div>

- **Risk-weight arbitrage:** mortgages carry lower risk weights than auto loans under the RBC rule. Shifting to mortgages reduces required capital &mdash; a rational regulatory response
- Loan quality also deteriorated: **delinquency +8bp**, driven by a capacity channel &mdash; CUs with the largest profitability declines showed 2&times; the delinquency increase

```{r fig-rq4a, out.width="100%", results="asis"}
fig_include(file.path(fig_dir, "Figure5_Portfolio_Trends.png"))
```

</div>
<div class="col-r">

<div class="lbl">Comparison to the 2008 Financial Crisis</div>

<div class="cb">We ran the same DiD analysis using 2008 as the event, comparing how complex CUs responded to the financial crisis versus the RBC rule.</div>

| Outcome | 2008 Crisis | RBC Rule |
|:--------|:-----------:|:--------:|
| Capital accumulation | Similar | Similar |
| Auto loan exit | Not observed | **Persistent** |
| Loan growth contraction | Temporary | **No recovery** |
| Profitability decline | Temporary | **No recovery** |

Capital behavior resembles a crisis response &mdash; but the lending and profitability effects are **distinctly regulatory** in nature.

```{r fig-rq4b, out.width="100%", results="asis"}
fig_include(file.path(fig_dir, "Figure6_Crisis_vs_RBC_Capital.png"))
```

</div>
</div>

---

## Supporting Finding &mdash; Profitability {.slide-section}

<span class="bdg-intro">Supporting Finding</span>

**Profitability Declined Persistently &mdash; Largest Institutions Hit Hardest**

<div class="abar">Return on assets fell across complex CUs &mdash; with the largest institutions experiencing the sharpest and most sustained declines. This finding directly informs the financial stability question in RQ5.</div>

<div class="two-col">
<div class="col-l">

<div class="stat-row" style="margin-bottom:12px;">
<div class="stat-card red"><div class="sn">-26bp</div><div class="sl">Average ROA decline &mdash; all complex CUs (p &lt; 0.01)</div></div>
<div class="stat-card red"><div class="sn">-47bp</div><div class="sl">ROA decline for CUs above $10B in assets</div></div>
</div>

- ROA decline is **persistent** &mdash; no evidence of adaptation or recovery through the observation window
- Largest CUs (above $10B in assets) experience a decline **nearly double the average**
- CCULR provided **no significant relief** on lending outcomes; CCULR adopters showed worse baseline outcomes consistent with distressed self-selection

</div>
<div class="col-r">

<div class="cb"><strong>Why this matters:</strong> Lower profitability reduces a credit union's ability to build capital organically. A CU that cannot retain earnings must either raise rates further, cut lending, or draw down its capital ratio &mdash; all three patterns appear in the data.</div>

<div class="cb" style="margin-top:8px;"><strong>CCULR finding:</strong> The Community Credit Union Leverage Ratio was designed as a simpler compliance pathway. The data suggest institutions most likely to adopt it were already under financial stress &mdash; CCULR adoption did not cause their worse outcomes, but did not reverse them either.</div>

```{r fig-roa, out.width="100%", results="asis"}
fig_include(file.path(fig_dir, "Figure3d_EventStudy_ROA.png"))
```

</div>
</div>

---

## RQ 5 &mdash; Failure Risk from Removing the Rule {.slide-section}

<span class="bdg-rq">Research Question 5</span>

**Does removing the RBC rule create unacceptable failure risk?**

<div class="abar">Answer: Removing the rule carries <strong>real but quantifiable and modest risk</strong>. At 2008 GFC severity, the rule reduces failure risk by 3.4 percentage points &mdash; protecting approximately 24 institutions.</div>

<div class="two-col">
<div class="col-l">

<div class="lbl">Capital Adequacy Stress Test &mdash; At 2008 GFC Severity</div>

<div class="stat-row" style="margin-bottom:12px;">
<div class="stat-card"><div class="sn sm">19.1%</div><div class="sl">Pre-rule failure rate (baseline)</div></div>
<div class="stat-card green"><div class="sn sm">19.4%</div><div class="sl">With-rule rate (&minus;0.3pp incremental protection)</div></div>
<div class="stat-card red"><div class="sn sm">22.5%</div><div class="sl">Post-removal (+3.4pp &asymp; 24 institutions)</div></div>
</div>

- The rule's **incremental protection is only &minus;0.3pp** &mdash; complex CUs were already building substantial buffers before the rule
- Removal raises simulated failure risk by **+3.4pp at 2008 severity**, corresponding to approximately **24 additional institutions**
- Each protected institution represents approximately **$14 billion in assets** sheltered from distress

</div>
<div class="col-r">

<div class="stat-row" style="margin-bottom:12px;">
<div class="stat-card"><div class="sn sm">$0.4B&ndash;$2.2B</div><div class="sl">Probability-weighted NCUSIF cost of removal</div></div>
</div>

<div class="lbl">Stress Scenario Sensitivity</div>

| Recession Severity | Welfare Retained | Member Cost |
|:-------------------|:----------------:|:-----------:|
| Moderate recession | 87% | $284BN |
| Severe recession | 73% | $238BN |

<div class="cb" style="margin-top:8px;">Even in a severe recession, the estimated member cost of the rule substantially outweighs the estimated insurance benefit under any plausible combination of assumptions tested. No scenario produces a reversal of this finding.</div>

</div>
</div>

---

## Summary &mdash; Five Questions, Five Answers {.slide-section}

<span class="bdg-sum">Summary</span>

```{r summary-table}
library(knitr)
library(kableExtra)
df <- data.frame(
  RQ = c("RQ 1","RQ 2","RQ 3","RQ 4","RQ 5"),
  Question = c(
    "How much capital buffer before the rule?",
    "Did complex CUs hold more capital?",
    "What did it cost members?",
    "Did the rule change what loans CUs make?",
    "Does removing the rule create unacceptable failure risk?"
  ),
  Finding = c(
    "Critically thin -- 72bp average buffer above the 10% threshold",
    "Mixed -- NW ratio +46bp, but well-capitalized probability -3.7pp",
    "Rates rose on every product; lending contracted persistently. $325.9BN total, $865/member/yr",
    "Yes -- auto share -4.0pp, RE share +2.4pp. Pattern unique to RBC rule, not seen in 2008",
    "Real but modest -- +3.4pp failure risk, ~24 institutions, NCUSIF cost $0.4-$2.2BN"
  ),
  stringsAsFactors = FALSE
)
kable(df, col.names = c("RQ", "Question", "Core Finding and Key Numbers"),
      align = c("c","l","l")) |>
  kable_styling(bootstrap_options = c("striped","hover"),
                full_width = TRUE, font_size = 14) |>
  column_spec(1, bold = TRUE, color = "#1a2744", width = "5%") |>
  column_spec(2, width = "28%") |>
  column_spec(3, bold = TRUE, color = "#1a2744") |>
  row_spec(0, background = "#1a2744", color = "white", bold = TRUE)
```

<div style="margin-top:16px;padding:12px 16px;border-left:4px solid #c8972a;background:#ede8df;font-size:13px;color:#444;border-radius:0 4px 4px 0;">
<strong>Across all five research questions</strong>, the findings are consistent: complex credit unions responded to the rule's capital incentive primarily through higher loan rates, reduced lending volume, portfolio reallocation, and lower profitability &mdash; rather than through material improvements in financial stability outcomes at the system level.
</div>
