---
title: "The 2022 NCUA Risk-Based Capital Rule"
subtitle: "Empirical Findings from an Analysis of Complex Credit Unions"
author: "NCUA Office of the Chief Economist"
date: "April 2026"
output:
  revealjs::revealjs_presentation:
    theme: simple
    highlight: pygments
    center: false
    transition: fade
    slide_level: 2
    self_contained: true
    reveal_options:
      width: 1100
      height: 660
      margin: 0.04
      minScale: 0.5
      maxScale: 1.5
      controls: true
      progress: true
      history: true
      keyboard: true
    css: rbc_slides.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)

# Helper: include a figure if it exists, otherwise show a placeholder
fig_include <- function(path, caption = "", width = "100%", height = "auto") {
  if (file.exists(path)) {
    knitr::include_graphics(path)
  } else {
    fname <- basename(path)
    cat(sprintf(
      '<div class="fig-placeholder"><span class="fig-label">FIGURE</span>
       <span class="fig-name">%s</span>
       <span class="fig-caption">%s</span></div>',
      fname, caption
    ))
  }
}

# Base figure directory — adjust if needed
fig_dir <- "output/paper/figures/"
```

<!-- ============================================================
     SLIDE 1 — TITLE
============================================================ -->

## {.title-slide data-background-color="#1a2744"}

<div class="title-org">NCUA &nbsp;·&nbsp; Office of the Chief Economist</div>

<div class="title-main">The 2022 NCUA Risk-Based Capital Rule</div>

<div class="title-sub">Empirical Findings from an Analysis of Complex Credit Unions</div>

<div class="title-rule"></div>

<div class="title-meta">
Saurabh C. Datta, Ph.D. &nbsp;·&nbsp; Senior Economist<br>
NCUA Call Report Data, 2000–2025 &nbsp;·&nbsp; April 2026
</div>

<!-- ============================================================
     SLIDE 2 — WHAT WE STUDIED + METHOD EXPLAINER
============================================================ -->

## What We Studied — and How {data-background-color="#f5f1eb"}

::: {.slide-cols}

::: {.col-wide}

**The Data**

- **696 complex credit unions** — those with average assets ≥ \$500M, classified using 2021 data — are the treated group
- **~4,488 non-complex credit unions** serve as the comparison group
- **150,000+ credit union–quarters** of NCUA Call Report data, 2000–2025
- **12 outcomes** measured: capital, loan rates, lending volume, portfolio mix, profitability, and loan quality

---

**The Method: Difference-in-Differences (DiD)**

::: {.method-diagram}
::: {.md-box}
**Change in complex CUs**  
Before → after the rule took effect
:::
::: {.md-minus}−:::
::: {.md-box}
**Change in non-complex CUs**  
Same time window
:::
::: {.md-equals}=:::
::: {.md-result}
**The Rule's Effect**  
What changed because of the rule, not the economy
:::
:::

::: {.callout}
**Why this works:** Non-complex credit unions faced the same economic conditions — interest rates, inflation, loan demand — but were not subject to the RBC rule. Any difference in outcomes between the two groups, after accounting for pre-existing trends, isolates the rule's effect.
:::

:::

::: {.col-divider}:::

::: {.col-narrow}

**How to Read the Event Study Charts**

::: {.intro-box}
- **Horizontal axis:** quarters relative to the rule's effective date (quarter 0). Left = before; right = after.
- **Vertical axis:** estimated difference between complex and non-complex credit unions, above any pre-existing gap.
- **Before quarter 0:** dots should sit near zero — confirming the two groups moved together before the rule *(parallel trends check)*.
- **After quarter 0:** a persistent move up or down is the rule's estimated effect. The shaded band is the 95% confidence interval.
- **If the band crosses zero** at a given quarter, the effect at that point is not statistically distinguishable from no effect.
:::

```{r how-to-read-schematic, fig.height=2.5, fig.width=4.5, out.width="100%"}
library(ggplot2)
library(tibble)

quarters <- c(-4, -2, 0, 2, 4, 6)
est      <- c(0.05, -0.03, 0.00, -0.45, -0.58, -0.62)
se       <- c(0.10,  0.10, 0.00,  0.10,  0.10,  0.10)

d <- tibble(q = quarters, e = est, lo = est - 1.96*se, hi = est + 1.96*se)

ggplot(d, aes(x = q, y = e)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "#999", linewidth = 0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "#c8972a", linewidth = 0.7) +
  geom_ribbon(aes(ymin = lo, ymax = hi), fill = "#1a2744", alpha = 0.15) +
  geom_line(color = "#1a2744", linewidth = 1.2) +
  geom_point(color = "#1a2744", size = 2.5, fill = "white", shape = 21, stroke = 1.5) +
  annotate("text", x = 0.2, y = 0.22, label = "Rule\neffective", hjust = 0,
           color = "#c8972a", size = 2.8, lineheight = 0.9) +
  annotate("text", x = -3.8, y = 0.22, label = "Flat pre-period ✓",
           hjust = 0, color = "#555", size = 2.8) +
  annotate("text", x = 1.8, y = -0.38, label = "Persistent\npost-rule effect",
           hjust = 0, color = "#1a2744", size = 2.8, lineheight = 0.9) +
  scale_x_continuous(breaks = quarters, labels = paste0(ifelse(quarters > 0, "+", ""), quarters, "Q")) +
  labs(x = NULL, y = NULL,
       caption = "Schematic illustration — actual charts follow on subsequent slides") +
  theme_minimal(base_size = 9) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "#e8e0d4", linewidth = 0.4),
    plot.caption = element_text(color = "#aaa", size = 7, hjust = 0.5),
    plot.background = element_rect(fill = "white", color = "#c8c0b0", linewidth = 0.5),
    axis.text = element_text(color = "#666")
  )
```

:::

:::

<!-- ============================================================
     SLIDE 3 — RQ1
============================================================ -->

## RQ 1 &nbsp;·&nbsp; Pre-Rule Capital Buffer {data-background-color="#f5f1eb"}

<div class="rq-label">Research Question 1</div>

**How much capital buffer did complex credit unions hold before the rule?**

::: {.answer-banner}
Answer: The average complex credit union entered the rule with a capital buffer of just **72 basis points** above the new 10% well-capitalized threshold — critically thin.
:::

::: {.slide-cols}

::: {.col-wide}

**What the Data Show**

- The 696 complex credit unions were, on average, nominally compliant before the rule — but with very little margin to absorb a new, higher threshold
- A **72 basis point buffer** means that a credit union holding 10.72% net worth would need to reach 10% under the new definition — but the definition of "well-capitalized" shifted upward simultaneously
- The capital distribution was **tightly clustered just above the old threshold**, with a large mass of institutions just barely on the compliant side
- This thin margin directly explains why the rule had such large downstream effects — institutions had to act immediately to avoid falling below the new line

:::

::: {.col-divider}:::

::: {.col-narrow}

::: {.callout}
**What is a "basis point"?** One basis point (bp) is one one-hundredth of a percentage point. A 72bp buffer means a credit union holding 10.72% capital, when the threshold is 10.00%.
:::

::: {.stat-row}
::: {.stat-card}
**72bp**  
Average pre-rule capital buffer above the 10% threshold
:::
::: {.stat-card .red}
**696**  
Complex credit unions subject to the rule  
*(Fixed 2021 classification)*
:::
:::

::: {.callout style="margin-top:12px;"}
**Why it matters:** A rule requiring meaningful additional capital reserves would always be more costly when institutions start with little room to spare. The thin pre-rule buffer set the stage for the compliance pressures documented in RQ2–RQ4.
:::

:::

:::

<!-- ============================================================
     SLIDE 4 — RQ2
============================================================ -->

## RQ 2 &nbsp;·&nbsp; Did Capital Levels Rise? {data-background-color="#f5f1eb"}

<div class="rq-label">Research Question 2</div>

**Did the rule cause complex credit unions to hold more capital?**

::: {.answer-banner}
Answer: **Mixed.** Average capital levels rose — but the share of institutions meeting the new "well-capitalized" standard actually fell. The bar rose faster than capital could accumulate.
:::

::: {.slide-cols}

::: {.col-wide}

::: {.stat-row}
::: {.stat-card .green}
**+46bp**  
Increase in net worth ratio  
*(p < 0.01)*
:::
::: {.stat-card .red}
**−3.7pp**  
Change in probability of being "well-capitalized"  
*(p < 0.01)*
:::
:::

**The Paradox Explained**

- Capital levels rose in absolute terms — institutions did respond to the rule's incentive to build reserves
- But the **definition of "well-capitalized" changed** at the same time, raising the threshold institutions must reach
- Result: more capital on average, but **fewer institutions clearing the new bar**

::: {.callout style="margin-top:10px;"}
**An analogy:** Think of a class where every student studied harder and raised their test score — but the passing grade was raised even more. The average score went up, yet more students are now technically failing. Both facts are true simultaneously.
:::

:::

::: {.col-divider}:::

::: {.col-narrow}

**What the Event Study Shows**

```{r fig-rq2, out.width="100%", results="asis"}
fig_include(
  path    = file.path(fig_dir, "Figure3_EventStudy_Capital.png"),
  caption = "4-panel: NW ratio, well-cap probability, tier 1 ratio, risk-based capital ratio"
)
```

::: {.fig-caption-note}
Pre-period flat ✓ confirms parallel trends. Post-period divergence is attributed to the rule.
:::

:::

:::

<!-- ============================================================
     SLIDE 5 — RQ3 (Rates + Volume combined)
============================================================ -->

## RQ 3 &nbsp;·&nbsp; Member Cost: Higher Rates and Less Lending {data-background-color="#f5f1eb"}

<div class="rq-label">Research Question 3</div>

**What did the rule cost members — in higher loan rates and reduced lending?**

::: {.answer-banner}
Answer: Loan rates rose across every product and loan growth contracted persistently. The estimated member cost totals **\$325.9 billion** over four years — approximately **\$865 per member per year**.
:::

::: {.slide-cols}

::: {.col-wide}

**Loan Rate Increases** *(vs. non-complex credit unions)*

::: {.stat-row}
::: {.stat-card .red}
**+75bp**  
Mortgage spread
:::
::: {.stat-card .red}
**+59bp**  
Auto loan spread
:::
::: {.stat-card .red}
**+58bp**  
Commercial spread
:::
:::

- All increases are **statistically significant** and persist through 10+ quarters — no sign of convergence
- A 75bp mortgage increase on a \$300,000 loan adds roughly **\$135/month** in payments

```{r fig-rq3-spreads, out.width="100%", results="asis"}
fig_include(
  path    = file.path(fig_dir, "Figure4_EventStudy_Spreads.png"),
  caption = "Loan spread event study — all products"
)
```

:::

::: {.col-divider}:::

::: {.col-narrow}

**Lending Volume and Total Member Cost**

::: {.stat-row}
::: {.stat-card .red}
**−0.44pp/qtr**  
Quarterly loan growth reduction  
*Persistent — no recovery*
:::
::: {.stat-card .red}
**\$325.9B**  
Total estimated member welfare cost  
*4-year, gradual scenario*
:::
:::

- Loan growth fell **0.44 percentage points per quarter** — compounding over time into a substantial reduction in credit availability
- The \$325.9BN estimate aggregates the rate-increase effect across all member-borrowers of all 696 complex credit unions over four years
- On a per-member basis: approximately **\$865/year** in higher borrowing costs

```{r fig-rq3-growth, out.width="100%", results="asis"}
fig_include(
  path    = file.path(fig_dir, "Figure3b_EventStudy_LoanGrowth.png"),
  caption = "Loan growth event study"
)
```

:::

:::

<!-- ============================================================
     SLIDE 6 — RQ4 (Portfolio + Crisis comparison combined)
============================================================ -->

## RQ 4 &nbsp;·&nbsp; Portfolio Shift and the Crisis Comparison {data-background-color="#f5f1eb"}

<div class="rq-label">Research Question 4</div>

**Did the rule change what loans credit unions make — and how does this compare to past stress?**

::: {.answer-banner}
Answer: Yes. Complex credit unions shifted away from auto loans toward mortgages — consistent with **risk-weight arbitrage**. This shift is unique to the RBC rule and did not occur during the 2008 financial crisis.
:::

::: {.slide-cols}

::: {.col-wide}

**Portfolio Reallocation**

::: {.stat-row}
::: {.stat-card .red}
**−4.0pp**  
Auto loan share  
*(relative to controls)*
:::
::: {.stat-card .green}
**+2.4pp**  
Real estate loan share
:::
:::

- **Risk-weight arbitrage:** Under the RBC rule, auto loans carry a higher risk weight than mortgages. By shifting to mortgages, complex CUs reduce their risk-weighted capital requirement — a rational regulatory response
- Loan quality also deteriorated: **delinquency rates rose +8bp**, driven by a capacity channel — the CUs with the largest profitability declines showed 2× the delinquency increase

```{r fig-rq4-portfolio, out.width="100%", results="asis"}
fig_include(
  path    = file.path(fig_dir, "Figure5_Portfolio_Trends.png"),
  caption = "Portfolio reallocation — auto vs. real estate loan share"
)
```

:::

::: {.col-divider}:::

::: {.col-narrow}

**Comparison to the 2008 Financial Crisis**

::: {.callout}
We ran the same DiD analysis using 2008 as the event — measuring how complex CUs responded to the financial crisis versus the RBC rule.
:::

| Outcome | 2008 Crisis | RBC Rule |
|:--------|:------------|:---------|
| Capital accumulation | Similar magnitude | Similar magnitude |
| Auto loan exit | Not observed | Significant & persistent |
| Loan growth contraction | Temporary | Persistent — no recovery |
| Profitability decline | Temporary | Persistent — no recovery |

- Capital behavior during RBC resembles a crisis response — but the **lending and profitability effects are distinctly regulatory** in nature

```{r fig-rq4-crisis, out.width="100%", results="asis"}
fig_include(
  path    = file.path(fig_dir, "Figure6_Crisis_vs_RBC_Capital.png"),
  caption = "Crisis vs. RBC comparison — capital and portfolio"
)
```

:::

:::

<!-- ============================================================
     SLIDE 7 — PROFITABILITY (supporting, bridge to RQ5)
============================================================ -->

## Supporting Finding &nbsp;·&nbsp; Profitability {data-background-color="#f5f1eb"}

<div class="badge-intro-label">Supporting Finding</div>

**Profitability Declined Persistently — Largest Institutions Hit Hardest**

::: {.answer-banner}
Return on assets fell across complex credit unions — with the largest institutions experiencing the sharpest and most sustained declines. This finding is directly relevant to understanding the rule's financial stability implications in RQ5.
:::

::: {.slide-cols}

::: {.col-wide}

**Key Estimates**

::: {.stat-row}
::: {.stat-card .red}
**−26bp**  
Average ROA decline (all complex CUs)  
*(p < 0.01)*
:::
::: {.stat-card .red}
**−47bp**  
ROA decline for CUs > \$10B in assets  
*Effect nearly doubles at largest institutions*
:::
:::

- The ROA decline is **persistent** across all post-rule quarters — no evidence of adaptation or recovery through the observation window
- The largest complex CUs — those with over \$10 billion in assets — experience a decline **nearly double the average**
- CCULR (an alternative compliance pathway) provided **no statistically significant relief** on lending outcomes; CCULR adopters showed worse baseline outcomes, consistent with distressed self-selection

:::

::: {.col-divider}:::

::: {.col-narrow}

::: {.callout}
**Lower profitability reduces a credit union's ability to build capital organically.** A credit union that cannot retain earnings must either raise rates further, cut lending, or run down its capital ratio — all three patterns appear in the data.
:::

::: {.callout style="margin-top:12px;"}
**The CCULR finding:** The Community Credit Union Leverage Ratio was designed as a simpler compliance pathway. The data suggest that the institutions most likely to adopt it were already under financial stress — so CCULR adoption itself does not appear to have caused their worse outcomes, but it also did not reverse them.
:::

```{r fig-roa, out.width="100%", results="asis"}
fig_include(
  path    = file.path(fig_dir, "Figure3d_EventStudy_ROA.png"),
  caption = "ROA event study"
)
```

```{r fig-cculr, out.width="100%", results="asis"}
fig_include(
  path    = file.path(fig_dir, "Figure7_CCULR_EventStudy.png"),
  caption = "CCULR three-group event study"
)
```

:::

:::

<!-- ============================================================
     SLIDE 8 — RQ5
============================================================ -->

## RQ 5 &nbsp;·&nbsp; Failure Risk from Removing the Rule {data-background-color="#f5f1eb"}

<div class="rq-label">Research Question 5</div>

**Does removing the RBC rule create unacceptable failure risk?**

::: {.answer-banner}
Answer: Removing the rule carries **real but quantifiable and modest risk**. At 2008 financial crisis severity, the rule reduces failure risk by 3.4 percentage points — protecting approximately 24 institutions.
:::

::: {.slide-cols}

::: {.col-wide}

**Capital Adequacy Stress Test — At 2008 GFC Severity**

::: {.stat-row}
::: {.stat-card}
**19.1%**  
Pre-rule failure rate  
*Baseline — before the rule*
:::
::: {.stat-card .green}
**19.4%**  
With-rule failure rate  
*Rule adds −0.3pp incremental protection*
:::
::: {.stat-card .red}
**22.5%**  
Post-removal failure rate  
*+3.4pp increase ≈ 24 institutions*
:::
:::

- The rule's **incremental protection is only −0.3pp** relative to pre-rule capital levels — because complex CUs were already building substantial buffers before the rule
- Removing the rule raises simulated failure risk by **+3.4pp at 2008 severity**, corresponding to approximately **24 additional institutions**
- Each institution protected represents approximately **\$14 billion in assets** sheltered from distress

:::

::: {.col-divider}:::

::: {.col-narrow}

**Expected Insurance Cost**

::: {.stat-row}
::: {.stat-card}
**\$0.36B–\$2.17B**  
Probability-weighted NCUSIF cost of removal  
*Accounts for crisis probability and loss-given-default*
:::
:::

**Stress Scenario Sensitivity**

| Recession Severity | Welfare Retained | Member Cost |
|:-------------------|:----------------:|:-----------:|
| Moderate recession | 87% | \$284BN |
| Severe recession   | 73% | \$238BN |

::: {.callout style="margin-top:12px;"}
Even in a severe recession scenario, the estimated member cost of the rule substantially outweighs the estimated insurance benefit under any plausible combination of assumptions tested. No scenario produces a reversal of this finding.
:::

```{r fig-rq5, out.width="100%", results="asis"}
fig_include(
  path    = file.path(fig_dir, "Figure9_9_Net_Benefit_Analysis.png"),
  caption = "Net benefit analysis — cost vs. protection"
)
```

:::

:::

<!-- ============================================================
     SLIDE 9 — SUMMARY TABLE
============================================================ -->

## Five Questions. Five Answers. {data-background-color="#f5f1eb"}

<div class="badge-summary-label">Summary</div>

```{r summary-table}
library(knitr)
library(kableExtra)

df <- data.frame(
  RQ = c("RQ 1", "RQ 2", "RQ 3", "RQ 4", "RQ 5"),
  Question = c(
    "How much capital buffer before the rule?",
    "Did complex CUs hold more capital?",
    "What did it cost members in higher rates and lower lending?",
    "Did the rule change what loans CUs make?",
    "Does removing the rule create unacceptable failure risk?"
  ),
  Finding = c(
    "Critically thin — institutions had very little room to absorb a higher threshold",
    "Mixed — average capital rose, but fewer institutions are now "well-capitalized"",
    "Rates rose on every loan product; lending growth contracted persistently and has not recovered",
    "Yes — shifted from auto to mortgage lending (risk-weight arbitrage). Pattern did not occur during the 2008 crisis.",
    "Real but modest and quantifiable — the rule protects approximately 24 institutions at 2008 crisis severity"
  ),
  Key_Number = c(
    "72bp average buffer above 10% threshold",
    "NW ratio +46bp · Well-cap probability −3.7pp",
    "$325.9BN total · $865/member/yr · Loan growth −0.44pp/qtr",
    "Auto share −4.0pp · RE share +2.4pp",
    "Failure risk +3.4pp · ~24 institutions · NCUSIF cost $0.4–$2.2BN"
  ),
  stringsAsFactors = FALSE
)

kable(
  df,
  col.names = c("RQ", "Question", "Core Finding", "Key Number"),
  align     = c("c", "l", "l", "l"),
  escape    = FALSE
) |>
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width        = TRUE,
    font_size         = 13
  ) |>
  column_spec(1, bold = TRUE, width = "4%",  color = "#1a2744") |>
  column_spec(2, width = "22%") |>
  column_spec(3, width = "38%", bold = TRUE, color = "#1a2744") |>
  column_spec(4, width = "36%") |>
  row_spec(0, background = "#1a2744", color = "white", bold = TRUE)
```

::: {.summary-footer-row}

::: {.summary-callout}
**Across all five research questions**, the findings are consistent: complex credit unions responded to the rule's capital incentive, but the adjustment came primarily through higher loan rates, reduced lending volume, portfolio reallocation, and lower profitability — rather than through material improvements in financial stability outcomes at the system level.
:::

::: {.summary-meta-box}
**Data & Methods**  
NCUA Form 5300 · 2000–2025  
Difference-in-differences  
696 treated · ~4,488 controls  
150,000+ CU-quarters  
12 outcome variables
:::

:::
