---
title: "The Impact of the 2022 NCUA Risk-Based Capital Rule on Complex Credit Unions"
subtitle: "Research Briefing — Internal Use"
date: "April 2026"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 2
    number_sections: false
    df_print: kable
    code_folding: hide
    self_contained: true
    css: exec_briefing.css
    pandoc_args:
      - "--wrap=none"
  # pdf_document: uncomment below to enable PDF output (requires tinytex)
params:
  fig_dir: "output/paper/figures"
  table_dir: "output/paper/tables"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo        = FALSE,
  warning     = FALSE,
  message     = FALSE,
  out.width   = "100%",
  fig.align   = "center",
  dpi         = 150
)

# ── Core packages ──────────────────────────────────────────────────────────────
library(tidyverse)
library(knitr)
library(kableExtra)
library(scales)

# ── Figure directory (set via params or default) ───────────────────────────────
# Run 4_Paper_Tables.R before knitting to populate output/paper/figures/
FIG_DIR   <- params$fig_dir
TABLE_DIR <- params$table_dir

# ── Helper: safe figure inclusion ─────────────────────────────────────────────
# Returns knitr::include_graphics() if file exists; placeholder message if not.
fig_include <- function(filename) {
  path <- file.path(FIG_DIR, filename)
  if (isTRUE(file.exists(path))) {
    tryCatch(
      include_graphics(path),
      error = function(e) {
        knitr::asis_output(paste0(
          "<div style='border:2px dashed #DDE3EE;border-radius:8px;",
          "padding:32px;text-align:center;color:#9BA8C0;background:#F8F9FC'>",
          "<p><strong>", filename, "</strong></p>",
          "<p style='font-size:12px'>Could not load: ", conditionMessage(e), "</p>",
          "</div>"
        ))
      }
    )
  } else {
    knitr::asis_output(paste0(
      "\n\n<div style='border:2px dashed #DDE3EE;border-radius:8px;",
      "padding:32px;text-align:center;color:#9BA8C0;background:#F8F9FC'>",
      "<p>&#128202; Figure not found: <strong>", filename, "</strong></p>",
      "<p style='font-size:12px'>Run <code>4_Paper_Tables.R</code> ",
      "to generate figures.</p>",
      "</div>\n\n"
    ))
  }
}

# ── Helper: load a CSV table safely ───────────────────────────────────────────
load_table <- function(filename) {
  path <- file.path(TABLE_DIR, filename)
  if (file.exists(path)) read_csv(path, show_col_types = FALSE)
  else tibble(Note = paste("Table not found:", filename, "— run 4_Paper_Tables.R"))
}

# ── Color palette (consistent with paper) ────────────────────────────────────
COL_NAVY  <- "#1B3A6B"
COL_RED   <- "#C94040"
COL_AMBER <- "#E8A838"
COL_BLUE  <- "#4A7CB5"
```

<style>
body { font-family: "Source Sans 3", "Helvetica Neue", Arial, sans-serif; }
h1.title { font-family: Georgia, serif; font-size: 28px; color: #1B3A6B; }
h2 { font-family: Georgia, serif; color: #1B3A6B; border-bottom: 2px solid #DDE3EE;
     padding-bottom: 6px; margin-top: 40px; }
h3 { color: #2A3A5E; font-size: 14px; text-transform: uppercase;
     letter-spacing: 1.5px; margin-top: 24px; font-weight: 700; }
.score-row { display: flex; gap: 16px; flex-wrap: wrap; margin: 20px 0; }
.score-card { flex: 1; min-width: 160px; border-radius: 8px; padding: 18px 16px;
              border: 1px solid #DDE3EE; background: #F4F6FA; border-top: 4px solid; }
.score-val  { font-family: Georgia, serif; font-size: 26px; font-weight: 700;
              color: #1B3A6B; line-height: 1; margin-bottom: 4px; }
.score-lbl  { font-size: 11.5px; color: #6B7A99; line-height: 1.4; }
.legend-row { display: flex; gap: 10px; flex-wrap: wrap; padding: 14px 20px;
              background: white; border: 1px solid #DDE3EE; border-radius: 6px;
              margin: 16px 0; align-items: center; }
.leg-lbl { font-size: 10px; text-transform: uppercase; letter-spacing: 2px;
           color: #6B7A99; font-weight: 700; margin-right: 4px; }
</style>


# Summary of Findings {#exec-summary}

<div style="background:#1B3A6B;color:white;padding:36px 40px;border-radius:10px;margin-bottom:28px;position:relative;overflow:hidden"><p style="font-size:10px;text-transform:uppercase;letter-spacing:3px;color:#E8A838;font-weight:700;margin:0 0 14px">Internal Briefing &middot; NCUA Office of the Chief Economist &middot; April 2026</p><p style="font-size:15px;color:rgba(255,255,255,0.65);max-width:700px;margin:0;line-height:1.65">The 2022 NCUA Risk-Based Capital rule imposed significantly higher capital requirements on the 696 credit unions with average assets above $500 million. Using a rigorous difference-in-differences design that controls for macroeconomic conditions, pre-existing trends, and thousands of smaller credit unions unaffected by the rule, this analysis documents the rule&#39;s effects across five outcome domains: capital adequacy, loan pricing, credit availability, portfolio composition, and profitability.</p></div>
<div class="score-row"><div class="score-card" style="border-top-color:#C94040"><div class="score-val">+73bp</div><div class="score-lbl">Mortgage spread increase over Treasury benchmarks</div></div><div class="score-card" style="border-top-color:#C94040"><div class="score-val">$325.9B</div><div class="score-lbl">Foregone 4-year member welfare under repeal baseline</div></div><div class="score-card" style="border-top-color:#1B3A6B"><div class="score-val">&#x2212;26bp</div><div class="score-lbl">Return on assets decline (31% of pre-rule mean)</div></div><div class="score-card" style="border-top-color:#E8A838"><div class="score-val">&#x2212;0.3pp</div><div class="score-lbl">Rule&#39;s actual protection at 2008-crisis stress level</div></div></div>
<div class="legend-row"><span class="leg-lbl">Inference key:</span><span style="background:#FFF0F0;color:#C94040;border:1px solid #C9404033;padding:3px 10px;border-radius:3px;font-size:10.5px;font-weight:700;font-family:monospace">&#x25BC; Concern</span><span style="background:#FFFBF0;color:#B8821A;border:1px solid #E8A83833;padding:3px 10px;border-radius:3px;font-size:10.5px;font-weight:700;font-family:monospace">&#x25C6; Mixed</span><span style="background:#F0FFF5;color:#2E7D4F;border:1px solid #2E7D4F33;padding:3px 10px;border-radius:3px;font-size:10.5px;font-weight:700;font-family:monospace">&#x25B2; Positive</span><span style="background:#F0F5FF;color:#2A5A9B;border:1px solid #4A7CB533;padding:3px 10px;border-radius:3px;font-size:10.5px;font-weight:700;font-family:monospace">&#x25CF; Baseline</span><span style="background:#FFF8F0;color:#8B4513;border:1px solid #8B451333;padding:3px 10px;border-radius:3px;font-size:10.5px;font-weight:700;font-family:monospace">&#x2717; No effect</span><span style="font-size:11px;color:#6B7A99;font-style:italic;margin-left:4px">&#8212; appears in every Key Numbers and Model Results table</span></div>

**Central finding:** The rule imposed large, persistent costs on member-borrowers through higher loan rates, reduced credit availability, and deteriorating profitability — while providing only marginal protective benefit over and above what existed before the rule. An empirically calibrated threshold of **8.5%**, rather than 10%, would deliver equivalent protection against institutional failure at approximately one-third the compliance burden. Finding 12 explains *why* the rule's effects were larger and more persistent than the 2008 financial crisis itself — through five structural mechanisms that distinguish permanent regulatory cost changes from temporary economic shocks.

---

# Part I: Research Questions & Answers {#rqs}


---

## Research Question 1 {#rq1}

<div style="background:#1B3A6B;color:white;padding:20px 24px;border-radius:8px;margin-bottom:16px">
<p style="font-size:10px;text-transform:uppercase;letter-spacing:2.5px;opacity:0.7;margin:0 0 6px;font-family:monospace">Research Question 1</p>
<p style="font-family:Georgia,serif;font-size:18px;font-weight:700;margin:0;line-height:1.35">What was the average capital buffer for complex credit unions before the RBC rule?</p>
</div>

<div style="display:flex;gap:14px;margin-bottom:16px;flex-wrap:wrap">
<div style="flex:1;min-width:260px;border-left:5px solid #1B3A6B;background:#F4F6FA;padding:14px 18px;border-radius:0 6px 6px 0">
<p style="font-size:9.5px;text-transform:uppercase;letter-spacing:1.5px;color:#6B7A99;font-weight:700;margin:0 0 4px">Answer</p>
<p style="font-family:Georgia,serif;font-size:15px;font-weight:700;color:#1A2035;margin:0">72 basis points — critically thin</p>
</div>
<div style="background:#F0F5FF;border:1px solid #4A7CB544;padding:14px 18px;border-radius:6px;min-width:200px">
<p style="font-size:11px;font-weight:700;color:#2A5A9B;font-family:monospace;margin:0 0 4px">●&nbsp;Baseline</p>
<p style="font-size:11px;color:#2A5A9B;margin:0;line-height:1.4">Baseline context — explains why the rule had such bite</p>
</div>
</div>

Before the rule took effect, the average complex credit union — that is, any credit union with more than $500 million in assets, of which there were 696 — held a net worth ratio of 10.72%. The RBC rule required these institutions to maintain at least 10% to be classified as "well-capitalized." That left the average institution with a buffer of just 72 basis points (less than three-quarters of one percent) above the new threshold. Nearly half (47.6%) of complex credit unions were already below 10%, meaning they entered the rule period already falling short of the new standard and facing immediate pressure to build capital. This thin buffer is what gave the rule its bite — and explains both why some institutions struggled to comply and why compliance was so costly.

*→ See [Finding 1: Capital Built — But Many Institutions Still Can't Cross the Bar](#finding1)*


---

## Research Question 2 {#rq2}

<div style="background:#1B3A6B;color:white;padding:20px 24px;border-radius:8px;margin-bottom:16px">
<p style="font-size:10px;text-transform:uppercase;letter-spacing:2.5px;opacity:0.7;margin:0 0 6px;font-family:monospace">Research Question 2</p>
<p style="font-family:Georgia,serif;font-size:18px;font-weight:700;margin:0;line-height:1.35">Are complex credit unions now holding larger capital buffers?</p>
</div>

<div style="display:flex;gap:14px;margin-bottom:16px;flex-wrap:wrap">
<div style="flex:1;min-width:260px;border-left:5px solid #1B3A6B;background:#F4F6FA;padding:14px 18px;border-radius:0 6px 6px 0">
<p style="font-size:9.5px;text-transform:uppercase;letter-spacing:1.5px;color:#6B7A99;font-weight:700;margin:0 0 4px">Answer</p>
<p style="font-family:Georgia,serif;font-size:15px;font-weight:700;color:#1A2035;margin:0">Yes on average — but the average hides a critical distributional problem</p>
</div>
<div style="background:#FFFBF0;border:1px solid #E8A83844;padding:14px 18px;border-radius:6px;min-width:200px">
<p style="font-size:11px;font-weight:700;color:#B8821A;font-family:monospace;margin:0 0 4px">◆&nbsp;Mixed</p>
<p style="font-size:11px;color:#B8821A;margin:0;line-height:1.4">Average improved, pass rate declined — the rule worked for strong CUs, hurt thin-buffer CUs</p>
</div>
</div>

The average net worth ratio rose by 46 basis points relative to trend. However, the probability of being classified as 'well-capitalized' actually fell by 3.7 percentage points. Institutions that were already well above 10% built larger buffers, while institutions near the threshold — the ones the rule was most designed to protect — actually saw capital ratios decline. Among near-threshold credit unions ($400M–$600M in assets), capital ratios fell by 44 basis points.

*→ See [Finding 1: Capital Built — But Many Institutions Still Can't Cross the Bar](#finding1)*


---

## Research Question 3 {#rq3}

<div style="background:#C94040;color:white;padding:20px 24px;border-radius:8px;margin-bottom:16px">
<p style="font-size:10px;text-transform:uppercase;letter-spacing:2.5px;opacity:0.7;margin:0 0 6px;font-family:monospace">Research Question 3</p>
<p style="font-family:Georgia,serif;font-size:18px;font-weight:700;margin:0;line-height:1.35">Did loan rate spreads change significantly after the RBC rule?</p>
</div>

<div style="display:flex;gap:14px;margin-bottom:16px;flex-wrap:wrap">
<div style="flex:1;min-width:260px;border-left:5px solid #C94040;background:#F4F6FA;padding:14px 18px;border-radius:0 6px 6px 0">
<p style="font-size:9.5px;text-transform:uppercase;letter-spacing:1.5px;color:#6B7A99;font-weight:700;margin:0 0 4px">Answer</p>
<p style="font-family:Georgia,serif;font-size:15px;font-weight:700;color:#1A2035;margin:0">Yes — across every loan type, by large and economically meaningful amounts</p>
</div>
<div style="background:#FFF0F0;border:1px solid #C9404044;padding:14px 18px;border-radius:6px;min-width:200px">
<p style="font-size:11px;font-weight:700;color:#C94040;font-family:monospace;margin:0 0 4px">▼&nbsp;Concern</p>
<p style="font-size:11px;color:#C94040;margin:0;line-height:1.4">Complete cost pass-through to member-borrowers confirmed</p>
</div>
</div>

Spreads over Treasury benchmarks rose by 73 basis points for mortgages, 56 basis points for new auto loans, 76 basis points for used auto loans, and 66 basis points for commercial non-real-estate loans. All effects are statistically significant at the 0.1% level. The pass-through appears complete — credit unions raised rates to cover their higher capital costs rather than absorbing them, placing the entire regulatory burden on member-borrowers.

*→ See [Finding 2: Loan Rates Rose Sharply Across Every Product Type](#finding2)*


---

## Research Question 4 {#rq4}

<div style="background:#C94040;color:white;padding:20px 24px;border-radius:8px;margin-bottom:16px">
<p style="font-size:10px;text-transform:uppercase;letter-spacing:2.5px;opacity:0.7;margin:0 0 6px;font-family:monospace">Research Question 4</p>
<p style="font-family:Georgia,serif;font-size:18px;font-weight:700;margin:0;line-height:1.35">Did asset and loan growth change significantly?</p>
</div>

<div style="display:flex;gap:14px;margin-bottom:16px;flex-wrap:wrap">
<div style="flex:1;min-width:260px;border-left:5px solid #C94040;background:#F4F6FA;padding:14px 18px;border-radius:0 6px 6px 0">
<p style="font-size:9.5px;text-transform:uppercase;letter-spacing:1.5px;color:#6B7A99;font-weight:700;margin:0 0 4px">Answer</p>
<p style="font-family:Georgia,serif;font-size:15px;font-weight:700;color:#1A2035;margin:0">Yes — both fell sharply and have shown no recovery</p>
</div>
<div style="background:#FFF0F0;border:1px solid #C9404044;padding:14px 18px;border-radius:6px;min-width:200px">
<p style="font-size:11px;font-weight:700;color:#C94040;font-family:monospace;margin:0 0 4px">▼&nbsp;Concern</p>
<p style="font-size:11px;color:#C94040;margin:0;line-height:1.4">Permanent supply-side lending contraction — members losing access to credit</p>
</div>
</div>

Loan growth fell by approximately 0.44 percentage points per quarter — equivalent to about 1.75 percentage points per year. Asset growth fell by 0.47 percentage points per quarter. These effects appeared at the moment the rule took effect and have not reversed through the full 10+ quarter observation window. Near-threshold credit unions experienced an even larger contraction of 0.6 percentage points per quarter.

*→ See [Finding 3: Lending Growth Fell and Never Recovered](#finding3)*


---

## Research Question 5 {#rq5}

<div style="background:#2E7D4F;color:white;padding:20px 24px;border-radius:8px;margin-bottom:16px">
<p style="font-size:10px;text-transform:uppercase;letter-spacing:2.5px;opacity:0.7;margin:0 0 6px;font-family:monospace">Research Question 5</p>
<p style="font-family:Georgia,serif;font-size:18px;font-weight:700;margin:0;line-height:1.35">Does removing the RBC rule create unacceptable risk of institutional failure?</p>
</div>

<div style="display:flex;gap:14px;margin-bottom:16px;flex-wrap:wrap">
<div style="flex:1;min-width:260px;border-left:5px solid #2E7D4F;background:#F4F6FA;padding:14px 18px;border-radius:0 6px 6px 0">
<p style="font-size:9.5px;text-transform:uppercase;letter-spacing:1.5px;color:#6B7A99;font-weight:700;margin:0 0 4px">Answer</p>
<p style="font-family:Georgia,serif;font-size:15px;font-weight:700;color:#1A2035;margin:0">Some risk — but the threshold calibration shows 8.5% achieves the same protection at one-third the cost</p>
</div>
<div style="background:#FFFBF0;border:1px solid #E8A83844;padding:14px 18px;border-radius:6px;min-width:200px">
<p style="font-size:11px;font-weight:700;color:#B8821A;font-family:monospace;margin:0 0 4px">◆&nbsp;Mixed</p>
<p style="font-size:11px;color:#B8821A;margin:0;line-height:1.4">Repeal adds 3.4pp failure risk, but 8.5% threshold delivers equivalent protection to only 15% of CUs vs. 47% today</p>
</div>
</div>

Empirical stress tests using actual 2008 crisis data for 634 complex credit unions show that full repeal would increase the number of institutions falling below minimum capital requirements by 3.4 percentage points — approximately 23 additional institutions crossing into regulatory distress territory. However, this is not an argument for maintaining the current rule unchanged. The threshold calibration analysis (Finding 11) shows that a well-capitalized threshold of 8.5% would provide equivalent protection against institutional failure while constraining only 15% of complex credit unions — one-third the burden of the current 47%. The answer to whether repeal risk is acceptable depends entirely on which version of the rule is under consideration.

*→ See [Finding 11: The 10% Threshold Was Never Empirically Justified — The Data Say 8.5%](#finding11)*


---

# Part II: Findings 1–12 — The Full Evidence {#findings}

<div style="background:#F0F5FF;border:1px solid #4A7CB544;border-radius:8px;padding:18px 22px;margin-bottom:20px">
<p style="font-size:10px;text-transform:uppercase;letter-spacing:2px;color:#4A7CB5;font-weight:700;margin:0 0 10px">How to Read the Numbers in This Section</p>
<p style="font-size:13px;color:#2A3A5E;line-height:1.7;margin:0 0 8px">All estimates in Part II come from a <strong>difference-in-differences (DiD)</strong> statistical design. The idea is simple: we compare the 696 large (complex) credit unions affected by the RBC rule against roughly 4,500 smaller credit unions that were not affected. Both groups face the same economy, the same interest rate environment, and the same member behavior — so any difference in outcomes that opens up after the rule takes effect can be attributed to the rule itself, not to outside conditions.</p>
<p style="font-size:13px;color:#2A3A5E;line-height:1.7;margin:0 0 8px">When you see an estimate like <strong>+0.75 percentage points (pp)</strong> for mortgage spreads, it means: after the rule took effect, complex credit unions charged members 0.75 percentage points more on mortgages than comparable non-complex credit unions did — holding all other factors constant. When you see <strong>p &lt; 0.001</strong>, it means there is less than a 1-in-1,000 chance that this difference arose by random chance. Three stars (***) means the result clears that bar comfortably.</p>
<p style="font-size:13px;color:#2A3A5E;line-height:1.7;margin:0">The pre-period flatness you will see in each event-study chart — where the two groups track each other closely before the rule — is what validates the comparison. If complex and non-complex credit unions were already diverging before 2022, the design would be unreliable. The fact that they were not is what gives us confidence that what we measure after 2022 is genuinely caused by the rule.</p>
</div>

Each finding section contains:
(a) **Context** — why this question matters, in plain English  
(b) **Key Numbers** — the most important statistics, with an inference column  
(c) **Supporting Chart** — with a plain-English walk-through of what to look for  
(d) **Statistical Model Results** — the formal estimates with signal and meaning columns  
(e) **Bottom Line** — the plain-English conclusion for this finding

---


---

## Finding 1: Capital Built — But Many Institutions Still Can't Cross the Bar {#finding1}

<div style="border-left:6px solid #1B3A6B;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#1B3A6B;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 1</span>&ensp;<span style="color:#1B3A6B;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Capital</span>
</div>

### Context

The RBC rule required complex credit unions (those with more than $500 million in assets) to maintain a net worth ratio of at least 10% to be officially classified as **"well-capitalized"** — a higher bar than the prior 7% threshold. Being well-capitalized matters because it is not merely a label: it determines whether an institution can operate freely, pay dividends, expand, and avoid heightened supervisory scrutiny. An institution that falls below well-capitalized status faces a growing list of restrictions on what it can do. The question is whether this stricter requirement actually produced the capital-building it was designed to achieve.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>Average capital increase (trend-adj.)</td><td><strong>+0.46 percentage points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFFBF0;color:#B8821A;border:1px solid #E8A83844;font-size:10px;font-weight:700;font-family:monospace">◆&nbsp;Mixed</span><br><small style="color:#6B7A99">Rose, but masks a distribution problem</small></td></tr>
<tr><td>Well-capitalized probability change</td><td><strong>−3.7 percentage points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Fewer institutions actually passed the bar</small></td></tr>
<tr><td>Pre-rule capital buffer above 10%</td><td><strong>72 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0F5FF;color:#2A5A9B;border:1px solid #4A7CB544;font-size:10px;font-weight:700;font-family:monospace">●&nbsp;Baseline</span><br><small style="color:#6B7A99">Critically thin starting position</small></td></tr>
<tr><td>Near-threshold CU capital change</td><td><strong>−0.44 pp (declined)</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Thin-buffer CUs got worse, not better</small></td></tr>
<tr><td>Institutions studied</td><td><strong>696 complex CUs</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0F5FF;color:#2A5A9B;border:1px solid #4A7CB544;font-size:10px;font-weight:700;font-family:monospace">●&nbsp;Baseline</span><br><small style="color:#6B7A99">Full treatment group</small></td></tr>
</tbody></table>

### Supporting Chart

**How to read this chart:** This is a 4-panel figure. Each panel shows a different aspect of the capital story, all plotted on the same time axis: quarters before and after the rule took effect (quarter zero, marked by a dashed vertical line). In each panel, the vertical axis shows the DiD estimate — the difference between complex and non-complex credit unions — so a value of zero means both groups behaved identically. The dots before quarter zero should hover near zero if both groups were on similar trajectories before the rule; after quarter zero, any sustained movement away from zero reflects the rule's effect.

**Panel A — Net Worth Ratio:** This is the headline capital measure — the total equity of a credit union as a percentage of its total assets. A positive value after Q0 means complex credit unions built more capital relative to the control group. You will see the line trend upward after 2022, confirming the rule did raise average capital levels. But this is the *average* — it masks the distributional problem shown in Panel B.

**Panel B — Capital Buffer vs 10%:** This is the key panel for understanding *who* benefited. It does not show the net worth ratio itself — it shows the *gap* between each institution's net worth ratio and the 10% threshold. Think of it as headroom: a positive buffer means an institution is above the threshold; a negative buffer means it is below. When this DiD coefficient is negative after Q0, it means complex credit unions are seeing their headroom *shrink* relative to the control group — even as the average rises in Panel A. This is the paradox: the rule pushed the overall average up but compressed the buffer for the institutions nearest the threshold. Those institutions were being forced to absorb compliance costs (higher capital charges, tighter lending margins) that ate into their retained earnings faster than the rule helped them build capital. The strong institutions sailed further above 10%; the thin-buffer institutions drifted closer to or below it.

**Panel C — Well-Capitalized:** This shows the probability of an institution being classified as well-capitalized (i.e., above 10%). A negative value after Q0 means complex credit unions became *less* likely to achieve well-capitalized status — confirming the paradox. Despite the average rising (Panel A), fewer institutions actually crossed the bar that matters.

**Panel D — Loan Growth:** Shown here for context — the lending contraction that accompanied the capital pressure (covered in detail in Finding 3).

```{r fig1, error=TRUE, results='asis', fig.cap="Event study: capital outcomes (4-panel). Panel A: average NW ratio rises post-2022. Panel B: capital buffer vs. 10% threshold narrows — institutions nearest the bar lost headroom. Panel C: well-capitalized probability falls. Panel D: loan growth contracts persistently.", out.width="100%"}
fig_include("Figure3_EventStudy_Capital.png")
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Net worth ratio</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.463pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=156,121</td><td style="padding:8px 11px;background:#F4F6FA;color:#4A5568;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">↑ Mixed</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Average rose, but distribution shifted adversely</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Well-capitalized (LPM)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.037pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p=0.004</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=156,121</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Fewer CUs achieved well-capitalized status</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Near-threshold NW ratio</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.444pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p=0.003</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=19,140</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Opposite sign — rule hurt the most exposed CUs</td></tr>
</tbody></table></div>

### Bottom Line

<div style="border:1px solid #1B3A6B;background:#1B3A6B0D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#1B3A6B;font-weight:700;margin-bottom:8px">Bottom Line</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">The rule raised the average capital ratio by 46 basis points — which looks like progress. But the fraction of institutions actually achieving well-capitalized status fell by 3.7 percentage points. These two facts are not contradictory: institutions that were already comfortably above 10% built larger buffers, pulling the average up. Meanwhile, the institutions that were closest to the threshold — the ones the rule was most designed to protect — saw their capital ratios decline. Compliance costs consumed their retained earnings faster than they could rebuild them. The rule strengthened the strong and further strained the strained.</p>
</div>


---

## Finding 2: Loan Rates Rose Sharply Across Every Product Type {#finding2}

<div style="border-left:6px solid #C94040;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#C94040;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 2</span>&ensp;<span style="color:#C94040;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Member Cost</span>
</div>

### Context

When a credit union faces higher capital requirements, one way it adjusts is by charging higher interest rates on loans — passing the regulatory cost directly to members. We measure this as the spread over comparable Treasury rates, which removes any market-wide rate changes.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>Mortgage spread increase</td><td><strong>+73 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Members pay ~$1,875 more per year on avg mortgage</small></td></tr>
<tr><td>New auto loan spread</td><td><strong>+56 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Members pay ~$176 more per year on avg auto loan</small></td></tr>
<tr><td>Used auto loan spread</td><td><strong>+76 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Larger than new auto — reflects portfolio pressure</small></td></tr>
<tr><td>Commercial non-RE spread</td><td><strong>+66 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Business borrowers also affected</small></td></tr>
<tr><td>Effect for largest CUs (>$10B)</td><td><strong>+97bp auto spread</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Bigger institutions charged members more</small></td></tr>
</tbody></table>

### Supporting Chart

**How to read this chart:** The figure has three separate panels — Panel A for mortgage spreads, Panel B for new auto loan spreads, and Panel C for commercial non-real-estate loan spreads. In each panel, the horizontal axis is time in quarters relative to when the rule took effect (quarter zero, marked by a dashed vertical line). The vertical axis is the DiD estimate — the difference in loan pricing between complex and non-complex credit unions after stripping out any economy-wide rate movements. A value of zero means both groups are pricing identically. A positive value means complex credit unions are charging members more than comparable non-complex credit unions by that amount, in percentage points.

**What you will see:** In each panel, the line starts in negative territory before quarter zero — this reflects the fact that complex credit unions were actually charging *less* than non-complex ones in the pre-rule period, likely because their scale and member focus allowed them to offer more competitive rates. At quarter zero the line begins rising sharply, crosses through zero, and climbs into strongly positive territory in the post-rule period — where it stays. The shaded band around each line is the confidence interval, showing the range within which the true effect almost certainly falls. Because the shaded band stays well above zero throughout the post-rule period, we can be confident the increases are real and not statistical noise. The pattern is consistent across all three loan types: after the rule, members at complex credit unions started paying more than they would have at a comparable institution not subject to the rule — reversing a pre-rule pricing advantage.

```{r fig2, error=TRUE, results='asis', fig.cap="Loan rate spreads: complex vs. non-complex credit unions, mortgage, new auto, and commercial non-RE. Pre-rule: complex CUs charged less (negative values). Post-rule: complex CUs charge more (positive values) — a reversal driven by the rule.", out.width="100%"}
fig_include("Figure4_EventStudy_Spreads.png")
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Mortgage spread</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.750pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=24,927</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Large, highly significant, fully passed through</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">New auto spread</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.587pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=33,335</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Consistent across all institution sizes</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Used auto spread</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.779pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=33,678</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Largest spread effect in absolute terms</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Comm non-RE spread</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.618pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=9,057</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Business lending affected equally</td></tr>
</tbody></table></div>

### Bottom Line

<div style="border:1px solid #C94040;background:#C940400D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#C94040;font-weight:700;margin-bottom:8px">Bottom Line</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">The rate increases are large and consistent across every loan type we could measure. A 73 basis-point increase in mortgage spreads means a member borrowing $220,000 pays approximately $1,875 more per year — for the life of the loan. The increases appear immediately after the rule takes effect and do not reverse. Larger institutions charged even more: credit unions above $10 billion raised auto spreads by 97 basis points, well above the average.</p>
</div>


---

## Finding 3: Lending Growth Fell and Never Recovered {#finding3}

<div style="border-left:6px solid #C94040;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#C94040;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 3</span>&ensp;<span style="color:#C94040;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Credit Access</span>
</div>

### Context

Higher capital requirements can reduce the volume of lending — if a credit union must hold more capital per dollar of loans, it may simply make fewer loans. This affects how many members can access credit, not just the price they pay.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>Quarterly loan growth decline</td><td><strong>−0.44 pp per quarter</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Roughly 1.75pp per year in annualized terms</small></td></tr>
<tr><td>Quarterly asset growth decline</td><td><strong>−0.47 pp per quarter</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Broad-based balance-sheet contraction</small></td></tr>
<tr><td>Near-threshold loan growth</td><td><strong>−0.60 pp per quarter</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">50% larger for most constrained CUs</small></td></tr>
<tr><td>Comparison to 2008 crisis</td><td><strong>~60% as large as crisis</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFFBF0;color:#B8821A;border:1px solid #E8A83844;font-size:10px;font-weight:700;font-family:monospace">◆&nbsp;Mixed</span><br><small style="color:#6B7A99">Regulatory contraction rivals financial crisis</small></td></tr>
<tr><td>Duration of effect</td><td><strong>Persistent — no recovery</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">No reversal through 10+ quarters post-rule</small></td></tr>
</tbody></table>

### Supporting Chart

**How to read this chart:** The vertical axis shows the DiD coefficient for loan growth (QoQ log×100) — the difference in loan growth rate between complex and non-complex credit unions each quarter. Zero means the two groups were growing at identical rates. The horizontal axis is time relative to the rule's implementation date. The shaded band around the line is the 95% confidence interval — the wider it is, the less precise the estimate.

**What you will see:** The pre-period (left of quarter zero) shows some volatility — this is normal for quarterly growth rates, which are naturally noisy. What matters for causal identification is that the pre-period coefficients are statistically not different from zero on average — the dashed flat line at zero is inside the confidence band. After quarter zero, the line drops sharply into negative territory and the stars on the post-period dots confirm statistical significance. The line stays below zero through Q+8 and beyond, with no recovery trend. This persistent negative gap — complex credit unions growing their loan books more slowly than comparable non-complex ones, quarter after quarter — is the signature of a structural lending constraint, not a temporary adjustment.

```{r fig3, error=TRUE, results='asis', fig.cap="Loan growth event study. Sharp, persistent step-down at RBC implementation with no subsequent recovery — a structural lending constraint, not a temporary adjustment.", out.width="100%"}
fig_include("Figure3b_EventStudy_LoanGrowth.png")
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Loan growth</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.004***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=150,937</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Significant, persistent, no V-shape recovery</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Asset growth</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.005***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=150,937</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Broad contraction, not just loan-specific</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Near-threshold loans</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.006***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p=0.009</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=18,520</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Largest effect for most exposed institutions</td></tr>
</tbody></table></div>

### Bottom Line

<div style="border:1px solid #C94040;background:#C940400D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#C94040;font-weight:700;margin-bottom:8px">Bottom Line</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">The lending contraction is not a temporary adjustment — it is a permanent step-down in the growth trajectory of these institutions. Every quarter, complex credit unions grew their loan books about 44 basis points more slowly than comparable non-complex institutions. This effect is driven by supply — the credit unions constrained their lending — not by reduced member demand. Members who needed loans from these institutions were less likely to get them.</p>
</div>


---

## Finding 4: Portfolio Shifted Away from Auto Loans Toward Mortgages — Risk-Weight Arbitrage {#finding4}

<div style="border-left:6px solid #E8A838;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#E8A838;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 4</span>&ensp;<span style="color:#E8A838;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Portfolio Risk</span>
</div>

### Context

Under the RBC rule, not all loans are treated equally when counting toward capital requirements. The rule assigns each loan category a "capital cost" based on its assessed riskiness — auto loans are treated as more capital-intensive to hold, while qualifying residential mortgages are treated as significantly less so. This creates a straightforward financial incentive: a credit union under pressure to build capital can reduce how much capital it needs — without making a single new loan — simply by shifting its existing portfolio away from auto loans toward mortgages. The rule effectively places a higher regulatory price on auto lending than on mortgage lending.

**A question we anticipated: doesn't this just reflect the Federal Reserve's rate-raising cycle?** The Fed began raising rates aggressively in 2022, which could independently affect portfolio composition across all credit unions — not just the large ones subject to the RBC rule. Our design specifically accounts for this. Here is how: we do not simply measure how much complex credit unions shifted their auto loan share. We measure how much they shifted *relative to* the roughly 4,500 smaller non-complex credit unions that faced exactly the same interest rate environment but were not subject to the RBC rule. Any economy-wide effect of Fed rate changes — including any tendency for all credit unions to move away from auto loans — cancels out in this comparison, because it affects both groups equally. What remains after that cancellation is the **4.0 percentage-point** auto share decline — the portion of the shift that is specific to complex credit unions above and beyond what non-complex ones did in the same environment. That residual cannot be explained by interest rates.

To further rule out the possibility that this 4.0pp shift reflects some other form of general financial stress — not interest rates specifically, but economic pressure more broadly — we compared it against what happened during the 2008 financial crisis, when complex credit unions faced genuine severe stress. During the crisis, complex credit unions reduced their auto loan share by about 1.8 percentage points. Under the RBC rule, in a stable economic environment with no financial crisis, they reduced it by 4.0 percentage points — more than double. The difference between those two episodes, **2.4 percentage points**, is the amount by which the rule's portfolio effect exceeded even a major financial crisis. That 2.4pp gap is statistically confirmed (p < 0.001) and tells us the portfolio shift is regulatory in origin — not a response to economic conditions of any kind.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>Auto loan share decline</td><td><strong>−4.0 percentage points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Doubled the auto exit seen in 2008 crisis</small></td></tr>
<tr><td>Real estate loan share rise</td><td><strong>+2.4 percentage points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFFBF0;color:#B8821A;border:1px solid #E8A83844;font-size:10px;font-weight:700;font-family:monospace">◆&nbsp;Mixed</span><br><small style="color:#6B7A99">Flight to lower-risk-weight assets</small></td></tr>
<tr><td>MBL share decline</td><td><strong>−0.25 percentage points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Business lending also reduced</small></td></tr>
<tr><td>Auto exit vs. 2008 crisis</td><td><strong>2.2pp LARGER under RBC</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Regulatory-specific — not generic stress</small></td></tr>
<tr><td>Portfolio reversal by year 4</td><td><strong><30% reversal</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Structural — not reverting without rule change</small></td></tr>
</tbody></table>

### Supporting Chart

**How to read this chart:** This is a 4-panel figure showing actual portfolio shares over calendar time (not event-time coefficients). Each panel shows the share of total loans in a category from 2018 to 2026. Two lines appear in each panel: dark (complex credit unions) and red/flat (non-complex credit unions). The dashed vertical line marks the RBC effective date (2022 Q1). If the shift is regulatory-driven, you expect the dark line to diverge from the red line at the dashed vertical — while the red line stays flat.

**What you will see:** Panel A (Member Business Loans) and Panel B (Commercial Loans) show the dark line rising after 2022 while the red line stays flat — complex credit unions are growing their commercial exposure. Panel C (Real Estate Loans) shows the same pattern — rising after 2022 for complex institutions. Panel D (Auto Loans) shows the most dramatic shift: the dark line drops sharply after 2022 and keeps falling, while the red line stays elevated and flat. This is the fingerprint of risk-weight arbitrage — complex institutions are systematically exiting the highest-risk-weight asset (auto loans) and shifting toward lower-weight alternatives, driven purely by the regulatory cost structure.

```{r fig4, error=TRUE, results='asis', fig.cap="Portfolio composition around RBC. Auto share falls sharply and permanently; RE share rises — a textbook risk-weight arbitrage pattern not seen in the 2008 crisis.", out.width="100%"}
fig_include("Figure5_Portfolio_Trends.png")
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Auto share</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−4.004pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=155,946</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Largest effect in study — regulatory arbitrage</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">RE share</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+2.385pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=155,946</td><td style="padding:8px 11px;background:#FFFBF0;color:#B8821A;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">◆ Mixed</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Lower risk-weight, but more recession-sensitive</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">MBL share</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.247pp**</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p=0.026</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=155,946</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Business lending reduced — serves fewer small biz</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Auto: RBC vs crisis Wald</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">2.406pp larger***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">Wald test</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Confirms regulatory-specific, not generic, effect</td></tr>
</tbody></table></div>

### Bottom Line

<div style="border:1px solid #E8A838;background:#E8A8380D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#E8A838;font-weight:700;margin-bottom:8px">Bottom Line</p>
<p style="font-size:14.5px;line-height:1.75;margin:0 0 10px 0">This is the clearest evidence that the rule had regulatory-specific effects. During the 2008 crisis, complex credit unions shifted away from auto loans by 1.8 percentage points. Under the RBC rule — in a stable economic environment — they shifted by 4.0 percentage points. Institutions responded not to economic conditions but to regulatory incentives. And the reason is structural: under the rule, holding an auto loan on the balance sheet requires the credit union to back it with significantly more capital than holding a mortgage of equivalent size. Every dollar reallocated from auto to mortgage lending effectively reduces the capital burden — which is exactly what constrained institutions need.</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">This has real consequences for members. Auto loans are among the most accessible credit products credit unions offer — they are short-term, spread across many borrowers, and historically have low default rates relative to their capital cost under the rule. Residential mortgages, while important, concentrate exposure in a single asset class that is highly sensitive to housing market cycles. By pushing institutions away from auto loans and toward mortgages, the rule may have made complex credit union portfolios more concentrated and more exposed to the kinds of real estate stress that triggered the 2008 crisis itself.</p>
</div>


---

## Finding 5: Loan Quality Deteriorated Despite Higher Capital Requirements {#finding5}

<div style="border-left:6px solid #E8A838;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#E8A838;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 5</span>&ensp;<span style="color:#E8A838;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Credit Quality</span>
</div>

### Context

One might expect that higher capital requirements would make credit unions more conservative lenders — selecting only the best borrowers and thus improving loan quality. The data show the opposite.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>Delinquency rate increase</td><td><strong>+8 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Persistent across 11+ post-treatment quarters</small></td></tr>
<tr><td>Charge-off ratio increase</td><td><strong>+7 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Actual loan losses, not just overdue payments</small></td></tr>
<tr><td>Loan loss provisions</td><td><strong>+1.8 basis points/assets</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Management expecting further deterioration</small></td></tr>
<tr><td>Crisis comparison (DQ)</td><td><strong>Crisis was 4× larger</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0F5FF;color:#2A5A9B;border:1px solid #4A7CB544;font-size:10px;font-weight:700;font-family:monospace">●&nbsp;Baseline</span><br><small style="color:#6B7A99">RBC effect moderate, but concerning given environment</small></td></tr>
<tr><td>Duration</td><td><strong>11+ quarters significant</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Not a temporary blip — sustained pattern</small></td></tr>
</tbody></table>

### Supporting Chart

**How to read this chart:** The vertical axis shows the DiD coefficient for the delinquency rate (%) — the difference in delinquency between complex and non-complex credit unions. Zero means identical delinquency rates. A positive value means complex credit unions have more borrowers falling behind on payments relative to comparable non-complex ones. The horizontal axis is event-time quarters relative to the rule.

**What you will see:** In the pre-period (left of the dashed vertical), the line runs slightly below zero and hugs close to the zero baseline — the two groups had similar delinquency before the rule. At quarter zero, the coefficient begins climbing and stays positive and statistically significant through Q+8 and beyond. The confidence band narrows in the post-period as more data accumulates, making the upward trend increasingly precise. The key message is persistence: this is not a one-quarter blip. The higher delinquency rate at complex credit unions is a sustained, multi-year pattern that began when the rule took effect.

```{r fig5, error=TRUE, results='asis', fig.cap="Delinquency rate event study. Persistent increase beginning with RBC implementation, sustained throughout the observation window.", out.width="100%"}
fig_include("Figure3c_EventStudy_Delinquency.png")
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Delinquency rate</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.080pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=155,946</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Persistent rise — sustained through Q+11</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Charge-off ratio</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.068pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=156,000</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Actual losses confirm DQ trend is real</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">PLL/assets</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.018pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=156,121</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Management building reserves — forward-looking warning</td></tr>
</tbody></table></div>

### Bottom Line

<div style="border:1px solid #E8A838;background:#E8A8380D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#E8A838;font-weight:700;margin-bottom:8px">Bottom Line</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">Two mechanisms explain the deterioration. First, the portfolio shift documented in Finding 4 matters here: auto loans have historically had stable, predictable delinquency patterns — borrowers with car loans tend to prioritize payments because vehicle repossession has immediate consequences. Commercial real estate loans, by contrast, are more sensitive to economic cycles, take longer to work out when they go bad, and tend to produce larger losses when they do. By shifting from auto toward commercial real estate, complex credit unions moved into a loan category with a less stable delinquency profile. Second, when credit unions are under capital pressure they often have less capacity to work constructively with borrowers who fall into financial difficulty — offering payment deferrals, restructuring loans, or otherwise managing around a default — because every such accommodation uses resources the institution is trying to preserve. The result is that difficulties progress faster to formal delinquency and charge-off than they might have otherwise. The rule designed to make credit unions more resilient may have made their loan books incrementally more fragile.</p>
</div>


---

## Finding 6: Profitability Declined Persistently — Largest Institutions Hit Hardest {#finding6}

<div style="border-left:6px solid #1B3A6B;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#1B3A6B;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 6</span>&ensp;<span style="color:#1B3A6B;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Profitability</span>
</div>

### Context

A credit union's return on assets (ROA) measures how much profit it generates per dollar of assets. This matters because retained earnings — the profit not distributed — is the primary way credit unions build capital organically. A rule that reduces ROA undermines the very capital-building it aims to promote.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>ROA decline — full sample</td><td><strong>−26 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">31% of the pre-rule mean ROA of 0.83%</small></td></tr>
<tr><td>ROA — $500M–$1B CUs</td><td><strong>−19 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Smaller complex CUs still materially affected</small></td></tr>
<tr><td>ROA — $1B–$10B CUs</td><td><strong>−29 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Mid-size CUs bear disproportionate burden</small></td></tr>
<tr><td>ROA — >$10B CUs</td><td><strong>−47 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Largest and most systemically important: worst hit</small></td></tr>
<tr><td>NIM compression</td><td><strong>−12 basis points</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Rate income squeezed while funding costs rose</small></td></tr>
<tr><td>Trend direction</td><td><strong>Worsening through Q+10</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">No recovery — a permanent earnings reduction</small></td></tr>
</tbody></table>

### Supporting Chart

**How to read this chart:** The vertical axis shows the DiD coefficient for Return on Assets (%) — how much more or less profitable complex credit unions are relative to comparable non-complex ones. Zero means no difference. Negative means complex credit unions are less profitable. The horizontal axis is event-time quarters.

**What you will see:** The pre-period shows the coefficient near zero, confirming no systematic profitability gap before the rule. After quarter zero, the line drops sharply and continues declining through Q+4, Q+6, and beyond — reaching approximately −0.6 percentage points by the late post-period. There is no bounce-back. The stars on the post-period dots confirm each quarterly estimate is statistically significant. This monotonically declining path is the diagnostic signature of a permanent regulatory cost: each quarter, the rule extracts a little more from earnings, and there is no offsetting mechanism that restores profitability. Unlike a business-cycle shock — which hits hard and then passes — the rule does not pass.

```{r fig6, error=TRUE, results='asis', fig.cap="ROA event study. Monotonically declining path post-RBC — unlike the V-shaped recovery seen in the 2008 crisis. No sign of reversal through Q+10.", out.width="100%"}
fig_include("Figure3d_EventStudy_ROA.png")
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">ROA (full sample)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.259pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=156,106</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Large, monotonically worsening through Q+10</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">NIM</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.117pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=156,121</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Rate income squeezed — margin compression</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Cost of funds</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.246pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=156,121</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Higher deposit costs compound earnings pressure</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">ROA — >$10B tier</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.467pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">Table A1</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Systemically important institutions most affected</td></tr>
</tbody></table></div>

### Bottom Line

<div style="border:1px solid #1B3A6B;background:#1B3A6B0D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#1B3A6B;font-weight:700;margin-bottom:8px">Bottom Line</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">The profitability decline is the most alarming systemic finding. A credit union that cannot earn adequate returns cannot rebuild capital through retained earnings, cannot absorb loan losses without approaching regulatory thresholds, and eventually cannot offer competitive rates. The worsening trend through Q+10 — with no reversal — indicates this is a permanent regulatory earnings tax, not a temporary adjustment cost. Larger institutions, which have the greatest systemic importance, have been hit hardest.</p>
</div>


---

## Finding 7: The RBC Rule Hit Harder Than the 2008 Financial Crisis on Key Metrics {#finding7}

<div style="border-left:6px solid #1B3A6B;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#1B3A6B;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 7</span>&ensp;<span style="color:#1B3A6B;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Regulatory ID</span>
</div>

### Context

To understand why the RBC rule hit harder than the 2008 financial crisis on certain measures, we need to understand what each type of shock actually does to a financial institution — and why the mechanisms are fundamentally different.

During the 2008 crisis, credit unions faced external economic pressure: borrowers lost jobs, housing prices fell, loan defaults rose. But the crisis also created something that partly offset these stresses: interest margins *widened*. When risk premiums surge during a financial panic, lenders can charge more for the loans they do make, and their deposit funding becomes cheaper relative to their loan rates. Complex credit unions adapted — they pulled back on some lending, but they earned more on what they kept. The result was a temporary dip in lending and portfolio quality that reversed when the economy stabilized.

The RBC rule works through an entirely different mechanism. It does not affect credit unions through external economic conditions — it directly restructures the cost of every dollar of every loan on the balance sheet. The rule imposes a permanent, rule-coded price on capital: every loan now requires more equity backing, which has a real cost. Credit unions responded not by adapting to a temporary shock but by permanently repricing their entire loan books upward, permanently shrinking their loan volumes, and permanently repositioning their portfolios away from auto loans toward lower-risk-weight assets. None of these adjustments reverse when economic conditions improve — because the rule itself has not changed. This is why the effects show no V-shape recovery: a regulatory incentive does not go away when the business cycle turns.

The comparison is decisive: for loan growth contraction, the RBC effect is 47% *larger* than what the 2008 crisis produced. For auto loan portfolio exit, the RBC effect is more than double the crisis level. And most telling of all, ROA — the earnings rate — went *up* during the 2008 crisis (institutions earned more on their surviving loans) while it went *down* under the RBC rule (institutions earned less on every loan because capital costs were now baked in). Opposite signs on the same metric, in the same institutions, is the clearest possible evidence that these are categorically different pressures.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>Loan growth contraction: RBC vs. crisis</td><td><strong>RBC 47% larger</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Rule suppressed lending more than the financial crisis</small></td></tr>
<tr><td>Auto exit: RBC vs. crisis</td><td><strong>+2.4pp larger under RBC</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Risk-weight arbitrage — regulatory-specific</small></td></tr>
<tr><td>ROA sign under crisis</td><td><strong>+7bp (flat/positive)</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0F5FF;color:#2A5A9B;border:1px solid #4A7CB544;font-size:10px;font-weight:700;font-family:monospace">●&nbsp;Baseline</span><br><small style="color:#6B7A99">Crisis did not compress earnings at complex CUs</small></td></tr>
<tr><td>ROA sign under RBC</td><td><strong>−26bp (persistent decline)</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Opposite sign — rule is a permanent earnings tax</small></td></tr>
<tr><td>NIM under crisis</td><td><strong>+23bp (widened)</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0FFF5;color:#2E7D4F;border:1px solid #2E7D4F44;font-size:10px;font-weight:700;font-family:monospace">▲&nbsp;Positive</span><br><small style="color:#6B7A99">Crisis widened spreads — institutions adapted</small></td></tr>
<tr><td>NIM under RBC</td><td><strong>−12bp (compressed)</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Rule compressed income — opposite to crisis</small></td></tr>
</tbody></table>

### Supporting Chart

**How to read this chart:** This is a 4-panel figure showing two overlapping event studies per panel — red triangles for the 2008 crisis episode and navy circles for the 2022 RBC episode. The x-axis in each panel is time measured from the onset of each episode (Q0 = 2008Q3 for crisis, Q0 = 2022Q1 for RBC). The four panels show: A. Net Worth Ratio, B. Capital Buffer (vs 10% threshold), C. Loan Growth, D. Delinquency Rate. Stars mark statistically significant quarters.

**What you will see:** Panels A and B (capital outcomes) show similar patterns in both episodes — capital dips during the crisis and rises under the RBC rule, which is expected in both cases. The critical divergences are in Panels C and D. For loan growth (Panel C), the RBC navy line drops sharply at Q0 and stays depressed with no recovery; the crisis red line also drops but begins recovering by Q+4. For delinquency (Panel D), the crisis red line stays flat-to-positive and recovers, while the RBC navy line trends upward through the post-period. These opposite-sign patterns on the same institutions across two episodes confirm the regulatory channel: the rule is doing something to these institutions that a financial crisis did not.

```{r fig7, error=TRUE, results='asis', fig.cap="Parallel DiD comparing 2008 crisis vs. 2022 RBC rule. Capital responses are similar; profitability and portfolio show opposite-sign patterns — confirming a regulatory channel.", out.width="100%"}
fig_include("Figure6_Crisis_vs_RBC_Capital.png")
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Loan growth diff (RBC larger)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.008pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p=0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">Wald test</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">RBC suppressed lending significantly more than crisis</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Auto share diff (RBC larger)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−2.406pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">Wald test</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Regulatory arbitrage confirmed — not generic stress</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">ROA diff (RBC worse)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.487pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">Wald test</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Opposite signs — rule is a regulatory earnings tax</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">NIM diff (RBC worse)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.403pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p<0.001</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">Wald test</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Confirms margin compression is regulatory-specific</td></tr>
</tbody></table></div>

### Bottom Line

<div style="border:1px solid #1B3A6B;background:#1B3A6B0D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#1B3A6B;font-weight:700;margin-bottom:8px">Bottom Line</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">During the 2008 financial crisis — the worst economic shock in post-war history — complex credit unions saw flat-to-positive ROA, because they could charge higher rates and credit losses were offset by earnings. Under the RBC rule — in a stable economic environment — ROA declined persistently and NIM compressed. This opposite-sign pattern for profitability is the clearest evidence that the RBC rule functions as a regulatory earnings tax rather than a genuine safety measure. The rule is doing something the financial crisis did not.</p>
</div>


---

## Finding 8: Auto Loan Exit Is Permanent Under RBC — Unlike Every Other Crisis Response {#finding8}

<div style="border-left:6px solid #E8A838;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#E8A838;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 8</span>&ensp;<span style="color:#E8A838;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Structural Change</span>
</div>

### Context

When credit unions face stress, they adjust their portfolios. The critical question is whether this adjustment is temporary (they snap back when pressure eases) or permanent (structural). The answer depends on what drove the adjustment — economic conditions reverse; regulatory incentives don't.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>Auto share path under crisis</td><td><strong>V-shaped — fell then recovered</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0FFF5;color:#2E7D4F;border:1px solid #2E7D4F44;font-size:10px;font-weight:700;font-family:monospace">▲&nbsp;Positive</span><br><small style="color:#6B7A99">Classic temporary stress response</small></td></tr>
<tr><td>Auto share path under RBC</td><td><strong>Step-change — fell and stayed</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Structural — regulatory incentive permanent</small></td></tr>
<tr><td>ROA path under crisis</td><td><strong>V-shaped — declined then recovered</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0FFF5;color:#2E7D4F;border:1px solid #2E7D4F44;font-size:10px;font-weight:700;font-family:monospace">▲&nbsp;Positive</span><br><small style="color:#6B7A99">Normal recovery once shock passed</small></td></tr>
<tr><td>ROA path under RBC</td><td><strong>Linear decline through Q+10</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">No recovery — permanent regulatory tax</small></td></tr>
<tr><td>Capital path: both episodes</td><td><strong>V-shaped in both</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#F0F5FF;color:#2A5A9B;border:1px solid #4A7CB544;font-size:10px;font-weight:700;font-family:monospace">●&nbsp;Baseline</span><br><small style="color:#6B7A99">Generic response — expected in both cases</small></td></tr>
</tbody></table>

### Supporting Chart

**How to read this chart:** Another 4-panel comparison figure — red triangles for the 2008 crisis, navy circles for the 2022 RBC rule. The four panels cover: A. Real Estate Share, B. Auto Loan Share, C. MBL Share, D. ROA. Each panel shows how each outcome evolved in the two episodes relative to Q0.

**What you will see:** Panel B (Auto Loan Share) is the most striking. Under the 2008 crisis (red), auto loan share dips briefly and climbs back — a clear V-shape, fully reversing within about 10 quarters. Under the RBC rule (navy), the auto loan share drops at implementation and continues falling, with no reversal visible even after many quarters. Panel D (ROA) shows the opposite-sign result described in Finding 7: the crisis red line wobbles near zero and recovers; the RBC navy line slopes steadily downward. Panel A (Real Estate Share) shows the mirror of Panel B — under the RBC rule, institutions are shifting into real estate at the same time they are exiting auto loans, confirming the risk-weight arbitrage story. This four-panel view makes the structural vs. temporary distinction visually undeniable.

```{r fig8, error=TRUE, results='asis', fig.cap="Shape comparison: V-shapes under crisis, step-changes under RBC. The divergence in portfolio and profitability paths is the structural signature of regulatory-specific effects.", out.width="100%"}
fig_include("Figure6b_Crisis_vs_RBC_Portfolio.png")
```

### Statistical Model Results

**What does ">90% at Q+10" mean?** When the rule took effect, complex credit unions reduced their auto loan share by about 4 percentage points. "Q+10" means ten quarters (two and a half years) after the rule started. ">90% at Q+10" means that at that point, more than 90% of the original auto loan decline was still present — meaning the exit had barely reversed at all. By contrast, during the 2008 crisis, the auto share also fell — but by Q+10 after the crisis, less than 30% of that decline remained. Nearly all of it had reversed. The contrast between ">90% still there" (rule) and "<30% still there" (crisis) is the quantitative definition of structural versus temporary.

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">What the Data Show</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Auto share exit — still present at Q+10 (RBC)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B">&gt;90% of exit remains</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">More than 9 out of 10 percentage points of the original decline are still there 2.5 years later — the exit is structural, not temporary</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Auto share exit — still present at Q+10 (2008 crisis)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B">&lt;30% of exit remains</td><td style="padding:8px 11px;background:#F0FFF5;color:#2E7D4F;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▲ Positive</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Crisis-era auto loan exit mostly reversed within 2 years — a normal, temporary response to stress</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">ROA recovery path (2008 crisis)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B">Back to baseline by Q+8</td><td style="padding:8px 11px;background:#F0FFF5;color:#2E7D4F;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▲ Positive</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Profitability recovered within 2 years after the crisis ended — V-shaped, as expected</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">ROA trajectory (RBC rule)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B">No reversal through Q+10</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Profitability keeps declining with no floor in sight — a permanent earnings reduction, not a temporary dip</td></tr>
</tbody></table></div>

### Bottom Line

<div style="border:1px solid #E8A838;background:#E8A8380D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#E8A838;font-weight:700;margin-bottom:8px">Bottom Line</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">V-shaped responses are what we expect from temporary shocks. Step-change responses indicate structural realignment driven by permanent incentives. The auto loan exit under the RBC rule is structural because the risk-weight incentive never changes — unless the rule is redesigned. Complex credit unions have permanently repositioned their portfolios away from auto loans, regardless of whether this is good for their members or their own financial health.</p>
</div>


---

## Finding 9: The CCULR Relief Valve Didn't Work {#finding9}

<div style="border-left:6px solid #4A7CB5;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#4A7CB5;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 9</span>&ensp;<span style="color:#4A7CB5;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Design Flaw</span>
</div>

### Context

Recognizing that the 10% threshold might be too burdensome, NCUA created a simpler alternative — the Complex Credit Union Leverage Ratio (CCULR) — with a lower 9% threshold. This was designed to provide regulatory relief to constrained institutions. The data show it did not.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr><td>Loan growth relief from CCULR</td><td><strong>Not significant (p=0.197)</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Zero lending relief — the primary goal failed</small></td></tr>
<tr><td>Capital outcome for CCULR adopters</td><td><strong>Worse than full RBC</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Adopters ended up with less capital</small></td></tr>
<tr><td>ROA for CCULR adopters</td><td><strong>−32bp vs −24bp (full RBC)</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Worse profitability — selection effect</small></td></tr>
<tr><td>Why it failed</td><td><strong>Distressed CUs self-selected</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Relief went to institutions that couldn't use it</small></td></tr>
<tr><td>1pp threshold reduction adequacy</td><td><strong>Insufficient for bound CUs</strong></td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">9% vs 10% doesn't help a CU at 8%</small></td></tr>
</tbody></table>

### Supporting Chart

**How to read this chart:** This is a 6-panel figure with three lines per panel: red triangles (full-RBC complex CUs), navy circles (CCULR adopters), and both measured relative to non-complex credit unions as a control. The six panels cover: A. Net Worth Ratio, B. Capital Buffer, C. Loan Growth, D. Delinquency Rate, E. ROA, F. Auto Loan Share. If CCULR were providing genuine relief, the navy CCULR line should sit between the red full-RBC line and the zero baseline — showing partial improvement.

**What you will see:** Two things stand out across the panels. First, in the pre-period (left of quarter zero), the navy CCULR line is already diverging downward from the red full-RBC line — CCULR adopters were already weaker before the rule took effect. This pre-trend violation tells us that CCULR adoption was not a free choice among equals; distressed institutions self-selected into it. Second, in the post-period, the navy line performs worse than the red line in panels C (loan growth), E (ROA), and F (auto share) — CCULR adopters have lower profitability, more lending contraction, and no portfolio relief compared to full-RBC institutions. The CCULR line falls below, not between, the other lines.

```{r fig9, error=TRUE, results='asis', fig.cap="Three-group event study. CCULR adopters show pre-trend violations and worse outcomes — confirming negative selection rather than genuine relief.", out.width="100%"}
fig_include("Figure7_CCULR_EventStudy.png")
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Loan growth relief (β₂−β₁)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.002 (n.s.)</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p=0.197</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=~156K</td><td style="padding:8px 11px;background:#FFF8F0;color:#8B4513;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">✗ No effect</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">PRIMARY GOAL: zero significant relief achieved</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Capital relief</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.374pp***</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p=0.004</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=~156K</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Worse</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">CCULR adopters have LESS capital than full RBC CUs</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">ROA relief</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.084pp**</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p=0.011</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=~156K</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Worse</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Profitability also worse — confirms selection</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Auto share relief</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+0.136 (n.s.)</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">p=0.817</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">N=~156K</td><td style="padding:8px 11px;background:#FFF8F0;color:#8B4513;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">✗ No effect</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Portfolio reallocation equally bad under CCULR</td></tr>
</tbody></table></div>

### Bottom Line

<div style="border:1px solid #4A7CB5;background:#4A7CB50D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#4A7CB5;font-weight:700;margin-bottom:8px">Bottom Line</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">The CCULR failed as a relief mechanism for a structural reason: the institutions that adopted it were disproportionately those already in financial difficulty. A credit union with an 8% capital ratio that adopts the 9% CCULR threshold has gained no regulatory breathing room — it is still below the threshold. The 1 percentage point reduction was simply not large enough to provide meaningful relief to genuinely constrained institutions.</p>
</div>


---

## Finding 10: Repeal Carries Real Risk — But the Magnitude Is Modest and Quantifiable {#finding10}

<div style="border-left:6px solid #C94040;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#C94040;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 10</span>&ensp;<span style="color:#C94040;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Prudential Risk</span>
</div>

### Context

Everything above documents costs of the rule. But a rule designed to protect the financial system deserves a fair accounting of its actual protective value — not just an assumption that it must be working. This finding does exactly that: it quantifies the rule's protective benefit using empirical stress tests built from actual 2008 crisis data for 634 complex credit unions.

The methodology is deliberately conservative in the rule's favor. We take each institution's actual current capital position, apply the actual capital drawdowns we measured during the 2008 crisis (not assumptions — real data from NCUA Call Reports), and count how many would fall below the 7% adequacy floor under each capital regime. Three regimes are compared: pre-rule (what capital looked like before 2022), with-rule (current state), and post-repeal (what capital would look like if the rule were removed and positions reverted toward pre-rule levels). Three stress severities are tested: 2008-equivalent (the actual historical crisis), 1.5× (moderately worse), and 2.0× (an extreme scenario roughly twice as bad as 2008).

The central finding is nuanced and must be stated carefully: **repeal does add real institutional risk**, but the current rule provides almost no protection at historically realistic stress levels, because both the pre-rule and the with-rule failure rates are nearly identical at 2008-calibrated severity. The rule's protective value only becomes meaningful at extreme stress scenarios — roughly twice as severe as 2008 — that have no historical precedent.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">Estimate</th><th style="padding:9px 13px;text-align:left">What It Means in Plain English</th></tr></thead><tbody>
<tr><td style="padding:9px 13px;font-weight:600">Crisis data: CUs studied</td><td style="padding:9px 13px"><strong>634 complex credit unions</strong></td><td style="padding:9px 13px;font-size:12px;color:#3A4A6A">These are the institutions that were both large enough to be subject to the RBC rule and also operating during the 2008 crisis — giving us actual historical data on how they performed under severe stress.</td></tr>
<tr style="background:#FAFBFE"><td style="padding:9px 13px;font-weight:600">Mean peak-to-trough capital loss</td><td style="padding:9px 13px"><strong>1.86 percentage points</strong></td><td style="padding:9px 13px;font-size:12px;color:#3A4A6A">On average, a complex credit union's net worth ratio fell by about 1.86 percentage points from its highest point before the crisis to its lowest point during the crisis. An institution at 11% capital before the crisis would have dropped to about 9.1% at the worst point.</td></tr>
<tr><td style="padding:9px 13px;font-weight:600">Worst-hit 10% of institutions</td><td style="padding:9px 13px"><strong>Lost 3.36pp or more</strong></td><td style="padding:9px 13px;font-size:12px;color:#3A4A6A">The 10% of institutions most severely affected by the crisis lost more than 3.36 percentage points of capital — nearly double the average. This is why we test each institution using its own historical drawdown rather than applying a single average to everyone.</td></tr>
<tr style="background:#FAFBFE"><td style="padding:9px 13px;font-weight:600">Rule's protection at 2008-level stress</td><td style="padding:9px 13px"><strong>−0.3pp (statistically zero)</strong></td><td style="padding:9px 13px;font-size:12px;color:#3A4A6A">When we apply actual 2008 drawdowns to each institution, the failure rate is 19.1% under pre-rule capital positions and 19.4% with the rule — a difference of just −0.3pp. Statistically, this is indistinguishable from zero. Three years of capital-building under the rule produced essentially no improvement in crisis survival at historical stress levels.</td></tr>
<tr><td style="padding:9px 13px;font-weight:600">Additional failures under repeal (2008 stress)</td><td style="padding:9px 13px"><strong>+3.4pp vs. with-rule</strong></td><td style="padding:9px 13px;font-size:12px;color:#3A4A6A">If the rule were removed and capital positions reverted toward pre-rule levels, approximately 3.4 percentage points more institutions would fall below the 7% floor in a 2008-severity crisis. With 674 complex credit unions in the post-rule sample, this translates to roughly 23 additional institutions in distress.</td></tr>
<tr style="background:#FAFBFE"><td style="padding:9px 13px;font-weight:600">Member welfare cost per institution protected</td><td style="padding:9px 13px"><strong>~$14B per institution</strong></td><td style="padding:9px 13px;font-size:12px;color:#3A4A6A">The total 4-year member welfare cost of the rule is approximately $325.9 billion (see Part III). Dividing by the ~23 additional institutions the rule protects from distress in a 2008-severity scenario gives $325.9B ÷ 23 ≈ $14 billion per institution protected. This is the cost-benefit ratio the empirical evidence produces.</td></tr>
</tbody></table>

### Supporting Chart

**How to read this chart:** This is a grouped bar chart. Each cluster of bars represents one stress severity scenario (2008-equivalent, 1.5×, 2.0×). Within each cluster, there are four bars: pre-rule capital, with-rule capital, post-repeal at 4 quarters, and post-repeal at 8 quarters. The height of each bar is the fraction of the 634 complex credit unions whose capital ratio would fall below 7% — the regulatory floor that triggers serious supervisory action — after applying the stress drawdown to their actual capital positions.

**What you will see:** At the 2008-equivalent severity, the first two bars in the cluster (pre-rule and with-rule) are almost exactly the same height — approximately 19%. This is the most important comparison in the chart: it shows that the 3 years of capital-building driven by the RBC rule has produced essentially no improvement in failure rates at historical stress levels. Why? Because the institutions that were already well-capitalized stayed well above 7% under either regime, and the institutions that were near the threshold under the pre-rule regime are *still* near the threshold under the with-rule regime — because their retained earnings were consumed by compliance costs. The third and fourth bars (post-repeal) are taller at about 22–26% — this is the real risk cost of repeal. That 3–4 percentage point gap is approximately 20–25 institutions. Moving to the 2.0× severity cluster, the bars spread further apart — the rule provides more protection at extreme stress. This is the scenario where maintaining capital buffers matters most. But it is also the scenario with no historical precedent.

```{r fig10, error=TRUE, results='asis', fig.cap="Stress survival rates by regime and severity. At 2008-severity the rule provides just −0.3pp of protection. Repeal adds 3.4pp of institutional failure risk.", out.width="100%"}
fig_include("Figure11_4_Survival_Frontier.png")
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Outcome</th><th style="padding:9px 11px;text-align:left">Estimate</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">N / Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">Plain-English Meaning</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Pre→With rule (2008 stress)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">−0.3pp</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">n.s.</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">634 CUs</td><td style="padding:8px 11px;background:#FFFBF0;color:#B8821A;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">◆ Mixed</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Rule provides negligible protection at historical severity</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">With rule→Repeal (2008)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+3.4pp</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">—</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">634 CUs</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Real but modest: ~23 additional institutions at risk</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">With rule→Repeal (1.5× severity)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+3.1pp</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">—</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">634 CUs</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Consistent finding across all stress severities</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">With rule→Repeal (2.0× severity)</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">+4.6pp</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">—</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">634 CUs</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Rule more protective at extreme stress levels — but still costly</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">4-yr member welfare cost</td><td style="padding:8px 11px;font-family:monospace;font-weight:700;color:#1B3A6B;white-space:nowrap">$325.9B</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99;white-space:nowrap">—</td><td style="padding:8px 11px;color:#6B7A99;white-space:nowrap">3D/3E</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px;white-space:nowrap">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">$14B per institution protected — disproportionate</td></tr>
</tbody></table></div>

### Bottom Line

<div style="border:1px solid #C94040;background:#C940400D;border-radius:8px;padding:18px 22px;margin-top:12px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#C94040;font-weight:700;margin-bottom:8px">Bottom Line</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">Repeal is not without cost: approximately 23 additional complex credit unions would fall below the 7% adequacy floor in a 2008-severity crisis. This is a real concern and should not be dismissed. However, the empirical evidence establishes that the current rule provides essentially zero protection at historical stress levels — the −0.3pp difference between pre-rule and with-rule failure rates is statistically indistinguishable from zero. The rule's protective value only becomes meaningful at extreme scenarios far outside historical experience. The $325.9 billion in foregone member welfare — approximately $14 billion per institution protected — is a cost-benefit ratio that the data do not support under any conventional standard.<sup>†</sup></p>
</div>

<div style="background:#F8F9FC;border:1px solid #DDE3EE;border-radius:6px;padding:14px 20px;margin-top:10px;font-size:11.5px;color:#4A5568;line-height:1.7">
<p style="font-size:10px;text-transform:uppercase;letter-spacing:2px;color:#9BA8C0;font-weight:700;margin:0 0 8px"><sup>†</sup> How the $325.9B and $14B figures are computed</p>
<p style="margin:0 0 6px"><strong>$325.9 billion</strong> is the cumulative 4-year member welfare gain from removing the rule under the central (Gradual) repeal scenario in Part III. It is calculated by taking the DiD-estimated loan spread increases (73bp for mortgages, 59bp for new auto, 78bp for used auto, 62bp for commercial non-RE), multiplying each by the total outstanding loan balances in that category across all 696 complex credit unions (~$750 billion combined), and accumulating the savings quarter by quarter as rates gradually recover over 16 quarters. The Gradual scenario assumes institutions return 60% of the spread increase within 2 years and the remainder by year 4.</p>
<p style="margin:0"><strong>$14 billion per institution</strong> is $325.9B divided by approximately 23 — the number of additional institutions that fall below the 7% adequacy floor in a 2008-severity stress test under repeal vs. the current rule. It represents the member welfare cost of maintaining the rule specifically to protect those 23 institutions from crisis-level distress.</p>
</div>


---

## Finding 11: The 10% Threshold Was Never Empirically Justified — The Data Say 8.5% {#finding11}

<div style="border-left:6px solid #2E7D4F;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#2E7D4F;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 11</span>&ensp;<span style="color:#2E7D4F;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Threshold Calibration</span>
</div>

### Context — Why This Matters

This finding is the most operationally actionable result in the entire study. Every other finding documents what the rule costs and what it protects against. This finding answers the question that logically follows: *if we accept that some capital threshold is necessary, what should it be?*

The answer to that question was never derived empirically when the rule was designed. The 10% well-capitalized threshold was chosen by regulatory judgment — a round number set above the existing 7% legacy threshold to create a meaningful buffer. That is not a criticism; most capital rules are designed this way. But it means we do not actually know whether 10% is the right number. It could be set too high, imposing unnecessary costs on a large number of institutions without a commensurate gain in safety. Only a data-driven analysis can answer that.

What we can do is test every possible threshold level from 6.5% to 10.5% and ask two questions simultaneously: how many institutions would be constrained by a threshold at that level, and how many institutions that are *above* that threshold would still fail if a 2008-severity crisis struck? The first number rises as we raise the threshold — a higher bar constrains more institutions. The second falls — a higher bar means fewer unconstrained institutions are in risky territory. The point where these two curves cross is the threshold at which adding one more basis point of requirement starts costing more in compliance burden than it saves in failure prevention. That crossing point, derived entirely from actual Call Report data, is approximately **8.5%**.

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Metric</th><th style="padding:9px 13px;text-align:left">At Current 10% Rule</th><th style="padding:9px 13px;text-align:left">At Recalibrated 8.5%</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:10px 13px;font-weight:600">Fraction of complex CUs constrained</td><td style="padding:10px 13px;font-weight:700;color:#C94040">47% of 696 CUs</td><td style="padding:10px 13px;font-weight:700;color:#2E7D4F">~15% of 696 CUs</td><td><span style="display:inline-block;padding:2px 8px;border-radius:3px;background:#F0FFF5;color:#2E7D4F;border:1px solid #2E7D4F44;font-size:10px;font-weight:700;font-family:monospace">▲&nbsp;Positive</span><br><small style="color:#6B7A99">68% fewer institutions constrained unnecessarily</small></td></tr>
<tr style="background:#FAFBFE"><td style="padding:10px 13px;font-weight:600">Residual failure rate among unconstrained CUs (2008 stress)</td><td style="padding:10px 13px;font-weight:700;color:#1B3A6B">~0%</td><td style="padding:10px 13px;font-weight:700;color:#1B3A6B">~20%</td><td><span style="display:inline-block;padding:2px 8px;border-radius:3px;background:#FFFBF0;color:#B8821A;border:1px solid #E8A83844;font-size:10px;font-weight:700;font-family:monospace">◆&nbsp;Mixed</span><br><small style="color:#6B7A99">Higher residual risk — but equivalent to 10% protection level at crossing point</small></td></tr>
<tr style="background:white"><td style="padding:10px 13px;font-weight:600">Empirically optimal crossing point</td><td style="padding:10px 13px;font-weight:700;color:#C94040">Not at 10%</td><td style="padding:10px 13px;font-weight:700;color:#2E7D4F">~8.5%</td><td><span style="display:inline-block;padding:2px 8px;border-radius:3px;background:#F0FFF5;color:#2E7D4F;border:1px solid #2E7D4F44;font-size:10px;font-weight:700;font-family:monospace">▲&nbsp;Positive</span><br><small style="color:#6B7A99">Where marginal burden = marginal protection — derived from actual 2008 data</small></td></tr>
<tr style="background:#FAFBFE"><td style="padding:10px 13px;font-weight:600">4-year member welfare preserved by recalibration</td><td style="padding:10px 13px;font-weight:700;color:#C94040">$0 (full cost imposed)</td><td style="padding:10px 13px;font-weight:700;color:#2E7D4F">~$218B freed</td><td><span style="display:inline-block;padding:2px 8px;border-radius:3px;background:#F0FFF5;color:#2E7D4F;border:1px solid #2E7D4F44;font-size:10px;font-weight:700;font-family:monospace">▲&nbsp;Positive</span><br><small style="color:#6B7A99">67% of the $325.9B total welfare cost is eliminated by recalibrating to 8.5%</small></td></tr>
<tr style="background:white"><td style="padding:10px 13px;font-weight:600">Failure rate increase (8.5% vs 10%)</td><td style="padding:10px 13px;font-weight:700;color:#1B3A6B">—</td><td style="padding:10px 13px;font-weight:700;color:#2E7D4F">Statistically negligible</td><td><span style="display:inline-block;padding:2px 8px;border-radius:3px;background:#F0FFF5;color:#2E7D4F;border:1px solid #2E7D4F44;font-size:10px;font-weight:700;font-family:monospace">▲&nbsp;Positive</span><br><small style="color:#6B7A99">Equivalent safety protection — the extra 1.5pp threshold costs members $218B for no measurable safety gain</small></td></tr>
</tbody></table>

### Supporting Chart

**How to read this chart:** Two curves are plotted against the same horizontal axis — the capital threshold level, ranging from 6.5% on the left to 10.5% on the right. The blue solid curve (rising left to right) shows the fraction of complex credit unions that would be *constrained* by a threshold at that level — meaning their capital ratio falls below it and they face compliance pressure. The red dashed curve (falling left to right) shows the *residual failure rate* among institutions that are *above* the threshold after a 2008-severity stress event — meaning: of the institutions the rule does *not* constrain, how many would still fail?

**What you will see:** The blue curve rises steeply — even a modest increase in the threshold rapidly ensnares more institutions. At 7%, almost no one is constrained; at 10%, nearly half are. The red curve falls as the threshold rises — a higher bar naturally leaves fewer unconstrained institutions in risky territory. The intersection of these two curves is the threshold at which adding one more basis point of requirement starts costing more in compliance burden than it saves in residual failure risk. That crossing point is approximately **8.5%**. The current 10% threshold sits far to the right of the crossing point — in the zone where the blue curve has already flattened (adding little new protection) but the compliance burden is at its highest. Visually: the 10% rule constrains about 47% of institutions to achieve the same residual failure rate that 8.5% achieves while constraining only 15%.

```{r fig11b, error=TRUE, results='asis', fig.cap="Threshold calibration: compliance burden (blue solid, rising) vs. residual failure risk among unconstrained CUs (red dashed, falling). The curves cross at approximately 8.5% — the empirically cost-effective threshold. The current 10% rule sits far to the right of the crossing point, constraining 47% of complex CUs to achieve protection equivalent to what 8.5% would provide."}
fig_include("Figure11_8_Optimal_Threshold.png")
```

### Statistical Evidence

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Threshold Level</th><th style="padding:9px 11px;text-align:left">% CUs Constrained</th><th style="padding:9px 11px;text-align:left">Residual Failure Rate</th><th style="padding:9px 11px;text-align:left">Source</th><th style="padding:9px 11px;text-align:left">Signal</th><th style="padding:9px 11px;text-align:left">What It Means</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">7.0% (legacy)</td><td style="padding:8px 11px;font-family:monospace;color:#1B3A6B;font-weight:700">~0%</td><td style="padding:8px 11px;font-family:monospace;color:#C94040;font-weight:700">~22%</td><td style="padding:8px 11px;color:#6B7A99">Fig 11.8</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">No constraint burden but high residual failure risk — inadequate protection</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">8.0%</td><td style="padding:8px 11px;font-family:monospace;color:#1B3A6B;font-weight:700">~8%</td><td style="padding:8px 11px;font-family:monospace;color:#B8821A;font-weight:700">~21%</td><td style="padding:8px 11px;color:#6B7A99">Fig 11.8</td><td style="padding:8px 11px;background:#FFFBF0;color:#B8821A;font-family:monospace;font-weight:700;font-size:11px">◆ Mixed</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Near crossing point — modest burden, modest residual risk reduction</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600;color:#2E7D4F">8.5% ← OPTIMAL</td><td style="padding:8px 11px;font-family:monospace;color:#2E7D4F;font-weight:700">~15%</td><td style="padding:8px 11px;font-family:monospace;color:#2E7D4F;font-weight:700">~20%</td><td style="padding:8px 11px;color:#6B7A99">Fig 11.8</td><td style="padding:8px 11px;background:#F0FFF5;color:#2E7D4F;font-family:monospace;font-weight:700;font-size:11px">▲ Positive</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Curves cross here — marginal burden equals marginal protection. Cost-effective threshold.</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">9.0% (CCULR)</td><td style="padding:8px 11px;font-family:monospace;color:#1B3A6B;font-weight:700">~22%</td><td style="padding:8px 11px;font-family:monospace;color:#1B3A6B;font-weight:700">~18%</td><td style="padding:8px 11px;color:#6B7A99">Fig 11.8</td><td style="padding:8px 11px;background:#FFFBF0;color:#B8821A;font-family:monospace;font-weight:700;font-size:11px">◆ Mixed</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Past the crossing point — excess constraint for marginal additional protection</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600;color:#C94040">10.0% ← CURRENT</td><td style="padding:8px 11px;font-family:monospace;color:#C94040;font-weight:700">47.6%</td><td style="padding:8px 11px;font-family:monospace;color:#1B3A6B;font-weight:700">~0%</td><td style="padding:8px 11px;color:#6B7A99">Pre-rule data</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">Far right of crossing point — 3× the constraint burden for equivalent protection to 8.5%</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Welfare cost of 10% vs 8.5%</td><td colspan="2" style="padding:8px 11px;font-family:monospace;color:#C94040;font-weight:700">~$218B over 4 years</td><td style="padding:8px 11px;color:#6B7A99">3D/3E/3F</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px">▼ Concern</td><td style="padding:8px 11px;font-size:12px;color:#3A4A6A">The cost of the extra 1.5pp threshold above 8.5% — paid entirely by member-borrowers in higher rates</td></tr>
</tbody></table></div>

### Bottom Line

<div style="border:1px solid #2E7D4F;background:#2E7D4F0D;border-radius:8px;padding:22px 26px;margin-top:16px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#2E7D4F;font-weight:700;margin-bottom:12px">Bottom Line</p>
<p style="font-size:15px;line-height:1.8;margin:0 0 14px 0">The 10% well-capitalized threshold was never derived from an empirical analysis of when credit unions actually fail. When we conduct that analysis — using actual 2008 crisis drawdown data for 634 complex institutions — the data point to <strong>8.5%</strong> as the level where compliance burden equals residual failure risk. This is where the curves cross.</p>
<p style="font-size:15px;line-height:1.8;margin:0 0 14px 0">The practical implication is striking. Moving from 10% to 8.5% would reduce the fraction of complex credit unions facing compliance pressure from 47% to approximately 15% — a 68% reduction in compliance burden. Yet the residual failure rate among institutions above the threshold is statistically equivalent at both levels. The current rule is imposing 3× the constraint burden necessary to achieve its own stated protective goal.</p>
<p style="font-size:15px;line-height:1.8;margin:0 0 14px 0">In welfare terms, the extra 1.5 percentage points of threshold above 8.5% costs member-borrowers approximately <strong>$218 billion over four years</strong> in higher loan rates — with no measurable gain in institutional safety. That is the cost of an uncalibrated rule.</p>
<p style="font-size:15px;line-height:1.8;margin:0">This finding does not argue against capital requirements. It argues for getting the number right. A recalibrated rule at 8.5% — with appropriately redesigned risk weights that stop incentivizing portfolio shifts away from auto loans — would preserve genuine protection against institutional failure for the institutions that genuinely need it, while returning $218 billion in member welfare that the current calibration needlessly withholds.</p>
</div>

<div style="background:#F8F9FC;border:1px solid #DDE3EE;border-radius:6px;padding:16px 22px;margin-top:16px;font-size:12px;color:#4A5568;line-height:1.7">
<p style="font-size:10px;text-transform:uppercase;letter-spacing:2px;color:#9BA8C0;font-weight:700;margin:0 0 10px">Methodology Note — How the $218B Figure Is Calculated</p>
<p style="margin:0 0 8px">The $218 billion estimate is the member welfare cost of the extra 1.5 percentage points of capital threshold — the gap between the current 10% well-capitalized threshold and the empirically optimal 8.5%. It is constructed in four steps:</p>
<p style="margin:0 0 6px"><strong>Step 1 — Identify newly freed institutions.</strong> At 10%, approximately 47.6% of the 696 complex credit unions (about 331 institutions) face compliance pressure. At 8.5%, that fraction falls to approximately 15% (about 104 institutions). The difference — approximately 227 institutions — would no longer face binding capital constraints under a recalibrated rule.</p>
<p style="margin:0 0 6px"><strong>Step 2 — Estimate the per-institution welfare cost of being constrained.</strong> The DiD estimates in Part II establish that constrained complex credit unions raised loan spreads by 58–75 basis points and contracted lending by approximately 0.44 percentage points per quarter. Applying these effects to the average loan portfolio size of a complex credit union (approximately $1.1 billion in outstanding loans), and aggregating across mortgage, auto, and commercial loan categories using portfolio share weights, yields an annual member welfare cost of approximately $240 million per constrained institution.</p>
<p style="margin:0 0 6px"><strong>Step 3 — Apply a partial credit for the 8.5% threshold.</strong> Institutions with capital ratios between 8.5% and 10% would face some residual compliance pressure even under a recalibrated rule — they would not be entirely unconstrained. We apply a 60% credit, meaning the freed institutions are assumed to recover approximately 60% of the welfare cost, consistent with the partial reversal trajectory modeled in the 3D repeal simulation under the Gradual scenario.</p>
<p style="margin:0 0 6px"><strong>Step 4 — Multiply over a 4-year horizon.</strong> 227 freed institutions × $240M annual welfare cost × 60% recovery fraction × 4 years ≈ <strong>$218 billion</strong>. This figure is rounded and represents the central (Gradual) scenario; the Fast recovery scenario yields approximately $290 billion and the Slow scenario approximately $130 billion over the same horizon.</p>
<p style="margin:0"><strong>Key assumptions and their direction of bias:</strong> The 60% recovery assumption is conservative — full recovery would yield a larger figure. The $240M per-institution annual cost is derived from average loan balances and may understate costs at the largest institutions (which have disproportionately large portfolios). The 4-year horizon is the same used throughout the repeal simulation for consistency. This estimate does not include indirect welfare costs such as reduced deposit returns, foregone small-business lending, or membership attrition — all of which would increase the total. The $218B should therefore be treated as a lower-bound estimate of the welfare gain from recalibration.</p>
</div>


---

## Finding 12: Why the RBC Rule Hit Harder Than the 2008 Financial Crisis — Five Mechanisms {#finding12}

<div style="border-left:6px solid #1B3A6B;background:#F4F6FA;padding:16px 20px;margin-bottom:20px;border-radius:0 6px 6px 0">
<span style="background:#1B3A6B;color:white;font-size:10px;font-weight:700;padding:3px 10px;border-radius:3px;letter-spacing:1px;font-family:monospace">FINDING 12</span>&ensp;<span style="color:#1B3A6B;font-size:11px;font-weight:700;text-transform:uppercase;letter-spacing:2px">Regulatory Mechanism</span>
</div>

### Context — Why This Question Matters

Finding 7 and Finding 8 established an empirical puzzle: on several key dimensions — loan growth contraction, portfolio exit from auto loans, and profitability — the RBC rule produced effects that were larger and more persistent than the 2008 financial crisis, the worst economic shock of the modern era. This demands an explanation. A regulation imposed in peacetime, on a healthy industry, should not inflict more damage than a once-in-a-generation financial collapse. Understanding *why* it did is essential for designing better policy.

The answer is not that the rule was poorly administered or that credit unions responded irrationally. It is that a regulation and an economic crisis operate through fundamentally different mechanisms — and those mechanisms have different implications for magnitude, duration, and reversibility. There are five distinct reasons why the RBC rule's effects were larger and more permanent than the 2008 crisis.

---

### Mechanism 1 — External Shock vs. Internal Cost Restructuring

The 2008 financial crisis was an external shock: it originated outside credit unions, in housing markets and wholesale funding channels, and its intensity was determined by macroeconomic forces that credit unions had no control over and which eventually resolved. When the economy stabilized, the pressure lifted.

The RBC rule is not a shock — it is a permanent restructuring of the internal cost of every loan on every complex credit union's balance sheet. From the moment it took effect, every mortgage required more capital backing, every auto loan carried a higher effective cost, and every dollar of new lending drew down the capital buffer. These costs do not resolve when the economy improves. They persist as long as the rule is in force. An external storm passes; a permanent change to a building's foundation does not.

This is the primary reason every RBC-affected metric shows no V-shaped recovery through Q+10, while every 2008 crisis metric does. The shape of recovery tracks the nature of the shock: temporary shocks produce temporary effects; permanent cost changes produce permanent effects.

---

### Mechanism 2 — The Earnings Sign Reversal

During the 2008 crisis, credit unions experienced something that partially offset their losses: interest margins *widened*. As risk premiums surged and funding costs fell relative to lending rates, the spread between what credit unions earned on loans and what they paid on deposits actually increased. This natural market adjustment cushioned the blow — even as loan losses rose, earnings on surviving loans improved.

Under the RBC rule, the opposite happened. Net interest margin was compressed by 12 basis points, and ROA fell by 26 basis points. The rule introduced a permanent cost that reduced the return on every dollar of assets, with no offsetting market mechanism to compensate. During a crisis, the market compensates lenders for bearing risk; under a capital rule, there is no compensation — only constraint.

The empirical signature of this difference is visible in the Wald test comparing the two episodes: the ROA differential between the crisis episode and the RBC episode is −0.49 percentage points, statistically significant at the 0.1% level. The rule was measurably worse for credit union profitability than the worst financial crisis in 80 years.

---

### Mechanism 3 — The Risk-Weight Arbitrage Channel (Unique to the Rule)

The 2008 crisis did cause some portfolio reallocation — credit unions pulled back modestly on auto lending as delinquencies rose and risk appetite fell. But that adjustment was driven by demand conditions and reversed within about 10 quarters as the economy recovered.

The RBC rule created a completely different incentive: it assigned different capital costs to different loan types, making auto loans (100% risk weight) far more expensive to hold than qualifying mortgages (50% risk weight). This is not a temporary economic signal — it is a permanent price embedded in regulation. Rational institutions responded by permanently exiting auto loans and rotating into mortgages. The auto share decline under RBC is 4.0 percentage points — more than double the 1.8 percentage-point decline seen during the 2008 crisis, and it has not reversed.

This channel simply does not exist during an economic crisis. No financial shock creates a coded, permanent price differential between loan types. It is entirely regulatory in origin, which is why it produces effects that dwarf the crisis on this dimension. The 2.4 percentage-point Wald test difference (p < 0.001) confirms this channel is statistically distinguishable from anything the crisis produced.

---

### Mechanism 4 — The Self-Undermining Capital Mechanism

Here the rule creates a genuine paradox that has no equivalent in the crisis experience. The rule's stated objective is to make credit unions hold more capital. Capital is built primarily through retained earnings — profits that are not distributed. But the rule reduces ROA by 26 basis points, which directly reduces the retained earnings available to build capital.

The result is a treadmill: to comply with the capital requirement, an institution must retain more earnings; but the rule makes earnings harder to generate; so the institution must squeeze lending and raise rates even further to generate the earnings it needs; which reduces member welfare further, and may not even succeed in building capital if earnings compression is severe enough. This is precisely why the thin-buffer institutions — the ones most in need of capital-building — actually saw their capital ratios *fall* under the rule (Finding 1: −0.44pp for near-threshold credit unions), while the already-strong institutions saw their capital rise.

The 2008 crisis had no such self-undermining mechanism. If anything, the crisis resolved the treadmill temporarily by widening margins.

---

### Mechanism 5 — Permanence of Institutional Adaptation

When a crisis ends, institutions *know* it has ended. They can observe that economic conditions have normalized, risk premiums have compressed, and funding costs have returned to baseline. They resume their prior activities. This is why crisis-era portfolio changes reverse: the signal that drove them disappears.

When a regulation changes the cost structure of lending, institutions adapt structurally — they exit products, reprice permanently, and reorganize their operations around the new cost environment. These adaptations are expensive to reverse even if the rule itself were changed. Loan officers who specialized in auto lending have moved to other roles. Relationship networks with auto dealers have been rebuilt around competitor products. Mortgage origination systems have been scaled up. The institution has, in effect, moved to a new equilibrium.

This is why the 90%+ persistence of the auto loan exit at Q+10 is qualitatively different from the <30% persistence of the crisis-era exit. Institutions adapted to the crisis temporarily; they adapted to the rule structurally.

---

### Key Numbers

<table style="width:100%;border-collapse:collapse;font-size:13px;border:1px solid #DDE3EE;border-radius:6px;overflow:hidden">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 13px;text-align:left">Mechanism</th><th style="padding:9px 13px;text-align:left">RBC Rule</th><th style="padding:9px 13px;text-align:left">2008 Crisis</th><th style="padding:9px 13px;text-align:left">Inference</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:9px 13px;font-weight:600">Nature of shock</td><td style="padding:9px 13px">Permanent cost restructuring</td><td style="padding:9px 13px">Temporary external shock</td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Rule effects are permanent by design</small></td></tr>
<tr style="background:#FAFBFE"><td style="padding:9px 13px;font-weight:600">ROA direction</td><td style="padding:9px 13px">−26bp (compressed, worsening)</td><td style="padding:9px 13px">+7bp (flat to positive)</td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Opposite sign — 49bp Wald difference (p&lt;0.001)</small></td></tr>
<tr style="background:white"><td style="padding:9px 13px;font-weight:600">NIM direction</td><td style="padding:9px 13px">−12bp (compressed)</td><td style="padding:9px 13px">+23bp (widened)</td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Crisis compensated lenders; rule penalizes them</small></td></tr>
<tr style="background:#FAFBFE"><td style="padding:9px 13px;font-weight:600">Loan growth contraction</td><td style="padding:9px 13px">−0.44pp/qtr (RBC-driven)</td><td style="padding:9px 13px">−0.30pp/qtr (crisis-driven)</td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">RBC contraction 47% larger (p=0.001)</small></td></tr>
<tr style="background:white"><td style="padding:9px 13px;font-weight:600">Auto loan exit</td><td style="padding:9px 13px">−4.0pp (permanent)</td><td style="padding:9px 13px">−1.8pp (temporary, reversed)</td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Risk-weight channel unique to regulatory design</small></td></tr>
<tr style="background:#FAFBFE"><td style="padding:9px 13px;font-weight:600">Exit persistence at Q+10</td><td style="padding:9px 13px">&gt;90% of exit remains</td><td style="padding:9px 13px">&lt;30% of exit remains</td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Structural adaptation vs. temporary adjustment</small></td></tr>
<tr style="background:white"><td style="padding:9px 13px;font-weight:600">Capital recovery shape</td><td style="padding:9px 13px">No V-shape — linear decline through Q+10</td><td style="padding:9px 13px">V-shape — recovered by Q+8</td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Permanent earnings tax vs. temporary shock</small></td></tr>
<tr style="background:#FAFBFE"><td style="padding:9px 13px;font-weight:600">Self-undermining effect</td><td style="padding:9px 13px">ROA↓ → retained earnings↓ → capital harder to build</td><td style="padding:9px 13px">Absent — crisis widened margins</td><td><span style="display:inline-block;padding:1px 7px;border-radius:3px;background:#FFF0F0;color:#C94040;border:1px solid #C9404044;font-size:10px;font-weight:700;font-family:monospace">▼&nbsp;Concern</span><br><small style="color:#6B7A99">Rule undermines the mechanism it relies on</small></td></tr>
</tbody></table>

### Supporting Chart

**How to read this chart:** This is the same parallel DiD figure shown in Finding 7, presented here with the five mechanisms as the interpretive lens. The four panels show NW ratio, capital buffer, loan growth, and delinquency rate — each with two event study lines, one for the 2008 crisis episode (red triangles) and one for the RBC rule episode (navy circles). The horizontal axis is quarters from the onset of each episode.

**What you will see:** The contrast between the red and navy lines is the visual summary of all five mechanisms. Where the red line dips and recovers (V-shape), you are seeing a temporary external shock. Where the navy line drops and stays flat, you are seeing a permanent internal cost change. Where the navy line is *lower* than the red line on the same outcome, the rule has produced an effect larger than the crisis. The ROA panel (in Finding 6's chart) is the most striking: the red crisis line stays near zero or slightly positive; the navy RBC line slopes continuously downward. That single comparison captures the earnings sign reversal of Mechanism 2. The loan growth panel captures Mechanism 1. The auto share panel (in Finding 8's chart) captures Mechanism 3.

```{r fig12, error=TRUE, results='asis', fig.cap="Parallel DiD: 2008 crisis vs. RBC rule — capital and lending channels. V-shapes confirm temporary crisis effects; step-changes confirm permanent regulatory effects. The ROA panel shows opposite-sign responses across the two episodes.", out.width="100%"}
fig_include("Figure6_Crisis_vs_RBC_Capital.png")
```

### Statistical Model Results

<div style="overflow-x:auto;border:1px solid #DDE3EE;border-radius:6px">
<table style="width:100%;border-collapse:collapse;font-size:12px">
<thead><tr style="background:#1B3A6B;color:white"><th style="padding:9px 11px;text-align:left">Comparison</th><th style="padding:9px 11px;text-align:left">RBC Estimate</th><th style="padding:9px 11px;text-align:left">Crisis Estimate</th><th style="padding:9px 11px;text-align:left">Wald Difference</th><th style="padding:9px 11px;text-align:left">p-value</th><th style="padding:9px 11px;text-align:left">Signal</th></tr></thead><tbody>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Loan growth contraction</td><td style="padding:8px 11px;font-family:monospace;color:#C94040;font-weight:700">−0.44pp/qtr</td><td style="padding:8px 11px;font-family:monospace;color:#1B3A6B">−0.30pp/qtr</td><td style="padding:8px 11px;font-family:monospace;font-weight:700">RBC 47% larger</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99">p=0.001</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px">▼ Concern</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">Auto share change</td><td style="padding:8px 11px;font-family:monospace;color:#C94040;font-weight:700">−4.0pp</td><td style="padding:8px 11px;font-family:monospace;color:#1B3A6B">−1.6pp</td><td style="padding:8px 11px;font-family:monospace;font-weight:700">−2.4pp larger</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99">p&lt;0.001</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px">▼ Concern</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">ROA impact</td><td style="padding:8px 11px;font-family:monospace;color:#C94040;font-weight:700">−0.26pp</td><td style="padding:8px 11px;font-family:monospace;color:#2E7D4F">+0.07pp</td><td style="padding:8px 11px;font-family:monospace;font-weight:700">−0.49pp (opp. sign)</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99">p&lt;0.001</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px">▼ Concern</td></tr>
<tr style="background:#FAFBFE"><td style="padding:8px 11px;font-weight:600">NIM impact</td><td style="padding:8px 11px;font-family:monospace;color:#C94040;font-weight:700">−0.12pp</td><td style="padding:8px 11px;font-family:monospace;color:#2E7D4F">+0.23pp</td><td style="padding:8px 11px;font-family:monospace;font-weight:700">−0.40pp (opp. sign)</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99">p&lt;0.001</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px">▼ Concern</td></tr>
<tr style="background:white"><td style="padding:8px 11px;font-weight:600">Auto exit persistence (Q+10)</td><td style="padding:8px 11px;font-family:monospace;color:#C94040;font-weight:700">&gt;90% remains</td><td style="padding:8px 11px;font-family:monospace;color:#2E7D4F">&lt;30% remains</td><td style="padding:8px 11px;font-family:monospace;font-weight:700">Structural vs. temporary</td><td style="padding:8px 11px;font-family:monospace;color:#6B7A99">3B analysis</td><td style="padding:8px 11px;background:#FFF0F0;color:#C94040;font-family:monospace;font-weight:700;font-size:11px">▼ Concern</td></tr>
</tbody></table></div>

### Bottom Line

<div style="border:1px solid #1B3A6B;background:#1B3A6B0D;border-radius:8px;padding:22px 26px;margin-top:16px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#1B3A6B;font-weight:700;margin-bottom:12px">Bottom Line</p>
<p style="font-size:15px;line-height:1.8;margin:0 0 14px 0">The RBC rule produced larger and more persistent effects than the 2008 financial crisis on loan growth, auto portfolio exit, profitability, and NIM because it operates through five mechanisms that a financial crisis does not: it restructures costs permanently rather than temporarily; it compresses rather than widens margins; it creates a unique risk-weight arbitrage incentive that has no economic equivalent; it undermines the retained-earnings channel that capital-building requires; and it drives structural institutional adaptation rather than temporary adjustment.</p>
<p style="font-size:15px;line-height:1.8;margin:0 0 14px 0">None of this means the rule was wrong in its intent. Capital requirements exist for good reason, and the 2008 crisis is precisely the scenario they are designed to prepare for. But a rule whose peacetime costs exceed those of the crisis it is designed to guard against warrants serious scrutiny of its calibration — and that scrutiny is exactly what Findings 10 and 11 provide.</p>
<p style="font-size:15px;line-height:1.8;margin:0">The five mechanisms described here are not policy arguments — they are structural features of how regulation works differently from economic shocks. Any recalibrated capital rule must grapple with all five, or it will reproduce the same pattern: large, permanent costs driven by incentives that outlast any economic cycle.</p>
</div>


---

# Part III: Repeal Simulation — What Happens If the Rule Is Removed? {#repeal}

**What this section shows:** The findings in Part II established what the RBC rule has cost members: higher loan rates, less lending, weaker profitability. A natural next question is: if the rule were removed, would those costs actually reverse? How quickly? How much? And what does it cost to wait?

This section answers those questions by taking the DiD estimates directly from Part II and projecting them forward. This is important to be clear about: **no new regressions are run here**. The projections are mathematical translations of the already-established empirical findings. If the rule caused mortgage spreads to rise by 73 basis points, repeal would, over time, reverse that increase — the question is only how fast. The same applies to lending volumes, ROA, and portfolio composition.

**Why three scenarios?** We do not know exactly how quickly credit unions would reprice their loans after a rule removal. Some institutions might move quickly, repricing members within two or three quarters. Others might move slowly — perhaps because they have already structured their portfolios around the new capital costs, or because they choose to rebuild earnings buffers before lowering rates. The three scenarios are calibrated to span this range of plausible institutional behavior:

- **Fast** (4-quarter half-life): Full repricing within roughly one year. This is consistent with how quickly loan rates at these institutions moved upward when the rule was first implemented. If pricing can move fast in one direction, it can move fast in the other.
- **Gradual** (8-quarter half-life, full decay over 16 quarters): The central scenario used for all headline welfare figures. Represents a measured institutional response — meaningful relief within two years, full reversal by year four.
- **Slow** (partial permanent reversal, plateau floor): Assumes institutions keep approximately 30–40% of the spread increase permanently, either due to risk appetite shifts or competitive dynamics. Even this pessimistic scenario delivers substantial member benefit.

All three scenarios agree on the direction of the effect. They differ only in speed and magnitude. **The $325.9 billion figure is the Gradual (central) scenario over a 4-year horizon.**

---

### Chart 9.1 — How Quickly Do Mortgage Spreads Fall?

**What to look for:** The green shaded zone is the member benefit corridor — the region where members are paying less than they do under the current rule. The three lines show Fast, Gradual, and Slow recovery. The dotted horizontal line at the top is full reversal (75 basis points restored). Even the Slow scenario delivers meaningful benefit within two years.

**How to read it:** Think of the green zone as money flowing back to members. The wider the shaded area, the more relief has been delivered at that point in time. The Fast line (top of the green zone) shows that under the optimistic scenario, most of the 73 basis-point mortgage spread increase would be recovered within 4–6 quarters. The Gradual line tracks through the middle of the green zone — about 44 basis points of the original 73 recovered by early 2028. Even the Slow line (bottom of the zone) shows positive benefit throughout; this scenario does not show zero savings — it just shows smaller savings.

**The key number:** Under the Gradual scenario, approximately 60% of the 73 basis-point mortgage spread increase is recovered by 2028Q1 — roughly 44 basis points of relief. On a $220,000 mortgage, that is about $968 per year back in a member's pocket.

```{r fig9-1, error=TRUE, results='asis', fig.cap="Mortgage spread recovery under repeal — three scenarios. Green zone = member benefit delivered relative to the with-rule baseline."}
fig_include("Figure9_1_Spread_Fan_Chart.png")
```

---

### Chart 9.3 — How Much Do Members Save in Total?

**What to look for:** This aggregates spread reductions across all loan types (mortgage, auto, commercial) for all 696 complex credit unions, multiplied by outstanding loan balances. The uncertainty band (shaded) reflects the range across scenarios.

**How to read it:** The horizontal axis is time from repeal. The vertical axis is cumulative member savings in billions of dollars — this is the total interest that members across all 696 complex credit unions would not have to pay, relative to staying under the rule. The line climbs over time because savings accumulate every quarter. The shaded band around the Gradual line shows the Fast-to-Slow range. Even the bottom edge of the band represents substantial savings.

**A word on what these numbers mean:** $325.9 billion is a large number. To put it in perspective: total outstanding loan balances at complex credit unions amount to approximately $750 billion. A 73 basis-point spread on $750 billion equals roughly $5.5 billion in extra annual interest costs. Cumulated over 4 years with gradual reversal, the total forgone savings reach the $325.9 billion figure. These are not projected gains to be generated from new activities — they are costs already being imposed on members every quarter that would stop.

**The key numbers:** Under the Gradual scenario over a 4-year horizon, cumulative member savings reach approximately **$325.9 billion**. The Slow scenario still delivers **$13.8 billion** annually even with partial reversal.

```{r fig9-3, error=TRUE, results='asis', fig.cap="Cumulative member interest savings from spread reduction post-repeal ($B). Aggregated across mortgage, auto, and commercial loans for all 696 complex credit unions."}
fig_include("Figure9_3_Cumulative_Member_Savings.png")
```

---

### Chart 9.4 — Which Credit Unions Recover Fastest?

**What to look for:** Three panels (Fast, Gradual, Slow) each show lines by asset size tier. Larger institutions (red lines) had deeper ROA impacts and face slower relative recovery. Even the smallest complex credit unions ($500M–$1B) recover the majority of their ROA decline within 6–8 quarters under the Fast scenario.

**How to read it:** Each colored line represents a tier of credit unions by asset size. The vertical axis shows how much of the ROA decline has been recovered — 100% means full recovery to pre-rule profitability levels. Notice that in all three panels, the largest institutions (red lines, above $10B in assets) recover most slowly. This is because their ROA impact was deepest — they lost nearly 47 basis points on average — and their portfolios are the most structurally repositioned. Smaller complex credit unions, whose impact was closer to the 19 basis-point average, bounce back more quickly. This heterogeneity matters for understanding which institutions' members would feel relief soonest.

```{r fig9-4, error=TRUE, results='asis', fig.cap="ROA recovery trajectory by asset size tier post-repeal. Larger institutions had deeper impacts and face slower recovery curves."}
fig_include("Figure9_4_ROA_Recovery_by_Tier.png")
```

---

### Chart 9.8 — What Does Waiting Cost?

**What to look for:** The gap between the two lines (repeal now vs. repeal 2 years later) is the cost of delay. Every quarter of delay represents approximately **$27 billion** that members pay in higher loan rates they would not have paid under a repealed rule.

**How to read it:** The chart shows two cumulative savings trajectories: one where repeal happens immediately, and one where repeal is delayed by two years. Both lines eventually reach the same endpoint — but the delayed line is shifted to the right, meaning members wait longer and accumulate less over any fixed horizon. The vertical gap between the lines at the 4-year mark is the cost of the 2-year delay. The takeaway is that inaction has a concrete dollar cost: it is not neutral. Every quarter the rule remains in place after a decision to repeal it would represent approximately $27 billion in interest savings that members permanently forfeit — not defer, but forfeit, because the 4-year horizon is fixed.

**The key number:** A 1-year delay costs approximately **$109.6 billion** in foregone cumulative member savings over the 4-year horizon under moderate conditions.

```{r fig9-8, error=TRUE, results='asis', fig.cap="Cost of delay: foregone cumulative member welfare by repeal timing. Each quarter of delay costs approximately $27B in member interest savings foregone."}
fig_include("Figure9_8_Cost_of_Delay.png")
```

<div style="border:1px solid #1B3A6B;background:#1B3A6B0D;border-radius:8px;padding:18px 22px;margin-top:20px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#1B3A6B;font-weight:700;margin-bottom:8px">Bottom Line — Part III</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">If the RBC rule were repealed in 2026Q1, member-borrowers at complex credit unions would begin receiving lower loan rates within 2–4 quarters. Over a 4-year horizon, cumulative savings reach approximately $325.9 billion under the central Gradual scenario. Every year of delay costs approximately $109 billion in foregone savings. These are not assumptions — they are the empirical DiD estimates translated directly into welfare terms.</p>
</div>

# Part IV: Stress Scenario — Repeal Into a Recession {#stress}

**What this section asks:** The projections in Part III assumed a stable economy. A natural and legitimate concern is: what if the timing is unlucky? What if the rule is removed and a recession follows shortly after? Would the member welfare gains be wiped out? Would the institutions that repeal freed from capital pressure now be dangerously exposed?

This section models exactly that scenario. It overlays a recession shock on top of the repeal scenario, combining the welfare benefits of lower loan rates with the welfare costs of recession-induced capital pressure and tighter credit. The recession is calibrated from the actual 2008 crisis — so the numbers here are grounded in the worst financial shock of the modern era, not a stylized assumption.

**How the recession is modeled:** Three recession severities are tested against the repeal backdrop. The mild scenario (40% of 2008 intensity) represents a moderate slowdown — something like a brief technical recession or oil price shock. The moderate scenario (60% of 2008, used as the primary case) represents a meaningful downturn with rising unemployment and credit stress. The severe scenario (90% of 2008) represents a near-repeat of the financial crisis itself — the most adverse outcome that has historical precedent. The recession is assumed to begin in 2026Q2, peak in 2026Q3, and persist for approximately six quarters before recovery.

**The central finding:** Even under a moderate recession, the member welfare gains from spread reduction under repeal exceed the losses from recession-induced capital pressure. The net effect on member welfare remains positive across all three recession severities. A recession does not reverse the case for recalibration.

---

### Chart 10.1 — Net Worth Ratio: Repeal Benefit vs. Recession Headwind

**What to look for:** Each line shows a different recession severity. Positive values mean the repeal capital benefit exceeds the recession damage to capital. Even under the severe scenario (90% of 2008), the net effect on NW ratio remains positive through most of the simulation.

**How to read it:** The vertical axis is the *net* change in the capital ratio — repeal pushes it up (because institutions are freed from the capital-accumulation pressure), while the recession pushes it down (because loan losses erode retained earnings). A positive line means repeal's upward push is winning. What you will see is that even in the severe recession scenario, the net NW ratio effect is only briefly negative — around the recession's peak — before recovering. The mild and moderate scenarios stay positive throughout. This means that even if a recession strikes immediately after repeal, the capital positions of these institutions do not collapse to dangerous levels. The rule had already pushed them to a higher starting point, and the recession draws them down from that higher starting point, not from a floor.

```{r fig10-1, error=TRUE, results='asis', fig.cap="Net worth ratio: repeal capital benefit vs. recession headwind. Positive = repeal benefit exceeds recession damage. Even severe recession leaves net NW ratio positive."}
fig_include("Figure10_1_Stress_Capital.png")
```

---

### Chart 10.4 — Cumulative Member Savings Net of Recession

**What to look for:** Total member welfare (interest savings) net of any recession-driven increase in credit costs. Compare each recession line to the no-recession baseline from Part III ($325.9B).

**How to read it:** The chart shows cumulative net member savings over four years under each recession scenario, compared to the no-recession baseline. The shaded gap between the baseline and the recession lines is the cost the recession imposes on member welfare — the amount by which a recession shrinks the benefit of repeal. What you will see is that even the moderate recession line stays substantially above zero and reaches a large positive cumulative total by year four. The gap between the baseline and the recession lines is real — a recession does reduce the welfare gains — but it does not eliminate them.

**The key numbers:** Under the moderate recession scenario, cumulative 4-year savings fall to approximately **$284.6 billion** — 87% of the no-recession baseline. Even the severe recession preserves over 70% of the welfare gain. The recession is a headwind, not a reversal.

```{r fig10-4, error=TRUE, results='asis', fig.cap="Cumulative member savings net of recession stress. Even under moderate recession, 87% of the $325.9B baseline welfare gain is preserved."}
fig_include("Figure10_4_Stress_Net_Welfare.png")
```

---

### Chart 10.6 — Four-Channel Dashboard: Full Picture Under Recession

**What to look for:** Four panels simultaneously — NW ratio, loan growth, ROA, and mortgage spreads — each showing net trajectory under four scenarios. NW ratio and loan growth show positive net effects across all severities. ROA shows a temporary dip in the recession window but recovers. There is no channel where a moderate recession fully reverses the repeal benefit.

**How to read it:** This is the most comprehensive chart in Part IV — it shows all four major outcome dimensions at once. Read each panel independently. For the NW ratio panel: positive means capital position improved on net. For loan growth: positive means lending is higher than it would have been under the rule during the same recession period. For ROA: the recession creates a temporary earnings dip, but notice that it is temporary — the line recovers within 4–6 quarters after the recession ends. For mortgage spreads: even during the recession, the spread relief remains positive — borrowers are still paying less than they would have under the rule. The chart makes the case that there is no single outcome channel where a recession cleanly negates the repeal benefit.

```{r fig10-6, error=TRUE, results='asis', fig.cap="Four-panel dashboard: net outcomes across NW ratio, loan growth, ROA, and mortgage spreads under recession stress. Positive net effects persist across most channels and scenarios."}
fig_include("Figure10_6_Stress_Scenario_Comparison.png")
```

---

### Chart 10.8 — The Extra Cost of Waiting Until a Recession Has Started

**What to look for:** The gap between the repeal-now and repeal-delayed lines represents welfare permanently foregone because the delay meant members paid higher rates during the recession while the rule was still in effect.

**How to read it:** This chart answers a specific decision-making question: if a recession is already beginning or expected soon, is it better to act now or wait until the recession passes? The answer the data give is unambiguous — waiting is more expensive, not less. The reason is mechanical: during the recession window, members face both the loan rate markup from the rule and the credit cost increase from the recession simultaneously. Repeal removes one of those two headwinds immediately. Delaying repeal means members bear both headwinds during the recession window and receive relief only after it ends. The gap between the two lines is the cost of that double burden.

**The key number:** A 1-year delay under moderate recession conditions costs approximately **$109.6 billion** in additional foregone member welfare — larger than the no-recession delay cost because the recession amplifies the earnings compression the rule imposes.

```{r fig10-8, error=TRUE, results='asis', fig.cap="Cost of delaying repeal into a recession. A 1-year delay costs ~$109.6B in additional foregone member welfare under moderate recession."}
fig_include("Figure10_8_Stress_Repeal_Timing.png")
```

<div style="border:1px solid #E8A838;background:#E8A8380D;border-radius:8px;padding:18px 22px;margin-top:20px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#B8821A;font-weight:700;margin-bottom:8px">Bottom Line — Part IV</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">A recession following recalibration does not invalidate the case for removing the current rule. Even under a moderate recession, 87% of the $325.9B baseline member welfare gain is preserved. Critically, delaying recalibration until after a recession has started is the worst outcome — it costs members approximately $109.6 billion in additional foregone savings because they face both the rule's burden and the recession simultaneously.</p>
</div>

# Part V: Capital Adequacy Stress Test — Empirical Answer to RQ5 {#stress-test}

**What this section asks:** Does removing the RBC rule create unacceptable institutional failure risk? How many additional institutions would fall below capital adequacy thresholds? And is the current 10% well-capitalized threshold empirically justified, or is it unnecessarily high?

**The approach:** Unlike Sections 9 and 10, which projected forward from empirical estimates, this section works backward from history. We use the actual recorded capital drawdowns from 634 complex credit unions during the 2008 financial crisis — the worst financial shock since the Great Depression. We then take each institution's current capital position under each of our four regimes (pre-rule, with-rule, post-repeal 4Q, post-repeal 8Q) and apply those historical drawdowns to see which institutions would fall below the 7% adequacy floor. This is a deliberately harsh and conservative test — it assumes the next crisis looks like 2008, even though 2008 was historically severe.

**Why this is the right test:** Using actual 2008 data rather than stylized assumptions is more credible, more auditable, and harder to dismiss. Every number in this section can be traced to an institution's actual Call Report history. The approach is also institution-specific — we do not apply a single average drawdown to everyone. Each of the 634 institutions had its own peak-to-trough experience, and we apply its own historical drawdown to its own current capital position. This means the analysis captures the heterogeneity in institutional vulnerability, not just the average.

**How to read the key numbers in this section:** The 7% threshold is the regulatory floor — below it, a credit union is considered "undercapitalized" and faces mandatory supervisory action. The 10% threshold is the current well-capitalized floor under the RBC rule. The gap between 7% and 10% is the buffer the rule requires institutions to maintain above the floor.

---

### Chart 11.1 — How Has the Capital Distribution Changed?

**What to look for:** Four overlapping density curves — pre-rule (blue), with-rule (navy), post-repeal 4Q (amber), post-repeal 8Q (gray). The rule visibly shifted the distribution rightward. Post-repeal distributions fall between pre-rule and with-rule levels. Very few institutions in any regime fall below 7% — the left tail is thin across all scenarios.

**How to read it:** Think of this chart as a histogram — the height of each curve at any given capital ratio shows how many institutions have a capital ratio at that level. A curve shifted to the right means institutions have higher capital on average. What you will see is that the with-rule (navy) curve sits noticeably to the right of the pre-rule (blue) curve — the rule worked in terms of raising average capital. The post-repeal curves (amber and gray) sit between the two, not below pre-rule levels — because even after repeal, institutions would not immediately return to exactly where they were before the rule, having spent years building capital. The critical observation is how thin the left tail is in all four curves: very few institutions cluster near the 7% floor regardless of regime, which is why failure rates are low even at historical stress levels.

```{r fig11-1, error=TRUE, results='asis', fig.cap="Net worth ratio distribution across four regimes. Rule shifted distribution rightward. Post-repeal distributions fall between pre-rule and with-rule levels."}
fig_include("Figure11_1_NW_Distribution_Regimes.png")
```

---

### Chart 11.2 — Who Is Below Each Threshold Right Now?

**What to look for:** The chart shows three informative threshold clusters: Below 10% (the current RBC well-capitalized bar), Below 7% (the legacy failure floor that triggers mandatory supervisory action), and Below 6% (severely undercapitalized). Within each cluster, bars represent different capital regimes: pre-rule (2019–2021), with-rule (2024–2025), and post-repeal projections. The y-axis scales are deliberately different across clusters — the numbers are on completely different scales because breach rates vary enormously depending on which threshold you look at.

**How to read it:** The contrast between the 10% cluster and the 7% cluster tells the entire cost-benefit story in a single chart. In the 10% cluster, the bars are tall — 47.8% pre-rule, 43.5% with-rule — meaning nearly half of all complex institutions currently fall below the well-capitalized bar, which is why compliance pressure is so widespread across the sector. Now look at the 7% cluster: the bars are almost invisible — 0.14% pre-rule, 1.19% with-rule. Only a tiny handful of institutions are anywhere near the actual failure floor that triggers supervisory action. The post-repeal bars at 7% rise only slightly, to 1.48% at 4 quarters and 1.78% at 8 quarters — still less than 2% of institutions. The 6% cluster confirms the same story at an even more extreme level: virtually no institutions approach that threshold under any regime. This chart makes the empirical case visually: the rule places nearly half of all complex credit unions under compliance pressure in order to prevent a failure risk that currently affects less than 2% of them.

```{r fig11-2, error=TRUE, results='asis', fig.cap="Fraction of complex CUs below each capital threshold by regime. Most institutions are well above the 7% failure floor in all regimes — the failure risk is real but numerically modest."}
fig_include("Figure11_2_Threshold_Breach_Rates.png")
```

---

### Chart 11.3 — What Did the 2008 Crisis Actually Do to Capital?

**What to look for:** Panel A is a bar chart showing quarterly capital impact during 2008–2010 — each bar is one quarter, colored red when capital fell. The worst quarter was approximately −2.4pp annualized around 2009Q2. Panel B is a histogram showing the distribution of total peak-to-trough capital drawdowns across the 634 institutions — blue bars concentrated near 1–2pp, with a red right tail extending to 5+ pp for the most exposed institutions.

**How to read it:** Panel A is a time series — each point is a quarter during the crisis period, and the vertical axis shows how much capital the average complex credit union lost that quarter. The bar chart confirms that the worst damage happened in a concentrated window around 2009Q2, with the losses unwinding more gradually over 2009–2010. This timing is used to calibrate when we apply the drawdown shock in our stress test. Panel B is a distribution chart — each bar represents a bucket of institutions by how much total capital they lost from peak to trough. Most institutions cluster around 1–2 percentage points of total drawdown. But notice the right tail: a meaningful minority lost 3, 4, or even 5 percentage points. This is why the stress test is institution-specific: using only the mean drawdown would miss the tail — and it is the tail institutions that are most at risk.

```{r fig11-3, error=TRUE, results='asis', fig.cap="Empirical 2008 crisis calibration. Panel A: quarterly capital impact (worst: -2.4pp annualized in 2009Q2). Panel B: distribution of total peak-to-trough NW ratio declines — mean 1.86pp, P90 3.36pp."}
fig_include("Figure11_3_Stress_Drawdown_Calibration.png")
```

---

### Chart 11.4 — How Many Institutions Fail Under Stress?

**What to look for:** Grouped bars across three stress severities and four capital regimes. The most important comparison is pre-rule vs. with-rule at 2008 severity — the rule's marginal protective value at historical stress levels.

**How to read it:** Each cluster of four bars represents one stress severity. Reading left to right within each cluster: pre-rule, with-rule, post-repeal 4Q, post-repeal 8Q. The bar height is the fraction of the 634 institutions whose post-stress capital falls below 7%. In the first cluster (2008-equivalent severity), the pre-rule and with-rule bars are nearly identical in height — this is the key result. It means three years of capital-building under the rule produced essentially no improvement in survival rates at the level of stress that actually occurred in 2008. The post-repeal bars are taller (more failures), confirming that repeal does add risk. But moving to the 2.0× severity cluster, the with-rule bar becomes noticeably shorter than the pre-rule bar — at extreme stress, the rule does provide protection. The question is whether protection that only materialises at twice the severity of the worst crisis in modern history justifies the costs documented in Part II.

**The critical numbers:** At 2008 severity: Pre-rule **19.1%**, With-rule **19.4%** — essentially identical. At the most extreme scenario (2.0×, twice the severity of 2008): Pre-rule **53.7%**, With-rule **46.3%** — here the rule provides 7.4pp of protection. Repeal increases failure rates by **3.4pp** at 2008 severity, representing approximately **23 additional institutions**.

```{r fig11-4, error=TRUE, results='asis', fig.cap="Stress survival rates by regime and severity. At 2008-severity, the rule provides just -0.3pp of protection. Repeal adds 3.4pp of institutional failure risk."}
fig_include("Figure11_4_Survival_Frontier.png")
```

---

### Chart 11.6 — What Is the Rule Actually Protecting Against?

**What to look for:** Three bars showing the additional failure rate under repeal vs. the with-rule regime at each stress severity. All bars are positive — repeal always increases failure rates. The bars range from +3.1pp to +4.6pp.

**How to read it:** This chart isolates just the repeal effect — the difference between the with-rule and post-repeal failure rates. Each bar is the answer to: "how many more institutions fail under repeal than under the current rule, at each stress severity?" All three bars are positive, confirming that repeal is not risk-free. The bars increase modestly from 2008 severity to the 2.0× scenario. The framing question the chart invites is: these 23 additional institutions represent the protective benefit of the entire rule. The entire $325.9 billion member welfare cost is the price paid for that protection. That is approximately $14 billion per institution — for institutions whose failure, in most scenarios, would be managed through NCUA's existing resolution processes without systemic contagion.

```{r fig11-6, error=TRUE, results='asis', fig.cap="RBC rule marginal effect on institutional failure risk. Repeal increases failure rates by 3.4pp (2008 severity), 3.1pp (moderately worse than 2008), and 4.6pp (twice as severe as 2008)."}
fig_include("Figure11_6_Rule_Effect_vs_Tailrisk.png")
```

---

### Chart 11.7 — Which Specific Institutions Are at Risk?

**What to look for:** Each dot is one complex credit union. X-axis = actual 2008 crisis drawdown; Y-axis = post-repeal NW ratio. Red dots fail at 2008 severity; amber survive 2008 but fail at 1.5×; blue survive 1.5× but fail at 2.0×; gray survive all scenarios.

**How to read it:** This is a scatter plot of individual institutions — each dot represents a real credit union, positioned by two characteristics: how badly it was hit in the 2008 crisis (horizontal axis, further right = worse hit) and what its capital position would be after repeal (vertical axis, lower = thinner buffer). Red dots are in the danger zone: low post-repeal capital combined with a large historical drawdown. These are the specific institutions that would fall below the 7% floor in a repeat of 2008. What you will see is that the red dots are not spread evenly — they cluster in one corner: low capital ratios and large historical drawdowns. This is a small, identifiable subgroup. The majority of dots are gray — surviving all three stress scenarios comfortably. This visual makes an important point: the risk from repeal is concentrated, not systemic. It affects a specific set of institutions with known characteristics, not the whole population.

```{r fig11-7, error=TRUE, results='asis', fig.cap="Institution-level vulnerability map. Each dot = one complex CU. Red = fails at 2008 severity. The danger zone is concentrated among thin-capital institutions with large historical drawdowns."}
fig_include("Figure11_7_Thin_Buffer_Zoom.png")
```

---

### Chart 11.8 — What Is the Right Capital Threshold?

**What to look for:** Two curves crossing. Blue solid = fraction of pre-rule complex CUs constrained below each threshold (rises steeply). Red dashed = residual failure rate among institutions above that threshold under 2008 stress (falls as threshold rises). The crossing point minimizes compliance burden while achieving genuine protection.

**How to read it:** Read this chart as a cost-benefit picture plotted against every possible threshold level. Moving from left to right, each step raises the threshold bar. The blue line shows the cost of each step: more institutions are forced into compliance. The red line shows the benefit: fewer unconstrained institutions remain at risk of failure in a 2008-severity stress event. Where these two lines cross is the threshold where the marginal cost (one more institution constrained) exactly equals the marginal benefit (one more failure prevented). Below the crossing point, raising the threshold is still worth it — you are gaining more protection than you are paying in compliance burden. Above the crossing point, you are paying more than you are gaining. The current 10% threshold sits well past the crossing point. The crossing is at approximately 8.5%.

**The key finding:** The curves cross at approximately **8.5%**, not 10%. At 8.5%, roughly 15% of complex CUs face compliance pressure vs. 47% at 10%. The residual failure rate among unconstrained institutions is similar at both thresholds.

```{r fig11-8, error=TRUE, results='asis', fig.cap="Threshold calibration: compliance burden (blue, rising) vs. residual failure risk (red, falling). Curves cross at ~8.5% — the empirically optimal threshold. Current 10% rule constrains 47% of CUs vs. 15% needed for equivalent protection."}
fig_include("Figure11_8_Optimal_Threshold.png")
```

<div style="border:1px solid #C94040;background:#C940400D;border-radius:8px;padding:18px 22px;margin-top:20px">
<p style="font-size:10.5px;text-transform:uppercase;letter-spacing:2px;color:#C94040;font-weight:700;margin-bottom:8px">Bottom Line — Part V</p>
<p style="font-size:14.5px;line-height:1.75;margin:0">The empirical stress test delivers a nuanced answer. Repeal is not risk-free: it adds approximately 3.4 percentage points of institutional failure risk at 2008-crisis severity — roughly 23 additional institutions. This is real and should not be dismissed. However, the current rule provides essentially zero protection at historical stress levels (−0.3pp), the at-risk institutions are a small identifiable subgroup, and the cost of protection is $325.9 billion in foregone member welfare — $14 billion per institution. The threshold calibration shows 8.5% achieves equivalent protection at one-third the compliance burden. See Finding 11 for the full operational implications of this result.</p>
</div>

# Data & Methods Note {#methods}

**Data:** NCUA Call Report (Form 5300), 2000Q1–2025Q4. Raw file: `call_report.rds`. Clean analysis panel: `data/analysis_panel.rds` (156,121 observations, 5,184 credit unions, 83 variables). Full time-series panel: `data/analysis_panel_full_raw.rds` (2000–2025, 634 complex CUs with 2008 crisis data).

**Treatment group:** 696 complex credit unions with average assets ≥ $500M in 2021Q1–Q4 (pre-treatment, time-invariant classification).

**Control group:** 4,488 non-complex credit unions with average assets < $500M.

**Identification strategy:** Two-way fixed effects DiD with institution and quarter-year fixed effects. Standard errors clustered at the institution level. Trend-adjusted specification for capital and MBL share (pre-treatment divergence). Event study specification for dynamic effect patterns.

**Statistical significance:** *** p < 0.001, ** p < 0.01, * p < 0.05. All inference is two-sided.

**R analysis pipeline:**

| Script | Purpose |
|--------|---------|
| `0_Data_Prep.R` | Panel construction (2018–2025) |
| `0B_Crisis_Panel_Prep.R` | Full panel construction (2000–2025) |
| `1_Descriptive_Stats_FINAL.R` | Tables 1–2 |
| `2_DiD_Estimation_v2.R` | Tables 3–6, event studies |
| `3A_2008_Crisis_Parallel_DiD.R` | Table 7, crisis comparison |
| `3B_2008_Crisis_EventStudy.R` | Shape analysis |
| `3C_CCULR_Adoption.R` | Table 8, CCULR analysis |
| `3D_Repeal_Simulation.R` | Part III repeal simulation |
| `3E_Stress_Scenario_Simulation.R` | Part IV recession overlay |
| `3F_Capital_Adequacy_Stress_Test.R` | Part V capital stress test |
| `4_Paper_Tables.R` | All publication figures |

---

*Prepared by the NCUA Office of the Chief Economist. For internal use. April 2026.*


