---
title: "The 2022 NCUA Risk-Based Capital Rule"
subtitle: "Empirical Findings from an Analysis of Complex Credit Unions"
author: "Saurabh C. Datta, Ph.D. -- Senior Economist, NCUA Office of the Chief Economist"
date: "April 2026"
output:
  revealjs::revealjs_presentation:
    theme: white
    highlight: pygments
    center: false
    transition: fade
    slide_level: 2
    self_contained: true
    reveal_options:
      width: 1100
      height: 660
      margin: 0.04
      controls: true
      progress: true
      history: true
      keyboard: true
    css: rbc_slides.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)

fig_include <- function(path, caption = "") {
  if (file.exists(path)) {
    knitr::include_graphics(path)
  } else {
    fname <- basename(path)
    cat(paste0(
      "<div class=\"fig-ph\">",
      "<span class=\"fig-ph-label\">FIGURE: ", fname, "</span>",
      "</div>"
    ))
  }
}

fig_dir <- "output/paper/figures/"
```

## {data-background-color="#1a2744"}

<div style="padding:40px 50px;">
<p style="font-size:12px;font-weight:700;letter-spacing:0.18em;text-transform:uppercase;color:#c8972a;margin-bottom:18px;">NCUA &middot; Office of the Chief Economist</p>
<p style="font-family:Georgia,serif;font-size:38px;font-weight:700;color:#ffffff;line-height:1.2;margin-bottom:14px;">The 2022 NCUA Risk-Based Capital Rule</p>
<p style="font-family:Georgia,serif;font-size:22px;color:#e8c36a;margin-bottom:28px;line-height:1.3;">Empirical Findings from an Analysis of Complex Credit Unions</p>
<hr style="border:none;border-top:1px solid rgba(200,151,42,0.4);width:460px;margin:0 0 24px 0;">
<p style="font-size:14px;color:#8fa4c8;line-height:2.0;">Saurabh C. Datta, Ph.D. &middot; Senior Economist<br>NCUA Call Report Data, 2000--2025 &middot; April 2026</p>
</div>

## What We Studied -- and How

<table style="width:100%;border-collapse:collapse;font-size:13px;">
<tr>
<td style="width:52%;padding-right:20px;vertical-align:top;">

<p class="label-row">The Data</p>

- **696 complex credit unions** -- avg assets above $500M, classified using 2021 data
- **~4,488 non-complex credit unions** as the comparison group
- **150,000+ CU-quarters** of NCUA Call Report data, 2000--2025
- **12 outcomes**: capital, loan rates, volume, portfolio mix, profitability, loan quality

<hr class="slide-hr">

<p class="label-row">The Method: Difference-in-Differences (DiD)</p>

<table class="method-tbl">
<tr>
<td class="method-box"><strong>Change in<br>complex CUs</strong><br><span>Before vs. after rule</span></td>
<td class="method-op">&minus;</td>
<td class="method-box"><strong>Change in<br>non-complex CUs</strong><br><span>Same time window</span></td>
<td class="method-op">=</td>
<td class="method-result"><strong>The Rule's Effect</strong><br><span>What changed because of the rule, not the economy</span></td>
</tr>
</table>

<p class="callout"><strong>Why this works:</strong> Non-complex CUs faced the same economic conditions but were not subject to the RBC rule. Differences between the two groups isolate the rule's effect.</p>

</td>
<td style="width:2px;background:#c8c0b0;padding:0;"></td>
<td style="width:46%;padding-left:20px;vertical-align:top;">

<p class="label-row">How to Read the Event Study Charts</p>

<div class="intro-box">
<ul style="margin:0;padding:0;list-style:none;">
<li><strong>Horizontal axis:</strong> quarters relative to the rule's effective date (Q0). Left = before; right = after.</li>
<li><strong>Vertical axis:</strong> estimated difference between complex and non-complex CUs above any pre-existing gap.</li>
<li><strong>Before Q0:</strong> dots near zero confirm the two groups moved together pre-rule (parallel trends check).</li>
<li><strong>After Q0:</strong> a persistent move up or down is the rule's estimated effect. Shaded band = 95% confidence interval.</li>
<li><strong>Band crosses zero:</strong> effect not statistically distinguishable from no effect at that quarter.</li>
</ul>
</div>

```{r schematic, fig.height=2.2, fig.width=4.0, out.width="100%"}
library(ggplot2)
library(tibble)
q <- c(-4,-2,0,2,4,6)
e <- c(0.04,-0.02,0.00,-0.44,-0.57,-0.61)
s <- c(0.10, 0.10,0.00, 0.10, 0.10, 0.10)
d <- tibble(q=q,e=e,lo=e-1.96*s,hi=e+1.96*s)
ggplot(d,aes(x=q,y=e))+
  geom_hline(yintercept=0,linetype="dashed",color="#aaa",linewidth=0.5)+
  geom_vline(xintercept=0,linetype="dashed",color="#c8972a",linewidth=0.8)+
  geom_ribbon(aes(ymin=lo,ymax=hi),fill="#1a2744",alpha=0.15)+
  geom_line(color="#1a2744",linewidth=1.2)+
  geom_point(color="#1a2744",size=2.5,fill="white",shape=21,stroke=1.5)+
  annotate("text",x=0.2,y=0.20,label="Rule effective",hjust=0,color="#c8972a",size=2.6)+
  annotate("text",x=-3.8,y=0.20,label="Flat pre-period",hjust=0,color="#555",size=2.5)+
  annotate("text",x=2.0,y=-0.35,label="Persistent effect",hjust=0,color="#1a2744",size=2.6)+
  scale_x_continuous(breaks=q,labels=paste0(ifelse(q>0,"+",""),q,"Q"))+
  labs(x=NULL,y=NULL,caption="Schematic only -- actual charts on following slides")+
  theme_minimal(base_size=9)+
  theme(panel.grid.minor=element_blank(),
        panel.grid.major=element_line(color="#ece8e0",linewidth=0.4),
        plot.caption=element_text(color="#aaa",size=7,hjust=0.5),
        plot.background=element_rect(fill="white",color="#c8c0b0",linewidth=0.5),
        axis.text=element_text(color="#666"))
```

</td>
</tr>
</table>

## RQ 1 &nbsp;&middot;&nbsp; Pre-Rule Capital Buffer

<p class="rq-badge">Research Question 1</p>

**How much capital buffer did complex credit unions hold before the rule?**

<div class="answer-bar">Answer: The average complex credit union entered the rule with a buffer of just <strong>72 basis points</strong> above the new 10% threshold -- critically thin.</div>

<table style="width:100%;border-collapse:collapse;font-size:13px;margin-top:10px;">
<tr>
<td style="width:54%;padding-right:20px;vertical-align:top;">

<p class="label-row">What the Data Show</p>

- The 696 complex CUs were nominally compliant before the rule -- but with very little margin to absorb a new, higher threshold
- A **72 basis point buffer** means holding 10.72% net worth when the definition of well-capitalized shifted upward simultaneously
- Capital distribution was **tightly clustered just above the old threshold** -- a large mass barely on the compliant side
- This thin margin directly explains why the rule had such large downstream effects

</td>
<td style="width:2px;background:#c8c0b0;padding:0;"></td>
<td style="width:44%;padding-left:20px;vertical-align:top;">

<p class="callout"><strong>What is a basis point?</strong> One bp = one one-hundredth of a percentage point. A 72bp buffer means 10.72% capital against a 10.00% threshold.</p>

<table class="stat-tbl" style="margin-top:12px;">
<tr>
<td class="stat-cell"><div class="stat-num">72bp</div><div class="stat-desc">Average pre-rule capital buffer above the 10% threshold</div></td>
<td class="stat-cell stat-red"><div class="stat-num">696</div><div class="stat-desc">Complex CUs subject to the rule (fixed 2021 classification)</div></td>
</tr>
</table>

<p class="callout" style="margin-top:10px;"><strong>Why it matters:</strong> A rule requiring additional capital is always more costly when institutions start with little room to spare. The thin buffer set the stage for the pressures in RQ2--RQ4.</p>

</td>
</tr>
</table>

## RQ 2 &nbsp;&middot;&nbsp; Did Capital Levels Rise?

<p class="rq-badge">Research Question 2</p>

**Did the rule cause complex credit unions to hold more capital?**

<div class="answer-bar">Answer: <strong>Mixed.</strong> Average capital rose -- but fewer institutions are now meeting the well-capitalized standard. The bar rose faster than capital could accumulate.</div>

<table style="width:100%;border-collapse:collapse;font-size:13px;margin-top:10px;">
<tr>
<td style="width:54%;padding-right:20px;vertical-align:top;">

<table class="stat-tbl" style="margin-bottom:14px;">
<tr>
<td class="stat-cell stat-green"><div class="stat-num">+46bp</div><div class="stat-desc">Net worth ratio increase (p &lt; 0.01)</div></td>
<td class="stat-cell stat-red"><div class="stat-num">-3.7pp</div><div class="stat-desc">Change in well-capitalized probability (p &lt; 0.01)</div></td>
</tr>
</table>

<p class="label-row">The Paradox Explained</p>

- Capital rose -- institutions did respond to the rule's incentive
- But the **definition of well-capitalized changed** simultaneously, raising the bar
- Result: more capital on average, but **fewer institutions clearing the new bar**

<p class="callout" style="margin-top:8px;"><strong>Analogy:</strong> Every student studied harder and raised their score -- but the passing grade was raised even more. Average scores up, more students technically failing. Both facts are true simultaneously.</p>

</td>
<td style="width:2px;background:#c8c0b0;padding:0;"></td>
<td style="width:44%;padding-left:20px;vertical-align:top;">

<p class="label-row">Event Study -- Capital</p>

```{r fig-rq2, out.width="100%", results="asis"}
fig_include(file.path(fig_dir,"Figure3_EventStudy_Capital.png"),
            "NW ratio, well-cap prob, tier 1, risk-based capital -- 4-panel")
```

<p class="fig-note">Pre-period flat confirms parallel trends. Post-period divergence attributed to the rule.</p>

</td>
</tr>
</table>

## RQ 3 &nbsp;&middot;&nbsp; Member Cost: Higher Rates and Less Lending

<p class="rq-badge">Research Question 3</p>

**What did the rule cost members in higher loan rates and reduced lending?**

<div class="answer-bar">Answer: Loan rates rose across every product; lending contracted persistently. Estimated member cost: <strong>$325.9 billion over 4 years</strong> -- approximately <strong>$865 per member per year</strong>.</div>

<table style="width:100%;border-collapse:collapse;font-size:13px;margin-top:10px;">
<tr>
<td style="width:54%;padding-right:20px;vertical-align:top;">

<p class="label-row">Loan Rate Increases vs. Non-Complex CUs</p>

<table class="stat-tbl" style="margin-bottom:10px;">
<tr>
<td class="stat-cell stat-red"><div class="stat-num" style="font-size:20px;">+75bp</div><div class="stat-desc">Mortgage spread</div></td>
<td class="stat-cell stat-red"><div class="stat-num" style="font-size:20px;">+59bp</div><div class="stat-desc">Auto loan spread</div></td>
<td class="stat-cell stat-red"><div class="stat-num" style="font-size:20px;">+58bp</div><div class="stat-desc">Commercial spread</div></td>
</tr>
</table>

- All increases **statistically significant**, persist 10+ quarters with no convergence
- A 75bp mortgage increase on $300,000 adds roughly **$135 per month**

```{r fig-rq3a, out.width="100%", results="asis"}
fig_include(file.path(fig_dir,"Figure4_EventStudy_Spreads.png"),"Loan spread event study")
```

</td>
<td style="width:2px;background:#c8c0b0;padding:0;"></td>
<td style="width:44%;padding-left:20px;vertical-align:top;">

<p class="label-row">Lending Volume and Total Cost</p>

<table class="stat-tbl" style="margin-bottom:10px;">
<tr>
<td class="stat-cell stat-red"><div class="stat-num" style="font-size:18px;">-0.44pp/qtr</div><div class="stat-desc">Loan growth reduction -- persistent, no recovery</div></td>
<td class="stat-cell stat-red"><div class="stat-num" style="font-size:20px;">$325.9B</div><div class="stat-desc">Total member welfare cost over 4 years</div></td>
</tr>
</table>

- Loan growth fell **0.44pp per quarter** -- compounding into substantial credit reduction
- $325.9BN aggregates the rate effect across all 696 complex CU member-borrowers over 4 years
- Per-member: approximately **$865/year** in higher borrowing costs

```{r fig-rq3b, out.width="100%", results="asis"}
fig_include(file.path(fig_dir,"Figure3b_EventStudy_LoanGrowth.png"),"Loan growth event study")
```

</td>
</tr>
</table>

## RQ 4 &nbsp;&middot;&nbsp; Portfolio Shift and the Crisis Comparison

<p class="rq-badge">Research Question 4</p>

**Did the rule change what loans credit unions make -- and how does this compare to past stress?**

<div class="answer-bar">Answer: Yes. Complex CUs shifted from auto loans toward mortgages -- consistent with <strong>risk-weight arbitrage</strong>. This shift is unique to the RBC rule; it did not occur during the 2008 financial crisis.</div>

<table style="width:100%;border-collapse:collapse;font-size:13px;margin-top:10px;">
<tr>
<td style="width:54%;padding-right:20px;vertical-align:top;">

<table class="stat-tbl" style="margin-bottom:10px;">
<tr>
<td class="stat-cell stat-red"><div class="stat-num">-4.0pp</div><div class="stat-desc">Auto loan share (vs. controls)</div></td>
<td class="stat-cell stat-green"><div class="stat-num">+2.4pp</div><div class="stat-desc">Real estate loan share</div></td>
</tr>
</table>

- **Risk-weight arbitrage:** mortgages carry lower risk weights than auto loans under RBC. Shifting to mortgages reduces required capital -- a rational regulatory response
- Loan quality deteriorated: **delinquency +8bp**, driven by a capacity channel -- CUs with the largest profitability declines showed 2x the delinquency increase

```{r fig-rq4a, out.width="100%", results="asis"}
fig_include(file.path(fig_dir,"Figure5_Portfolio_Trends.png"),"Portfolio reallocation")
```

</td>
<td style="width:2px;background:#c8c0b0;padding:0;"></td>
<td style="width:44%;padding-left:20px;vertical-align:top;">

<p class="label-row">Comparison to the 2008 Financial Crisis</p>

<p class="callout">We ran the same DiD using 2008 as the event, comparing complex CU responses to the financial crisis vs. the RBC rule.</p>

<table style="width:100%;border-collapse:collapse;font-size:12px;margin-top:8px;">
<thead><tr style="background:#1a2744;color:#fff;">
<th style="padding:6px 8px;text-align:left;">Outcome</th>
<th style="padding:6px 8px;text-align:center;">2008 Crisis</th>
<th style="padding:6px 8px;text-align:center;">RBC Rule</th>
</tr></thead>
<tbody>
<tr style="border-bottom:1px solid #ddd;"><td style="padding:6px 8px;">Capital accumulation</td><td style="text-align:center;">Similar</td><td style="text-align:center;">Similar</td></tr>
<tr style="background:#f5f1eb;border-bottom:1px solid #ddd;"><td style="padding:6px 8px;">Auto loan exit</td><td style="text-align:center;color:#666;">Not observed</td><td style="text-align:center;color:#9b2226;font-weight:600;">Persistent</td></tr>
<tr style="border-bottom:1px solid #ddd;"><td style="padding:6px 8px;">Loan growth</td><td style="text-align:center;color:#666;">Temporary</td><td style="text-align:center;color:#9b2226;font-weight:600;">No recovery</td></tr>
<tr style="background:#f5f1eb;"><td style="padding:6px 8px;">Profitability</td><td style="text-align:center;color:#666;">Temporary</td><td style="text-align:center;color:#9b2226;font-weight:600;">No recovery</td></tr>
</tbody>
</table>

<p style="font-size:12px;margin-top:8px;color:#444;">Capital behavior resembles a crisis response -- but lending and profitability effects are <strong>distinctly regulatory</strong> in nature.</p>

</td>
</tr>
</table>

## Supporting Finding &nbsp;&middot;&nbsp; Profitability

<p class="intro-badge">Supporting Finding</p>

**Profitability Declined Persistently -- Largest Institutions Hit Hardest**

<div class="answer-bar">Return on assets fell across complex CUs -- with the largest institutions hit hardest. This directly informs the financial stability implications in RQ5.</div>

<table style="width:100%;border-collapse:collapse;font-size:13px;margin-top:10px;">
<tr>
<td style="width:54%;padding-right:20px;vertical-align:top;">

<table class="stat-tbl" style="margin-bottom:12px;">
<tr>
<td class="stat-cell stat-red"><div class="stat-num">-26bp</div><div class="stat-desc">Average ROA decline -- all complex CUs (p &lt; 0.01)</div></td>
<td class="stat-cell stat-red"><div class="stat-num">-47bp</div><div class="stat-desc">ROA decline for CUs above $10B in assets</div></td>
</tr>
</table>

- ROA decline is **persistent** -- no evidence of adaptation or recovery through the observation window
- Largest CUs (above $10B) experience a decline **nearly double the average**
- CCULR provided **no significant relief** on lending outcomes; CCULR adopters showed worse baseline outcomes consistent with distressed self-selection

</td>
<td style="width:2px;background:#c8c0b0;padding:0;"></td>
<td style="width:44%;padding-left:20px;vertical-align:top;">

<p class="callout"><strong>Lower profitability reduces organic capital-building ability.</strong> A CU that cannot retain earnings must raise rates, cut lending, or draw down its capital ratio -- all three patterns appear in the data.</p>

<p class="callout" style="margin-top:8px;"><strong>CCULR finding:</strong> The Community Credit Union Leverage Ratio was designed as a simpler compliance pathway. Institutions most likely to adopt it were already under stress -- CCULR adoption did not cause their worse outcomes, but did not reverse them either.</p>

```{r fig-roa, out.width="100%", results="asis"}
fig_include(file.path(fig_dir,"Figure3d_EventStudy_ROA.png"),"ROA event study")
```

</td>
</tr>
</table>

## RQ 5 &nbsp;&middot;&nbsp; Failure Risk from Removing the Rule

<p class="rq-badge">Research Question 5</p>

**Does removing the RBC rule create unacceptable failure risk?**

<div class="answer-bar">Answer: Removing the rule carries <strong>real but quantifiable and modest risk</strong>. At 2008 GFC severity, the rule reduces failure risk by 3.4 percentage points -- protecting approximately 24 institutions.</div>

<table style="width:100%;border-collapse:collapse;font-size:13px;margin-top:10px;">
<tr>
<td style="width:54%;padding-right:20px;vertical-align:top;">

<p class="label-row">Capital Adequacy Stress Test -- At 2008 GFC Severity</p>

<table class="stat-tbl" style="margin-bottom:12px;">
<tr>
<td class="stat-cell"><div class="stat-num" style="font-size:18px;">19.1%</div><div class="stat-desc">Pre-rule failure rate (baseline)</div></td>
<td class="stat-cell stat-green"><div class="stat-num" style="font-size:18px;">19.4%</div><div class="stat-desc">With-rule rate (-0.3pp incremental protection)</div></td>
<td class="stat-cell stat-red"><div class="stat-num" style="font-size:18px;">22.5%</div><div class="stat-desc">Post-removal rate (+3.4pp, ~24 institutions)</div></td>
</tr>
</table>

- Rule's **incremental protection is only -0.3pp** -- complex CUs were already building buffers before the rule
- Removal raises simulated failure risk by **+3.4pp**, corresponding to **~24 institutions**
- Each protected institution represents approximately **$14 billion in assets**

</td>
<td style="width:2px;background:#c8c0b0;padding:0;"></td>
<td style="width:44%;padding-left:20px;vertical-align:top;">

<table class="stat-tbl" style="margin-bottom:12px;">
<tr>
<td class="stat-cell"><div class="stat-num" style="font-size:16px;">$0.4B--$2.2B</div><div class="stat-desc">Probability-weighted NCUSIF cost of removal</div></td>
</tr>
</table>

<p class="label-row">Stress Scenario Sensitivity</p>

<table style="width:100%;border-collapse:collapse;font-size:12px;margin-bottom:10px;">
<thead><tr style="background:#1a2744;color:#fff;">
<th style="padding:6px 8px;text-align:left;">Recession Severity</th>
<th style="padding:6px 8px;text-align:center;">Welfare Retained</th>
<th style="padding:6px 8px;text-align:center;">Member Cost</th>
</tr></thead>
<tbody>
<tr style="border-bottom:1px solid #ddd;"><td style="padding:6px 8px;">Moderate recession</td><td style="text-align:center;font-weight:600;">87%</td><td style="text-align:center;">$284BN</td></tr>
<tr style="background:#f5f1eb;"><td style="padding:6px 8px;">Severe recession</td><td style="text-align:center;font-weight:600;">73%</td><td style="text-align:center;">$238BN</td></tr>
</tbody>
</table>

<p class="callout">Even in a severe recession, the estimated member cost substantially outweighs the estimated insurance benefit under any plausible combination of assumptions. No scenario produces a reversal.</p>

</td>
</tr>
</table>

## Five Questions. Five Answers.

<p class="summary-badge">Summary</p>

```{r summary-table}
library(knitr)
library(kableExtra)
df <- data.frame(
  RQ = c("RQ 1","RQ 2","RQ 3","RQ 4","RQ 5"),
  Question = c(
    "How much capital buffer before the rule?",
    "Did complex CUs hold more capital?",
    "What did it cost members?",
    "Did the rule change what loans CUs make?",
    "Does removing the rule create unacceptable failure risk?"
  ),
  Finding = c(
    "Critically thin -- 72bp average buffer above the 10% threshold",
    "Mixed -- NW ratio +46bp, but well-capitalized probability -3.7pp",
    "Rates rose on every product; lending contracted persistently. $325.9BN total, $865/member/yr",
    "Yes -- auto share -4.0pp, RE share +2.4pp. Pattern unique to RBC rule, not seen in 2008",
    "Real but modest -- +3.4pp failure risk, ~24 institutions, NCUSIF cost $0.4-$2.2BN"
  ),
  stringsAsFactors = FALSE
)
kable(df, col.names=c("RQ","Question","Core Finding and Key Numbers"),
      align=c("c","l","l"), escape=TRUE) |>
  kable_styling(bootstrap_options=c("striped","hover"),
                full_width=TRUE, font_size=13) |>
  column_spec(1, bold=TRUE, width="5%", color="#1a2744") |>
  column_spec(2, width="28%") |>
  column_spec(3, width="67%", bold=TRUE, color="#1a2744") |>
  row_spec(0, background="#1a2744", color="white", bold=TRUE)
```

<p style="font-size:12.5px;color:#444;margin-top:12px;border-left:3px solid #c8972a;padding-left:12px;">
<strong>Across all five research questions</strong>, the findings are consistent: complex credit unions responded to the rule's capital incentive, but the adjustment came primarily through higher loan rates, reduced lending, portfolio reallocation, and lower profitability -- rather than material improvements in financial stability outcomes at the system level.
</p>
