---
title: "The 2022 NCUA Risk-Based Capital Rule"
subtitle: "Empirical Findings from an Analysis of Complex Credit Unions"
author: "Saurabh C. Datta, Ph.D. -- Senior Economist, NCUA Office of the Chief Economist"
date: "April 2026"
output:
  revealjs::revealjs_presentation:
    theme: white
    highlight: pygments
    center: false
    transition: fade
    slide_level: 2
    self_contained: true
    reveal_options:
      width: 1100
      height: 660
      margin: 0.04
      controls: true
      progress: true
      history: true
      keyboard: true
    css: rbc_slides.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, fig.align="center")
fig_include <- function(path, caption="") {
  if (file.exists(path)) { knitr::include_graphics(path) } else {
    cat(paste0('<div class="fig-ph"><span>FIGURE: ',basename(path),'</span></div>'))
  }
}
fig_dir <- "output/paper/figures/"
```

## {data-background-color="#1a2744"}

<table class="sw"><tr><td>
<p class="t-org">NCUA &middot; Office of the Chief Economist</p>
<p class="t-main">The 2022 NCUA Risk-Based Capital Rule</p>
<p class="t-sub">Empirical Findings from an Analysis of Complex Credit Unions</p>
<hr class="t-rule">
<p class="t-meta">Saurabh C. Datta, Ph.D. &middot; Senior Economist<br>NCUA Call Report Data, 2000--2025 &middot; April 2026</p>
</td></tr></table>

## What We Studied -- and How

<table class="sw"><tr><td>
<table class="two-col"><tr>
<td class="col-l">
<p class="lbl">The Data</p>
<ul>
<li><b>696 complex credit unions</b> -- avg assets above $500M, classified using 2021 data</li>
<li><b>~4,488 non-complex credit unions</b> as the comparison group</li>
<li><b>150,000+ CU-quarters</b> of NCUA Call Report data, 2000--2025</li>
<li><b>12 outcomes</b>: capital, loan rates, volume, portfolio mix, profitability, loan quality</li>
</ul>
<hr class="shr">
<p class="lbl">The Method: Difference-in-Differences (DiD)</p>
<table class="mtbl"><tr>
<td class="mbox"><b>Change in complex CUs</b><br><span>Before vs. after rule</span></td>
<td class="mop">&minus;</td>
<td class="mbox"><b>Change in non-complex CUs</b><br><span>Same time window</span></td>
<td class="mop">=</td>
<td class="mres"><b>The Rule's Effect</b><br><span>What changed because of the rule, not the economy</span></td>
</tr></table>
<p class="cbt"><b>Why this works:</b> Non-complex CUs faced the same economic conditions but were not subject to the RBC rule. Differences between the two groups, after accounting for pre-existing trends, isolate the rule's effect.</p>
</td>
<td class="col-div"></td>
<td class="col-r">
<p class="lbl">How to Read the Event Study Charts</p>
<div class="ibox">
<ul>
<li><b>Horizontal axis:</b> quarters relative to the rule's effective date (Q0). Left = before; right = after.</li>
<li><b>Vertical axis:</b> estimated difference between complex and non-complex CUs above any pre-existing gap.</li>
<li><b>Before Q0:</b> dots near zero confirm both groups moved together pre-rule (parallel trends check).</li>
<li><b>After Q0:</b> a persistent move is the rule's estimated effect. Shaded band = 95% confidence interval.</li>
<li><b>Band crosses zero:</b> effect not statistically distinguishable from no effect at that quarter.</li>
</ul>
</div>

```{r schematic, fig.height=2.2, fig.width=3.8, out.width="100%"}
library(ggplot2); library(tibble)
q<-c(-4,-2,0,2,4,6); e<-c(0.04,-0.02,0,-0.44,-0.57,-0.61); s<-c(.10,.10,0,.10,.10,.10)
d<-tibble(q=q,e=e,lo=e-1.96*s,hi=e+1.96*s)
ggplot(d,aes(x=q,y=e))+
  geom_hline(yintercept=0,linetype="dashed",color="#aaa",linewidth=0.5)+
  geom_vline(xintercept=0,linetype="dashed",color="#c8972a",linewidth=0.8)+
  geom_ribbon(aes(ymin=lo,ymax=hi),fill="#1a2744",alpha=0.15)+
  geom_line(color="#1a2744",linewidth=1.2)+
  geom_point(color="#1a2744",size=2.5,fill="white",shape=21,stroke=1.5)+
  annotate("text",x=0.2,y=0.18,label="Rule effective",hjust=0,color="#c8972a",size=2.5)+
  annotate("text",x=-3.8,y=0.18,label="Flat pre-period",hjust=0,color="#555",size=2.4)+
  annotate("text",x=2.0,y=-0.33,label="Persistent effect",hjust=0,color="#1a2744",size=2.5)+
  scale_x_continuous(breaks=q,labels=paste0(ifelse(q>0,"+",""),q,"Q"))+
  labs(x=NULL,y=NULL,caption="Schematic only -- actual charts on following slides")+
  theme_minimal(base_size=9)+theme(
    panel.grid.minor=element_blank(),
    panel.grid.major=element_line(color="#ece8e0",linewidth=0.4),
    plot.caption=element_text(color="#aaa",size=7,hjust=0.5),
    plot.background=element_rect(fill="white",color="#c8c0b0",linewidth=0.5),
    axis.text=element_text(color="#666"))
```

</td>
</tr></table>
</td></tr></table>

## RQ 1 &nbsp;&middot;&nbsp; Pre-Rule Capital Buffer

<table class="sw"><tr><td>
<span class="bdg-rq">Research Question 1</span>
<p class="rq-q"><b>How much capital buffer did complex credit unions hold before the rule?</b></p>
<table class="abar"><tr><td>Answer: The average complex credit union entered the rule with a buffer of just <b>72 basis points</b> above the new 10% threshold -- critically thin.</td></tr></table>
<table class="two-col"><tr>
<td class="col-l">
<p class="lbl">What the Data Show</p>
<ul>
<li>The 696 complex CUs were nominally compliant before the rule -- but with very little margin to absorb a new, higher threshold</li>
<li>A <b>72 basis point buffer</b> means holding 10.72% net worth when the definition of well-capitalized shifted upward simultaneously</li>
<li>Capital distribution was <b>tightly clustered just above the old threshold</b> -- a large mass barely on the compliant side</li>
<li>This thin margin directly explains why the rule had such large downstream effects -- institutions had to act immediately</li>
</ul>
</td>
<td class="col-div"></td>
<td class="col-r">
<p class="cbt"><b>What is a basis point?</b> One bp = one one-hundredth of a percentage point. A 72bp buffer means 10.72% capital against a 10.00% threshold.</p>
<table class="stbl"><tr>
<td class="sc"><div class="sn">72bp</div><div class="sd">Average pre-rule capital buffer above the 10% threshold</div></td>
<td class="sc sc-red"><div class="sn">696</div><div class="sd">Complex CUs subject to the rule (fixed 2021 classification)</div></td>
</tr></table>
<p class="cbt" style="margin-top:8px;"><b>Why it matters:</b> A rule requiring additional capital is always more costly when institutions start with little room to spare. The thin buffer set the stage for the pressures documented in RQ2--RQ4.</p>
</td>
</tr></table>
</td></tr></table>

## RQ 2 &nbsp;&middot;&nbsp; Did Capital Levels Rise?

<table class="sw"><tr><td>
<span class="bdg-rq">Research Question 2</span>
<p class="rq-q"><b>Did the rule cause complex credit unions to hold more capital?</b></p>
<table class="abar"><tr><td>Answer: <b>Mixed.</b> Average capital rose -- but fewer institutions are now meeting the well-capitalized standard. The bar rose faster than capital could accumulate.</td></tr></table>
<table class="two-col"><tr>
<td class="col-l">
<table class="stbl" style="margin-bottom:12px;"><tr>
<td class="sc sc-grn"><div class="sn">+46bp</div><div class="sd">Net worth ratio increase (p &lt; 0.01)</div></td>
<td class="sc sc-red"><div class="sn">-3.7pp</div><div class="sd">Well-capitalized probability change (p &lt; 0.01)</div></td>
</tr></table>
<p class="lbl">The Paradox Explained</p>
<ul>
<li>Capital rose -- institutions responded to the rule's incentive to build reserves</li>
<li>But the <b>definition of well-capitalized changed</b> simultaneously, raising the bar</li>
<li>Result: more capital on average, but <b>fewer institutions clearing the new bar</b></li>
</ul>
<p class="cbt" style="margin-top:8px;"><b>Analogy:</b> Every student studied harder and raised their score -- but the passing grade was raised even more. Average scores up, yet more students are technically failing. Both facts are simultaneously true.</p>
</td>
<td class="col-div"></td>
<td class="col-r">
<p class="lbl">Event Study -- Capital</p>

```{r fig-rq2, out.width="100%", results="asis"}
fig_include(file.path(fig_dir,"Figure3_EventStudy_Capital.png"),"NW ratio, well-cap prob, tier 1, risk-based capital")
```

<p class="fnote">Pre-period flat confirms parallel trends. Post-period divergence attributed to the rule.</p>
</td>
</tr></table>
</td></tr></table>

## RQ 3 &nbsp;&middot;&nbsp; Member Cost: Higher Rates and Less Lending

<table class="sw"><tr><td>
<span class="bdg-rq">Research Question 3</span>
<p class="rq-q"><b>What did the rule cost members in higher loan rates and reduced lending?</b></p>
<table class="abar"><tr><td>Answer: Loan rates rose across every product; lending contracted persistently. Estimated member cost: <b>$325.9 billion over 4 years</b> -- approximately <b>$865 per member per year</b>.</td></tr></table>
<table class="two-col"><tr>
<td class="col-l">
<p class="lbl">Loan Rate Increases vs. Non-Complex CUs</p>
<table class="stbl" style="margin-bottom:8px;"><tr>
<td class="sc sc-red"><div class="sn" style="font-size:20px;">+75bp</div><div class="sd">Mortgage spread</div></td>
<td class="sc sc-red"><div class="sn" style="font-size:20px;">+59bp</div><div class="sd">Auto loan spread</div></td>
<td class="sc sc-red"><div class="sn" style="font-size:20px;">+58bp</div><div class="sd">Commercial spread</div></td>
</tr></table>
<ul>
<li>All increases <b>statistically significant</b>, persist 10+ quarters with no convergence</li>
<li>A 75bp mortgage increase on $300,000 adds roughly <b>$135 per month</b></li>
</ul>

```{r fig-rq3a, out.width="100%", results="asis"}
fig_include(file.path(fig_dir,"Figure4_EventStudy_Spreads.png"),"Loan spread event study")
```

</td>
<td class="col-div"></td>
<td class="col-r">
<p class="lbl">Lending Volume and Total Cost</p>
<table class="stbl" style="margin-bottom:8px;"><tr>
<td class="sc sc-red"><div class="sn" style="font-size:17px;">-0.44pp/qtr</div><div class="sd">Loan growth -- persistent, no recovery</div></td>
<td class="sc sc-red"><div class="sn" style="font-size:18px;">$325.9B</div><div class="sd">Total member welfare cost over 4 years</div></td>
</tr></table>
<ul>
<li>Loan growth fell <b>0.44pp per quarter</b> -- compounding into substantial credit reduction</li>
<li>$325.9BN aggregates the rate effect across all 696 complex CU member-borrowers over 4 years</li>
<li>Per-member: approximately <b>$865/year</b> in higher borrowing costs</li>
</ul>

```{r fig-rq3b, out.width="100%", results="asis"}
fig_include(file.path(fig_dir,"Figure3b_EventStudy_LoanGrowth.png"),"Loan growth event study")
```

</td>
</tr></table>
</td></tr></table>

## RQ 4 &nbsp;&middot;&nbsp; Portfolio Shift and the Crisis Comparison

<table class="sw"><tr><td>
<span class="bdg-rq">Research Question 4</span>
<p class="rq-q"><b>Did the rule change what loans credit unions make -- and how does this compare to past stress?</b></p>
<table class="abar"><tr><td>Answer: Yes. Complex CUs shifted from auto loans toward mortgages -- consistent with <b>risk-weight arbitrage</b>. This shift is unique to the RBC rule; it did not occur during the 2008 financial crisis.</td></tr></table>
<table class="two-col"><tr>
<td class="col-l">
<table class="stbl" style="margin-bottom:8px;"><tr>
<td class="sc sc-red"><div class="sn">-4.0pp</div><div class="sd">Auto loan share (vs. controls)</div></td>
<td class="sc sc-grn"><div class="sn">+2.4pp</div><div class="sd">Real estate loan share</div></td>
</tr></table>
<ul>
<li><b>Risk-weight arbitrage:</b> mortgages carry lower risk weights than auto loans under RBC -- shifting to mortgages reduces required capital, a rational regulatory response</li>
<li>Loan quality deteriorated: <b>delinquency +8bp</b>, driven by a capacity channel -- CUs with the largest profitability declines showed 2x the delinquency increase</li>
</ul>

```{r fig-rq4a, out.width="100%", results="asis"}
fig_include(file.path(fig_dir,"Figure5_Portfolio_Trends.png"),"Portfolio reallocation")
```

</td>
<td class="col-div"></td>
<td class="col-r">
<p class="lbl">Comparison to the 2008 Financial Crisis</p>
<p class="cbt">We ran the same DiD using 2008 as the event, comparing complex CU responses to the financial crisis vs. the RBC rule.</p>
<table style="width:100%;border-collapse:collapse;font-size:12px;margin-top:6px;">
<thead><tr style="background:#1a2744;color:#fff;">
<th style="padding:6px 8px;text-align:left;font-size:11px;">Outcome</th>
<th style="padding:6px 8px;text-align:center;font-size:11px;">2008 Crisis</th>
<th style="padding:6px 8px;text-align:center;font-size:11px;">RBC Rule</th>
</tr></thead>
<tbody>
<tr style="border-bottom:1px solid #ddd;"><td style="padding:5px 8px;">Capital accumulation</td><td style="text-align:center;">Similar</td><td style="text-align:center;">Similar</td></tr>
<tr style="background:#f5f1eb;border-bottom:1px solid #ddd;"><td style="padding:5px 8px;">Auto loan exit</td><td style="text-align:center;color:#888;">Not observed</td><td style="text-align:center;color:#9b2226;font-weight:600;">Persistent</td></tr>
<tr style="border-bottom:1px solid #ddd;"><td style="padding:5px 8px;">Loan growth</td><td style="text-align:center;color:#888;">Temporary</td><td style="text-align:center;color:#9b2226;font-weight:600;">No recovery</td></tr>
<tr style="background:#f5f1eb;"><td style="padding:5px 8px;">Profitability</td><td style="text-align:center;color:#888;">Temporary</td><td style="text-align:center;color:#9b2226;font-weight:600;">No recovery</td></tr>
</tbody>
</table>
<p style="font-size:12px;margin-top:6px;color:#444;">Capital behavior resembles a crisis response -- but lending and profitability effects are <b>distinctly regulatory</b> in nature.</p>
</td>
</tr></table>
</td></tr></table>

## Supporting Finding &nbsp;&middot;&nbsp; Profitability

<table class="sw"><tr><td>
<span class="bdg-intro">Supporting Finding</span>
<p class="rq-q"><b>Profitability Declined Persistently -- Largest Institutions Hit Hardest</b></p>
<table class="abar"><tr><td>Return on assets fell across complex CUs -- with the largest institutions hit hardest. This directly informs the financial stability implications in RQ5.</td></tr></table>
<table class="two-col"><tr>
<td class="col-l">
<p class="lbl">Key Estimates</p>
<table class="stbl" style="margin-bottom:10px;"><tr>
<td class="sc sc-red"><div class="sn">-26bp</div><div class="sd">Average ROA decline -- all complex CUs (p &lt; 0.01)</div></td>
<td class="sc sc-red"><div class="sn">-47bp</div><div class="sd">ROA decline for CUs above $10B in assets</div></td>
</tr></table>
<ul>
<li>ROA decline is <b>persistent</b> -- no evidence of adaptation or recovery through the observation window</li>
<li>Largest CUs (above $10B) experience a decline <b>nearly double the average</b></li>
<li>CCULR provided <b>no significant relief</b> on lending outcomes; CCULR adopters showed worse baseline outcomes consistent with distressed self-selection</li>
</ul>
</td>
<td class="col-div"></td>
<td class="col-r">
<p class="cbt"><b>Lower profitability reduces organic capital-building ability.</b> A CU that cannot retain earnings must raise rates, cut lending, or draw down its capital ratio -- all three patterns appear in the data.</p>
<p class="cbt" style="margin-top:8px;"><b>CCULR finding:</b> The Community Credit Union Leverage Ratio was designed as a simpler compliance pathway. Institutions most likely to adopt it were already under stress -- CCULR adoption did not cause their worse outcomes, but did not reverse them either.</p>

```{r fig-roa, out.width="100%", results="asis"}
fig_include(file.path(fig_dir,"Figure3d_EventStudy_ROA.png"),"ROA event study")
```

</td>
</tr></table>
</td></tr></table>

## RQ 5 &nbsp;&middot;&nbsp; Failure Risk from Removing the Rule

<table class="sw"><tr><td>
<span class="bdg-rq">Research Question 5</span>
<p class="rq-q"><b>Does removing the RBC rule create unacceptable failure risk?</b></p>
<table class="abar"><tr><td>Answer: Removing the rule carries <b>real but quantifiable and modest risk</b>. At 2008 GFC severity, the rule reduces failure risk by 3.4 percentage points -- protecting approximately 24 institutions.</td></tr></table>
<table class="two-col"><tr>
<td class="col-l">
<p class="lbl">Capital Adequacy Stress Test -- At 2008 GFC Severity</p>
<table class="stbl" style="margin-bottom:10px;"><tr>
<td class="sc"><div class="sn" style="font-size:17px;">19.1%</div><div class="sd">Pre-rule failure rate (baseline)</div></td>
<td class="sc sc-grn"><div class="sn" style="font-size:17px;">19.4%</div><div class="sd">With-rule rate (-0.3pp incremental protection)</div></td>
<td class="sc sc-red"><div class="sn" style="font-size:17px;">22.5%</div><div class="sd">Post-removal (+3.4pp, ~24 institutions)</div></td>
</tr></table>
<ul>
<li>Rule's <b>incremental protection is only -0.3pp</b> -- complex CUs were already building buffers before the rule</li>
<li>Removal raises simulated failure risk by <b>+3.4pp</b>, corresponding to <b>~24 institutions</b></li>
<li>Each protected institution represents approximately <b>$14 billion in assets</b></li>
</ul>
</td>
<td class="col-div"></td>
<td class="col-r">
<table class="stbl" style="margin-bottom:10px;"><tr>
<td class="sc"><div class="sn" style="font-size:15px;">$0.4B--$2.2B</div><div class="sd">Probability-weighted NCUSIF cost of removal</div></td>
</tr></table>
<p class="lbl">Stress Scenario Sensitivity</p>
<table style="width:100%;border-collapse:collapse;font-size:12px;margin-bottom:8px;">
<thead><tr style="background:#1a2744;color:#fff;">
<th style="padding:5px 8px;text-align:left;font-size:11px;">Recession Severity</th>
<th style="padding:5px 8px;text-align:center;font-size:11px;">Welfare Retained</th>
<th style="padding:5px 8px;text-align:center;font-size:11px;">Member Cost</th>
</tr></thead>
<tbody>
<tr style="border-bottom:1px solid #ddd;"><td style="padding:5px 8px;">Moderate recession</td><td style="text-align:center;font-weight:600;">87%</td><td style="text-align:center;">$284BN</td></tr>
<tr style="background:#f5f1eb;"><td style="padding:5px 8px;">Severe recession</td><td style="text-align:center;font-weight:600;">73%</td><td style="text-align:center;">$238BN</td></tr>
</tbody>
</table>
<p class="cbt">Even in a severe recession, the estimated member cost substantially outweighs the estimated insurance benefit under any plausible combination of assumptions. No scenario produces a reversal.</p>
</td>
</tr></table>
</td></tr></table>

## Five Questions. Five Answers.

<table class="sw"><tr><td>
<span class="bdg-sum">Summary</span>

```{r summary-table}
library(knitr); library(kableExtra)
df <- data.frame(
  RQ=c("RQ 1","RQ 2","RQ 3","RQ 4","RQ 5"),
  Question=c(
    "How much capital buffer before the rule?",
    "Did complex CUs hold more capital?",
    "What did it cost members?",
    "Did the rule change what loans CUs make?",
    "Does removing the rule create unacceptable failure risk?"
  ),
  Finding=c(
    "Critically thin -- 72bp average buffer above the 10% threshold",
    "Mixed -- NW ratio +46bp, but well-capitalized probability -3.7pp",
    "Rates rose on every product; lending contracted persistently. $325.9BN total, $865/member/yr",
    "Yes -- auto share -4.0pp, RE share +2.4pp. Pattern unique to RBC rule, not seen in 2008",
    "Real but modest -- +3.4pp failure risk, ~24 institutions, NCUSIF cost $0.4-$2.2BN"
  ),
  stringsAsFactors=FALSE
)
kable(df,col.names=c("RQ","Question","Core Finding and Key Numbers"),
      align=c("c","l","l"),escape=TRUE) |>
  kable_styling(bootstrap_options=c("striped","hover"),full_width=TRUE,font_size=13) |>
  column_spec(1,bold=TRUE,width="5%",color="#1a2744") |>
  column_spec(2,width="28%") |>
  column_spec(3,width="67%",bold=TRUE,color="#1a2744") |>
  row_spec(0,background="#1a2744",color="white",bold=TRUE)
```

<p style="font-size:12.5px;color:#444;margin-top:10px;border-left:3px solid #c8972a;padding-left:10px;"><b>Across all five research questions</b>, the findings are consistent: complex credit unions responded to the rule's capital incentive primarily through higher loan rates, reduced lending, portfolio reallocation, and lower profitability -- rather than material improvements in financial stability outcomes at the system level.</p>

</td></tr></table>
