---
title: "The 2022 NCUA Risk-Based Capital Rule"
subtitle: "Empirical Findings from an Analysis of Complex Credit Unions"
author: "NCUA Office of the Chief Economist"
date: "April 2026"
output:
  ioslides_presentation:
    widescreen: true
    smaller: false
    transition: faster
    css: rbc_ios.css
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
fig_dir <- "output/paper/figures/"
fig_include <- function(path, caption="") {
  if (file.exists(path)) {
    knitr::include_graphics(path)
  } else {
    cat(paste0('<div class="fp"><span>', basename(path), '</span></div>'))
  }
}
```

## {.title-slide}

<div class="title-block">
<div class="title-org">NCUA &middot; Office of the Chief Economist</div>
<div class="title-main">The 2022 NCUA Risk-Based Capital Rule</div>
<div class="title-sub">Empirical Findings from an Analysis of Complex Credit Unions</div>
<div class="title-rule"></div>
<div class="title-meta">Saurabh C. Datta, Ph.D. &middot; Senior Economist<br>NCUA Call Report Data, 2000&ndash;2025 &middot; April 2026</div>
</div>

## What We Studied | and How

<div class="two-col">
<div class="col-l">
<div class="lbl">The Data</div>
<ul>
<li><b>696 complex credit unions</b> &mdash; avg assets above $500M, classified using 2021 data</li>
<li><b>~4,488 non-complex credit unions</b> as the comparison group</li>
<li><b>150,000+ CU-quarters</b> of NCUA Call Report data, 2000&ndash;2025</li>
<li><b>12 outcomes</b>: capital, loan rates, volume, portfolio mix, profitability, loan quality</li>
</ul>
<div class="lbl" style="margin-top:10px;">The Method: Difference-in-Differences (DiD)</div>
<div class="method-row">
<div class="mbox"><b>Change in complex CUs</b><br>Before vs. after rule</div>
<div class="mop">&minus;</div>
<div class="mbox"><b>Change in non-complex CUs</b><br>Same time window</div>
<div class="mop">=</div>
<div class="mres"><b>The Rule's Effect</b><br>What changed because of the rule, not the economy</div>
</div>
<div class="cb"><b>Why this works:</b> Non-complex CUs faced the same economic conditions but were not subject to the RBC rule. Any difference between the two groups, after accounting for pre-existing trends, isolates the rule's effect.</div>
</div>
<div class="col-div"></div>
<div class="col-r">
<div class="lbl">How to Read the Event Study Charts</div>
<div class="ib">
<ul>
<li><b>Horizontal axis:</b> quarters relative to the rule's effective date (Q0). Left = before; right = after.</li>
<li><b>Vertical axis:</b> estimated gap between complex and non-complex CUs above any pre-existing difference.</li>
<li><b>Before Q0:</b> dots near zero confirm both groups moved together &mdash; the parallel trends check.</li>
<li><b>After Q0:</b> a persistent move is the rule's estimated effect. Shaded band = 95% confidence interval.</li>
<li><b>If the band crosses zero</b> at a given quarter, the effect is not statistically distinguishable from no effect.</li>
</ul>
</div>
<svg viewBox="0 0 280 95" style="width:100%;margin-top:6px;border:1px solid #ddd;border-radius:3px;background:white;">
<line x1="18" y1="52" x2="270" y2="52" stroke="#ccc" stroke-width="0.8" stroke-dasharray="3,2"/>
<line x1="145" y1="8" x2="145" y2="86" stroke="#c8972a" stroke-width="1.2" stroke-dasharray="3,2"/>
<polygon points="55,47 100,49 145,52 188,40 230,35 270,33 270,45 230,47 188,52 145,60 100,57 55,59" fill="#1a2744" opacity="0.12"/>
<polyline points="55,48 100,50 145,52 188,41 230,36 270,34" fill="none" stroke="#1a2744" stroke-width="1.8"/>
<circle cx="55" cy="48" r="2.8" fill="#1a2744"/><circle cx="100" cy="50" r="2.8" fill="#1a2744"/>
<circle cx="145" cy="52" r="2.8" fill="#1a2744"/><circle cx="188" cy="41" r="2.8" fill="#1a2744"/>
<circle cx="230" cy="36" r="2.8" fill="#1a2744"/><circle cx="270" cy="34" r="2.8" fill="#1a2744"/>
<text x="55" y="94" font-size="7.5" text-anchor="middle" fill="#888" font-family="Arial">-4Q</text>
<text x="100" y="94" font-size="7.5" text-anchor="middle" fill="#888" font-family="Arial">-2Q</text>
<text x="145" y="94" font-size="7.5" text-anchor="middle" fill="#888" font-family="Arial">0</text>
<text x="188" y="94" font-size="7.5" text-anchor="middle" fill="#888" font-family="Arial">+2Q</text>
<text x="230" y="94" font-size="7.5" text-anchor="middle" fill="#888" font-family="Arial">+4Q</text>
<text x="270" y="94" font-size="7.5" text-anchor="middle" fill="#888" font-family="Arial">+6Q</text>
<text x="148" y="18" font-size="7" fill="#c8972a" font-family="Arial">Rule effective</text>
<text x="57" y="42" font-size="7" fill="#555" font-family="Arial">Flat pre-period</text>
<text x="158" y="36" font-size="7" fill="#1a2744" font-family="Arial">Persistent effect</text>
</svg>
<p style="font-size:10px;color:#aaa;text-align:center;margin-top:3px;">Schematic only &mdash; actual charts on following slides</p>
</div>
</div>

## RQ 1 | Pre-Rule Capital Buffer

<div class="rq-badge">Research Question 1</div>
<div class="rq-q">How much capital buffer did complex credit unions hold before the rule?</div>
<div class="abar">Answer: The average complex credit union entered the rule with a buffer of just <b>72 basis points</b> above the new 10% threshold &mdash; critically thin.</div>

<div class="two-col">
<div class="col-l">
<div class="lbl">What the Data Show</div>
<ul>
<li>The 696 complex CUs were nominally compliant before the rule &mdash; but with very little margin to absorb a new, higher threshold</li>
<li>A <b>72bp buffer</b> means holding 10.72% net worth when the definition of well-capitalized shifted upward simultaneously</li>
<li>Capital was <b>tightly clustered just above the old threshold</b> &mdash; a large mass barely on the compliant side</li>
<li>This thin margin directly explains the large downstream effects &mdash; institutions had to act immediately</li>
</ul>
</div>
<div class="col-div"></div>
<div class="col-r">
<div class="cb"><b>What is a basis point?</b> One bp = one one-hundredth of a percentage point. A 72bp buffer means 10.72% capital against a 10.00% threshold.</div>
<div class="stat-row">
<div class="stat-card"><div class="stat-num">72bp</div><div class="stat-lbl">Average pre-rule buffer above the 10% threshold</div></div>
<div class="stat-card red"><div class="stat-num">696</div><div class="stat-lbl">Complex CUs subject to the rule (fixed 2021 classification)</div></div>
</div>
<div class="cb" style="margin-top:8px;"><b>Why it matters:</b> A rule requiring additional capital is always more costly when institutions start with little room to spare. The thin buffer set the stage for the pressures in RQ2&ndash;RQ4.</div>
</div>
</div>

## RQ 2 | Did Capital Levels Rise?

<div class="rq-badge">Research Question 2</div>
<div class="rq-q">Did the rule cause complex credit unions to hold more capital?</div>
<div class="abar">Answer: <b>Mixed.</b> Average capital rose &mdash; but fewer institutions are now meeting the well-capitalized standard. The bar rose faster than capital could accumulate.</div>

<div class="two-col">
<div class="col-l">
<div class="stat-row" style="margin-bottom:12px;">
<div class="stat-card green"><div class="stat-num">+46bp</div><div class="stat-lbl">Net worth ratio increase (p &lt; 0.01)</div></div>
<div class="stat-card red"><div class="stat-num">-3.7pp</div><div class="stat-lbl">Well-capitalized probability change (p &lt; 0.01)</div></div>
</div>
<div class="lbl">The Paradox Explained</div>
<ul>
<li>Capital rose &mdash; institutions responded to the rule's incentive to build reserves</li>
<li>But the <b>definition of well-capitalized changed</b> simultaneously, raising the bar further</li>
<li>Result: more capital on average, but <b>fewer institutions clearing the new bar</b></li>
</ul>
<div class="cb" style="margin-top:8px;"><b>Analogy:</b> Every student studied harder and raised their score &mdash; but the passing grade was also raised. Average scores went up, yet more students are technically failing. Both facts are simultaneously true.</div>
</div>
<div class="col-div"></div>
<div class="col-r">
<div class="lbl">Event Study &mdash; Capital</div>
```{r fig-rq2, out.width="100%", results="asis"}
fig_include(file.path(fig_dir,"Figure3_EventStudy_Capital.png"))
```
<p class="fn">Pre-period flat confirms parallel trends. Post-period divergence attributed to the rule.</p>
</div>
</div>

## RQ 3 | Member Cost: Higher Rates and Less Lending

<div class="rq-badge">Research Question 3</div>
<div class="rq-q">What did the rule cost members in higher loan rates and reduced lending?</div>
<div class="abar">Answer: Loan rates rose across every product; lending contracted persistently. Estimated cost: <b>$325.9 billion over 4 years</b> &mdash; approximately <b>$865 per member per year</b>.</div>

<div class="two-col">
<div class="col-l">
<div class="lbl">Loan Rate Increases vs. Non-Complex CUs</div>
<div class="stat-row" style="margin-bottom:8px;">
<div class="stat-card red"><div class="stat-num sm">+75bp</div><div class="stat-lbl">Mortgage spread</div></div>
<div class="stat-card red"><div class="stat-num sm">+59bp</div><div class="stat-lbl">Auto loan spread</div></div>
<div class="stat-card red"><div class="stat-num sm">+58bp</div><div class="stat-lbl">Commercial spread</div></div>
</div>
<ul>
<li>All increases <b>statistically significant</b>, persist 10+ quarters with no convergence</li>
<li>A 75bp mortgage increase on $300,000 adds roughly <b>$135 per month</b></li>
</ul>
```{r fig-rq3a, out.width="100%", results="asis"}
fig_include(file.path(fig_dir,"Figure4_EventStudy_Spreads.png"))
```
</div>
<div class="col-div"></div>
<div class="col-r">
<div class="lbl">Lending Volume and Total Cost</div>
<div class="stat-row" style="margin-bottom:8px;">
<div class="stat-card red"><div class="stat-num sm">-0.44pp</div><div class="stat-lbl">Quarterly loan growth &mdash; no recovery</div></div>
<div class="stat-card red"><div class="stat-num sm">$325.9B</div><div class="stat-lbl">Total member welfare cost over 4 years</div></div>
</div>
<ul>
<li>Loan growth fell <b>0.44pp per quarter</b> &mdash; compounding into substantial credit reduction</li>
<li>$325.9BN aggregates the rate effect across all member-borrowers of all 696 complex CUs over 4 years</li>
<li>Per-member: approximately <b>$865/year</b> in higher borrowing costs</li>
</ul>
```{r fig-rq3b, out.width="100%", results="asis"}
fig_include(file.path(fig_dir,"Figure3b_EventStudy_LoanGrowth.png"))
```
</div>
</div>

## RQ 4 | Portfolio Shift and the Crisis Comparison

<div class="rq-badge">Research Question 4</div>
<div class="rq-q">Did the rule change what loans credit unions make &mdash; and how does this compare to past stress?</div>
<div class="abar">Answer: Yes. Complex CUs shifted from auto loans toward mortgages &mdash; consistent with <b>risk-weight arbitrage</b>. This shift is unique to the RBC rule; it did not occur during the 2008 financial crisis.</div>

<div class="two-col">
<div class="col-l">
<div class="stat-row" style="margin-bottom:8px;">
<div class="stat-card red"><div class="stat-num">-4.0pp</div><div class="stat-lbl">Auto loan share (vs. controls)</div></div>
<div class="stat-card green"><div class="stat-num">+2.4pp</div><div class="stat-lbl">Real estate loan share</div></div>
</div>
<ul>
<li><b>Risk-weight arbitrage:</b> mortgages carry lower risk weights than auto loans under RBC &mdash; shifting to mortgages reduces required capital, a rational regulatory response</li>
<li>Loan quality deteriorated: <b>delinquency +8bp</b>, driven by a capacity channel &mdash; CUs with the largest profitability declines showed 2x the delinquency increase</li>
</ul>
```{r fig-rq4a, out.width="100%", results="asis"}
fig_include(file.path(fig_dir,"Figure5_Portfolio_Trends.png"))
```
</div>
<div class="col-div"></div>
<div class="col-r">
<div class="lbl">Comparison to the 2008 Financial Crisis</div>
<div class="cb">We ran the same DiD using 2008 as the event, comparing responses to the financial crisis vs. the RBC rule.</div>
<table class="ct" style="margin-top:6px;">
<thead><tr><th>Outcome</th><th>2008 Crisis</th><th>RBC Rule</th></tr></thead>
<tbody>
<tr><td>Capital accumulation</td><td>Similar</td><td>Similar</td></tr>
<tr><td>Auto loan exit</td><td class="dim">Not observed</td><td class="bad">Persistent</td></tr>
<tr><td>Loan growth</td><td class="dim">Temporary</td><td class="bad">No recovery</td></tr>
<tr><td>Profitability</td><td class="dim">Temporary</td><td class="bad">No recovery</td></tr>
</tbody>
</table>
<p style="font-size:12px;margin-top:6px;color:#444;">Capital behavior resembles a crisis response &mdash; but lending and profitability effects are <b>distinctly regulatory</b> in nature.</p>
</div>
</div>

## Supporting Finding | Profitability

<div class="rq-badge intro">Supporting Finding</div>
<div class="rq-q">Profitability Declined Persistently &mdash; Largest Institutions Hit Hardest</div>
<div class="abar">Return on assets fell across complex CUs &mdash; with the largest institutions hit hardest. This directly informs the financial stability implications in RQ5.</div>

<div class="two-col">
<div class="col-l">
<div class="stat-row" style="margin-bottom:10px;">
<div class="stat-card red"><div class="stat-num">-26bp</div><div class="stat-lbl">Average ROA decline &mdash; all complex CUs (p &lt; 0.01)</div></div>
<div class="stat-card red"><div class="stat-num">-47bp</div><div class="stat-lbl">ROA decline for CUs above $10B in assets</div></div>
</div>
<ul>
<li>ROA decline is <b>persistent</b> &mdash; no evidence of adaptation or recovery through the observation window</li>
<li>Largest CUs (above $10B) experience a decline <b>nearly double the average</b></li>
<li>CCULR provided <b>no significant relief</b> on lending outcomes; CCULR adopters showed worse baseline outcomes consistent with distressed self-selection</li>
</ul>
</div>
<div class="col-div"></div>
<div class="col-r">
<div class="cb"><b>Lower profitability reduces organic capital-building ability.</b> A CU that cannot retain earnings must raise rates, cut lending, or draw down its capital ratio &mdash; all three patterns appear in the data.</div>
<div class="cb" style="margin-top:8px;"><b>CCULR finding:</b> The Community Credit Union Leverage Ratio was designed as a simpler compliance pathway. Institutions most likely to adopt it were already under stress &mdash; CCULR adoption did not cause their worse outcomes, but did not reverse them either.</div>
```{r fig-roa, out.width="100%", results="asis"}
fig_include(file.path(fig_dir,"Figure3d_EventStudy_ROA.png"))
```
</div>
</div>

## RQ 5 | Failure Risk from Removing the Rule

<div class="rq-badge">Research Question 5</div>
<div class="rq-q">Does removing the RBC rule create unacceptable failure risk?</div>
<div class="abar">Answer: Removing the rule carries <b>real but quantifiable and modest risk</b>. At 2008 GFC severity, the rule reduces failure risk by 3.4 percentage points &mdash; protecting approximately 24 institutions.</div>

<div class="two-col">
<div class="col-l">
<div class="lbl">Capital Adequacy Stress Test &mdash; At 2008 GFC Severity</div>
<div class="stat-row" style="margin-bottom:10px;">
<div class="stat-card"><div class="stat-num sm">19.1%</div><div class="stat-lbl">Pre-rule failure rate (baseline)</div></div>
<div class="stat-card green"><div class="stat-num sm">19.4%</div><div class="stat-lbl">With-rule (&minus;0.3pp incremental protection)</div></div>
<div class="stat-card red"><div class="stat-num sm">22.5%</div><div class="stat-lbl">Post-removal (+3.4pp, ~24 institutions)</div></div>
</div>
<ul>
<li>Rule's <b>incremental protection is only &minus;0.3pp</b> &mdash; complex CUs were already building buffers before the rule</li>
<li>Removal raises simulated failure risk by <b>+3.4pp</b>, corresponding to <b>~24 institutions</b></li>
<li>Each protected institution represents approximately <b>$14 billion in assets</b></li>
</ul>
</div>
<div class="col-div"></div>
<div class="col-r">
<div class="stat-row" style="margin-bottom:10px;">
<div class="stat-card"><div class="stat-num sm">$0.4B&ndash;$2.2B</div><div class="stat-lbl">Probability-weighted NCUSIF cost of removal</div></div>
</div>
<div class="lbl">Stress Scenario Sensitivity</div>
<table class="ct">
<thead><tr><th>Recession Severity</th><th style="text-align:center;">Welfare Retained</th><th style="text-align:center;">Member Cost</th></tr></thead>
<tbody>
<tr><td>Moderate recession</td><td style="text-align:center;font-weight:600;">87%</td><td style="text-align:center;">$284BN</td></tr>
<tr><td>Severe recession</td><td style="text-align:center;font-weight:600;">73%</td><td style="text-align:center;">$238BN</td></tr>
</tbody>
</table>
<div class="cb" style="margin-top:8px;">Even in a severe recession, the estimated member cost substantially outweighs the estimated insurance benefit under any plausible combination of assumptions. No scenario produces a reversal.</div>
</div>
</div>

## Five Questions. Five Answers.

<div class="rq-badge sum">Summary</div>

<table class="kbl">
<thead><tr><th>RQ</th><th>Question</th><th>Core Finding and Key Numbers</th></tr></thead>
<tbody>
<tr><td class="rqn">RQ 1</td><td>How much capital buffer before the rule?</td><td>Critically thin &mdash; 72bp average buffer above the 10% threshold</td></tr>
<tr><td class="rqn">RQ 2</td><td>Did complex CUs hold more capital?</td><td>Mixed &mdash; NW ratio +46bp, but well-capitalized probability &minus;3.7pp</td></tr>
<tr><td class="rqn">RQ 3</td><td>What did it cost members?</td><td>Rates rose on every product; lending contracted persistently. $325.9BN total, $865/member/yr</td></tr>
<tr><td class="rqn">RQ 4</td><td>Did the rule change what loans CUs make?</td><td>Yes &mdash; auto share &minus;4.0pp, RE share +2.4pp. Pattern unique to RBC rule, not seen in 2008</td></tr>
<tr><td class="rqn">RQ 5</td><td>Does removing the rule create unacceptable failure risk?</td><td>Real but modest &mdash; +3.4pp failure risk, ~24 institutions, NCUSIF cost $0.4&ndash;$2.2BN</td></tr>
</tbody>
</table>
<p style="font-size:12px;color:#444;margin-top:10px;border-left:3px solid #c8972a;padding-left:10px;"><b>Across all five research questions</b>, the findings are consistent: complex credit unions responded to the rule's capital incentive primarily through higher loan rates, reduced lending, portfolio reallocation, and lower profitability &mdash; rather than through material improvements in financial stability outcomes at the system level.</p>
